---
title: "2024/2025 Drought-net"
author: "Janelle Bohey"
date: "2026-01-27"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Packages

```{r}
library(bipartite)
library(vegan)
library(phytools)
library(taxize)
library(brranching)
library(rotl)
library(dendextend)
library(tidyverse)
library(heatmaply)
library(googlesheets4)
library(broom)
library(dplyr)
library(glmmTMB)


library(bipartite)
library(vegan)
library(phytools)
library(taxize)
library(rotl)
library(dendextend)
library(tidyverse)
library(heatmaply)
library(googlesheets4)
library(tibble)
library(lme4)
library(lmerTest)
library(broom)
library(tidyverse)
library(tidyr)
library(vegan)
library(dplyr)
library(broom.mixed)
library(car)
library(emmeans)
library(knitr)
```

#color pallate 
```{r}

treatment_colors <- c(
  "control" = "#9bd2ef",
  "rainout" = "#ce9a9b"
)

year_colors <- c(
  "2024" = "#4D4D4D",
  "2025" = "#A6A6A6"
)

species_colors <- c(
  "#1B9E77",
  "#D95F02",
  "#7570B3",
  "#E6AB02",
  "#66A61E",
  "#A6761D"
)

palette_all <- c(
  # Treatments
  "control" = "#0072B2",
  "rainout" = "#D55E00",
  # Plant species
  "species1" = "#009E73",
  "species2" = "#F0E442",
  "species3" = "#CC79A7",
  "species4" = "#56B4E9",
  "species5" = "#E69F00",
  "species6" = "#999999",
  # Insect groups
  "group1" = "#882255",
  "group2" = "#44AA99",
  "group3" = "#DDCC77",
  "group4" = "#117733"
)

palette_all <- c(
  # Treatments
  "control" = "#0072B2",
  "rainout" = "#D55E00",
  # Plant species
  "species1" = "#009E73",
  "species2" = "#F0E442",
  "species3" = "#CC79A7",
  "species4" = "#56B4E9",
  "species5" = "#DDAA33",
  "species6" = "#999999",
  # Insect groups
  "group1" = "#661100",
  "group2" = "#882255",
  "group3" = "#44AA99",
  "group4" = "#117733"
)

palette_new <- c(
  # Treatments
  "control" = "#0055A4",   # strong blue
  "rainout" = "#CC3300",   # red-orange
  # Plant species
  "species1" = "#FFCC00",  # bright yellow
  "species2" = "#00CC99",  # cyan-teal
  "species3" = "#FF66CC",  # pink
  "species4" = "#9966FF",  # purple
  "species5" = "#FF9933",  # orange
  "species6" = "#666666",  # medium gray
  # Insect groups
  "group1" = "#990000",    # dark red
  "group2" = "#009966",    # dark teal
  "group3" = "#CC9966",    # brown/tan
  "group4" = "#3333CC"     # dark blue
)

palette_gradient <- c(
  # Treatments
  "control" = "#0055A4",
  "rainout" = "#CC3300",
  # Plant species
  "flag" = "#330066",
  "congesta" = "#660099",
  "umbellatum" = "#9933CC",
  "potentilla" = "#CC66FF",
  "heterotheca" = "#DDAAFF",
  "heliomeris" = "#EEDCFF",
  # Insect groups
  "bombus" = "#00441B",
  "sol_bee" = "#006D2C",
  "fly" = "#238B45",
  "lepidoptera" = "#66C2A4"
)





```

```{r}
# Plants: 6-level purple gradient
#plants.color <- palette_gradient[grep("species", names(palette_gradient))]

# Insects: 4-level green gradient
#insects.color <- palette_gradient[grep("group", names(palette_gradient))]

```

#finds gradient of colors
```{r}
# library(ggplot2)
# library(colorspace)
# 
# # Function to generate gradient with perceptually uniform steps
# generate_contrast_gradient <- function(start_hex, end_hex, n) {
#   # colorRampPalette returns hex directly, using Lab space for perceptual uniformity
#   gradient <- colorRampPalette(c(start_hex, end_hex), space = "Lab")(n)
#   return(gradient)
# }
# 
# # Plant species: purple gradient (6 levels, dark → light)
# insect.color <- generate_contrast_gradient("#330066", "#CC99FF", 4)
# names(insects) <- paste0("species", 1:4)
# 
# # Insect groups: green gradient (4 levels, dark → light)
# plant.color <- generate_contrast_gradient("#00441B", "#66C2A4", 6)
# names(plants) <- paste0("group", 1:6)
# 
# # Treatments: fixed high-contrast colors
# treatments <- c("control" = "#0055A4", "rainout" = "#CC3300")
# 
# # Combine into full palette
# palette_contrast <- c(treatments, plant.color, insect.color)
# 
# # Visualize the palette
# colors.df <- data.frame(
#   category = names(palette_contrast),
#   color = palette_contrast,
#   x = 1:length(palette_contrast)
# )
# 
# ggplot(colors.df, aes(x = x, y = 1, fill = color)) +
#   geom_tile() +
#   scale_fill_identity() +
#   geom_text(aes(label = category), vjust = -0.5, size = 4) +
#   theme_void() +
#   theme(plot.margin = margin(10, 10, 10, 10))
```


#rainout treatments
```{r}
#treatments 
#loads in plot and transect codes paired with treatments (rainout, control)
rainout.treatments <- read_sheet("1YiOm8el0IGCvUHm8xDRxf8U9IWiwI8uNT6KGCzZP2sY") %>% 
                  mutate(transectplot = as.character(transectplot))
```

#soil moisture
```{r}
vwc.2024 <- read_sheet("1pMNaVLHJ2d5SBAP25xk5DjTMnbX1yAmYyWLtiJ_bQlw", na=c("","rocky")) %>% 
                  dplyr::select(-transect_plot) %>% 
                  mutate(plot = str_to_lower(plot), #plot to lowercase  
                  transectplot = paste0(transect, plot)) %>%     # combine transect + plot
  
                mutate(top_VWC=as.numeric(top_VWC), bottom_VWC=as.numeric(bottom_VWC)) %>%
                pivot_longer(c(top_VWC, bottom_VWC), names_to="position", values_to="VWC") %>% 
                group_by(plot, transect, transectplot, treatment, date) %>% 
                summarize(VWC = mean(VWC, na.rm=TRUE), .groups = "drop") %>% 
                filter(!is.nan(VWC)) %>%  #removes rows with no VWC measurements
                mutate(year = 2024)

vwc.2025 <- read_sheet("1WE6tQZkowFBW0A5h5wrFIRIXZ3p4bs5qPBXu1PPG5XU") %>% 
                  drop_na(vwc) %>% 
                  pivot_longer(vwc, names_to="position", values_to="VWC") %>%      
                  group_by(plot, transect, transectplot, date) %>% 
 
                  summarize(VWC = mean(VWC, na.rm=TRUE)) %>% 
                  left_join(rainout.treatments) %>% 
                  mutate(year = 2025)

vwc.all <- bind_rows(vwc.2024, vwc.2025) %>% 
              mutate(VWC = round(VWC, 2)) %>% 
              mutate(year = as.factor(year)) %>% 
              mutate(date = as.Date(date)) %>% 
              mutate(calendar_week = isoweek(date)) %>% 
              mutate(calendar_week = as.numeric(round(calendar_week))) %>%
              mutate(year_week = interaction(year, 
                                        calendar_week, drop = TRUE))
  
  
  
  #TO make sure 2024 only includes weeks 27-33:
              #Filter data to include only weeks actually have data for 2024
              #removes rows from the dataset where the year is 2024 and the calendar_week is not in the specified range (27-33)
             # filter(!(year == "2024" & !(calendar_week %in% c("27", "28", "29", "30", "31", "32", "33"))))

                          
             

vwc.all$calendar_week <- droplevels(vwc.all$calendar_week)

```

```{r}
#fixing calendar_ week issue: 
# 2024 & 2025 are missing Week 28...
#2025 also missing week 32

 vwc.all %>%
     group_by(year) %>%
     summarize(
         unique_weeks = n_distinct(calendar_week),
         weeks_list = paste(sort(unique(calendar_week)), collapse = ", ")
     )

sort(unique(vwc.all$date))

unique(vwc.all$calendar_week[vwc.all$year == 2024])
unique(vwc.all$calendar_week[vwc.all$year == 2025])

# vwc.all %>%
#   group_by(year, calendar_week) %>%
#   summarize(mean_VWC = mean(VWC, na.rm = TRUE), .groups = "drop")

levels(vwc.all$calendar_week)
```


##vwc models
```{r}
#Treatment by year models:
vwc.lmer.2025 <- lmer(VWC~treatment + (1|date) + (1|transectplot) , data = vwc.2025)
anova(vwc.lmer.2025) 

vwc.lmer.2024 <- lmer(VWC~treatment + (1|date) + (1|transectplot) , data = vwc.2024)
Anova(vwc.lmer.2024) 

#Cross Year Model:
vwc.lmer.all.years <- lmer(VWC ~ year + (1|transectplot)  , data = vwc.all)
anova(vwc.lmer.all.years)



#Full model (lmer):
vwc.lmer.all.years.treatment <- lmer(VWC ~  year * treatment  * poly(calendar_week,2,raw=TRUE) + (1|transectplot), data = vwc.all) #add calendar date

vwc.lmer.all.years.treatment <- lmer(VWC ~  year * treatment  * calendar_week  + (1|transectplot), data = vwc.all) #add calendar date

Anova(vwc.lmer.all.years.treatment)

#new 
newdat<-data.frame(expand.grid(year=as.factor(c(2024,2025)),calendar_week=(seq(from=22,to=33,by=1)),treatment=as.factor(c('control','rainout')),transectplot=NA))

newdat$pred<-predict(vwc.lmer.all.years.treatment,newdata=newdat,re.form=NA)
ggplot(aes(x=calendar_week,y=pred,color=treatment),data=newdat)+geom_line()+geom_point(aes(y=VWC),data=vwc.all)+facet_wrap(.~year)

# test24<-data.frame(calendar_week=seq(27,33,1),year=rep(as.factor(2024))
# test25<-data.frame(calendar_week=seq(22,33,1),year=as.factor(2025))
# testemmeans<-as.list(rbind(test24,test25))

emmeans(vwc.lmer.all.years.treatment,~treatment|calendar_week*year,at=list(cal_weeks))

emtrends(vwc.lmer.all.years.treatment,pairwise~year*treatment,var="calendar_week")

#---------------------

#vwc emmeans

cal_weeks <- sort(unique(vwc.all$calendar_week))

# Estimated marginal means by year
vwc.emm <- emmeans(vwc.lmer.all.years.treatment, ~ treatment | year )
summary(vwc.emm)

#do separate emmeans for each year then compare across years


#EMMeans for every treatment × week combination within each year 
vwc.emm.week <- emmeans(vwc.lmer.all.years.treatment, ~ treatment * calendar_week | year, at = list(calendar_week = cal_weeks))
summary(vwc.emm.week)
  
#Emmeans treatment differences within each year-week
vwc.emm.treatweek <- emmeans(vwc.lmer.all.years.treatment, ~ treatment | calendar_week, at = list(calendar_week = cal_weeks))
summary(vwc.emm.treatweek)

       

# Pairwise comparison of treatments within each year
# tukey gives adjusted p-values 
pairs(vwc.emm, adjust = "fdr")

#contrast between treatments in each week
#contrast(vwc.emm.week, method = "pairwise", adjust = "fdr")

#effect size- emmeans effect size function:
vwc.effect.size.em <- eff_size(vwc.emm,
         sigma = sigma(vwc.lmer.all.years.treatment),
         edf = df.residual(vwc.lmer.all.years.treatment))
vwc.effect.size.em


#Cohens D - effect size:

#install.packages("effsize")  # only need to run this once
#library(effsize)  
vwc.all.treatment.cohend <- cohen.d(
  vwc.all$VWC,
  vwc.all$treatment,
  pooled = TRUE # pooled standard deviation
)

vwc.all.treatment.cohend

#2025 vwc cohens D
vwc.2025.cohend <- cohen.d(
  vwc.2025$VWC,
  vwc.2025$treatment,
  pooled = TRUE # pooled standard deviation
)

vwc.2025.cohend

#2024 vwc cohens D
vwc.2024.cohend <- cohen.d(
  vwc.2024$VWC,
  vwc.2024$treatment,
  pooled = TRUE # pooled standard deviation
)

vwc.2024.cohend


#Cross year Cohens D
vwc.all.year.cohend <- cohen.d(
  vwc.all$VWC,
  vwc.all$year,
  pooled = TRUE # pooled standard deviation
)

vwc.all.year.cohend

```


##vwc plots
```{r}
vwc.2024 %>%
  group_by(treatment) %>%
  summarize(mean_VWC = mean(VWC, na.rm = TRUE),
            se = sd(VWC, na.rm = TRUE)/sqrt(n())) %>%
  ggplot(aes(x = treatment, y = mean_VWC)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean_VWC - se, ymax = mean_VWC + se), width = 0.2)

# Year Comparison 
vwc.all$treatment <- factor(vwc.all$treatment, levels = c("control", "rainout"))

vwc.all %>%
  group_by(year, treatment) %>%
  summarize(mean_VWC = mean(VWC, na.rm = TRUE),
            se = sd(VWC, na.rm = TRUE)/sqrt(n()),
            .groups = "drop") %>%
  ggplot(aes(x = treatment, y = mean_VWC, fill = treatment)) +
  scale_fill_manual(values = treatment_colors) +
  #scale_color_manual(values = treatment_colors) +
  scale_y_continuous(limits = c(0, NA)) +
  geom_col(color = "white") +
  geom_errorbar(aes(ymin = mean_VWC - se, ymax = mean_VWC + se), width = 0.2) +
  facet_wrap(~ year) +
  labs(x = "Treatment", y = "Soil volumetric water content (VWC)",
       title = "") +
  theme_classic() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none")

#raw values of VCW in 2024 and 2025:
vwc.all %>%
  ggplot(aes(x = treatment, y = VWC, color = treatment)) +
  geom_boxplot(fill = NA, linewidth = 0.8) +
  scale_color_manual(values = treatment_colors) +
  scale_y_continuous(limits = c(0, NA)) +
  facet_wrap(~ year) +
  labs(
    x = "Treatment",
    y = "Mean soil volumetric water content (VWC)",
    title = ""
  ) +
  theme_classic() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "gray80", linewidth = 0.3),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )


#By month VWC
ggplot(vwc.all %>% mutate(year = factor(year, levels = c(2024, 2025))), aes(x = date, y = VWC, color = treatment, group = treatment)) +
  geom_point(alpha = 0.6) +                            # raw measurements
  scale_color_manual(values = treatment_colors) +
  geom_line(stat = "summary", fun = mean, linewidth = 1) +  # mean line per treatment
  facet_wrap(~ year, ncol = 1, scales = "free_x") +   # top: 2024, bottom: 2025
  labs(x = "Date", y = "Soil volumetric water content (VWC)",
       color = "Treatment") +
  #scale_x_date(date_breaks = "1 month", date_labels = "%b")+
  theme_classic() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "gray80", size = 0.3),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```


#Flower Counts

## Data Cleaning
```{r}

focalspecies.2024 <- c("delphinium", "erigeron_flagellaris", "umbellatum", "eremogone_congesta", "potentilla_hippiana", "galium_boreale", "heterotheca_villosa", "heliomeris_multiflora")
focalspecies.2025 <- c("clay","delphinium", "umb", "flag", "pot", "congesta", "het", "helio", "galium" )

flowercounts.2024 <- read_sheet("1YR7lBNf5s9Drw9Ix61J8wUtKQO-XuZmXrRMhgktsc5k", sheet="2024") %>%
                      filter(plant!="na") %>%  #2024-07-02, 4f - no flowers open wrote na
                      mutate(transectplot=paste0(transect,plot)) %>%
                      left_join(rainout.treatments) %>% 
                      mutate(week_num = factor(week(date)-26)) %>% 
                      group_by(date, week_num, transect, plot,transectplot, plant, treatment) %>% 
                      filter(! (week_num== "1"& scribe =="sb")) %>% #removes sb's duplicate flower counts in week 1
                      summarize(flower_number=mean(flower_number)) %>% 
                      filter(plant%in% focalspecies.2024) %>% 
                      #rename plant names
                      mutate(plant = dplyr::recode(plant,
                                           "erigeron_flagellaris" ="flag" ,
                                           "eremogone_congesta" = "congesta",
                                           "potentilla_hippiana" = "potentilla",
                                           "galium_boreale" = "galium",
                                           "heterotheca_villosa" = "heterotheca",
                                           "heliomeris_multiflora" = "heliomeris")) %>% 
                      mutate(year = 2024) %>%   
                      ungroup()


flowercounts.2025 <- read_sheet("1ObUirLHnFpHzpepx_tiXLPFxXfSVSJ6s9IAa4_Kd25o", na="") %>% 
                      mutate(flower_number = as.numeric(flower_number)) %>%
                      left_join(rainout.treatments) %>% 
                      group_by(date, week_num_flowcounts, transect, plot,transectplot, plant, treatment) %>% 
                      #summarize(flower_number=mean(flower_number)) %>% 
                      #remove species with little or no data:
                      filter(plant%in% focalspecies.2025) %>% 
                      filter(!plant%in% "clay") %>% #remove claytonia becuase not measured in 2024
                       
                      mutate(plant = dplyr::recode(plant,
                                            "clay" = "claytonia",
                                            "umb" = "umbellatum",
                                            "pot" = "potentilla",
                                            "het" = "heterotheca",
                                            "helio" = "heliomeris")) %>% 
                        mutate(year = 2025) %>% 
                        ungroup()

setdiff( flowercounts.2025$plant, flowercounts.2024$plant)
library(lubridate)

all.flowercounts <- bind_rows(flowercounts.2024, flowercounts.2025) %>% 
                            select(-week_num, -week_num_flowcounts) %>% 
                            mutate(calendar_week = week(date))   #adds calendar week based on the date column

  

unique(all.flowercounts$plant)

```

##Flower Models - treatment
```{r}

#glmmTMB for flower number
#Zero-inflated negative binomial mixed model with plant-specific dispersion:

flowercounts.forglmm <- all.flowercounts %>% 
                            filter(!is.na(flower_number)) %>% #one galium has NA
                            mutate(flower_number = as.integer(flower_number)) %>% 
                            filter(!plant%in%"delphinium") %>% 
                            filter(!plant%in%"galium")

flower.glmm.full <- glmmTMB(flower_number ~ plant * treatment + (1|transect), 
                #ziformula = ~plant*treatment,
                dispformula = ~plant, #Variance differs among plant species, not much among years
                family = nbinom2, 
                data = flowercounts.forglmm)

AIC(flower.glmm.full)

summary(flower.glmm.full)

table(flowercounts.forglmm$plant, flowercounts.forglmm$year)



flowercounts.glmm.dropinteraction <- glmmTMB(flower_number ~ year + plant + (1|transect), 
                            #ziformula = ~ year + plant + flower_number, 
                            dispformula = ~plant, 
                            family = nbinom2, 
                            data = flowercounts.forglmm)
AIC(flowercounts.glmm.dropinteraction)
summary(flowercounts.glmm.dropinteraction)


anova(flower.glmm.full, flowercounts.glmm.dropinteraction) #will do a likelihood ratio test of the effect of the interaction.


```


##Flower Model - year
Does flower number differ by year (per species)
```{r}

avg.flowers.year <- all.flowercounts %>% group_by(plant, year) %>%
                            summarize(mean_flowers = mean(flower_number, na.rm = TRUE),
                                sd_flowers = sd(flower_number, na.rm = TRUE),
                                n = n())

ggplot(all.flowercounts, aes(x = factor(year), y = flower_number, fill = factor(year))) +
  geom_boxplot() +
  facet_wrap(~plant, scales = "free_y") +
  labs(x = "Year", y = "Flower number", fill = "Year")

all.flower.year.lmer <- lmer(flower_number ~ plant * factor(year) +treatment + (1| calendar_week), data = all.flowercounts)
summary(all.flower.year.lmer)

anova(all.flower.year.lmer)

all.flower.year.model <- all.flowercounts %>% filter(!plant%in% "delphinium") %>%  
  group_by(plant) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lmer(flower_number ~ factor(year) + (1|transect/transectplot), data = .x)),
    anova_table = map(model, ~ tidy(anova(.x)))  # convert ANOVA to tibble
  ) %>%
  select(plant, anova_table) %>%
  unnest(anova_table) %>%
  mutate(sig = if_else(p.value < 0.05, "*", ""))  # add * if p < 0.05
View(all.flower.year.model)

#plot flower number year to year

# all.flowercounts %>%
#   filter(!plant %in% "delphinium") %>%
#   ggplot(aes(x = factor(year),
#              y = flower_number)) +
#   geom_boxplot(outlier.shape = NA) +
#   geom_jitter(width = 0.15, alpha = 0.4) +
#   facet_wrap(~ plant, scales = "free_y") +
#   labs(x = "Year",
#        y = "Flower number") +
#   theme_classic()

flowernumber.sig.species <- data.frame(
  plant = c("congesta", "heterotheca", "heliomeris"),
  label = "*")

all.flowercounts %>%
  filter(!plant %in% "delphinium") %>%
  group_by(plant, year) %>%
  summarise(mean_flowers = mean(flower_number, na.rm = TRUE),
            se = sd(flower_number, na.rm = TRUE)/sqrt(n()),
            .groups = "drop") %>%
  ggplot(aes(x = factor(year),
             y = mean_flowers,
             group = plant)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean_flowers - se,
                    ymax = mean_flowers + se),
                width = 0.1) +
  facet_wrap(~ plant, scales = "free_y") +
  labs(x = "Year",
       y = "Mean flower number ± SE") +
  geom_text(data = flowernumber.sig.species,
            aes(x = 1.5, y = Inf, label = label),
            inherit.aes = FALSE,
            vjust = 1.5,
            size = 6) +
  theme_classic()


#but this model is looking across the entire year, do I need to downsample bc 2024 had much less data???
glmm.all.flower.year.model <- all.flowercounts %>%
  filter(!plant %in% "delphinium") %>%
  group_by(plant) %>%
  nest() %>%
  mutate(
    model = map(data, ~ {
      .x <- .x %>% mutate(flower_number = round(flower_number))
      glmer(flower_number ~ factor(year) + (1 | transectplot),
            family = poisson,
            data = .x)
    }),
    anova_table = map(model, ~ broom::tidy(.x, effects = "fixed"))
  ) %>%
  select(plant, anova_table) %>%
  unnest(anova_table) %>%
  mutate(sig = if_else(p.value < 0.05, "*", ""))
View(glmm.all.flower.year.model)


```

##Is flower timining different (onset of timing) between years.

```{r}

flower.timing <- all.flowercounts %>% 
          #filter(!plant%in% "delphinium") %>% #filter delph bc didn't do flower counts early enough in 2024
          group_by(plant, year) %>%
          summarize(
            first_week = min(calendar_week, na.rm = TRUE),  # flowering onset
            last_week  = max(calendar_week, na.rm = TRUE),  # flowering end
            duration   = last_week - first_week + 1,        # flowering duration in weeks
            .groups = "drop")


#Pivot wider to compare 2024 vs 2025
flower.timing.wide <- flower.timing %>%
                      pivot_wider(
                      names_from = year,
                      values_from = c(first_week, last_week, duration),
                      names_sep = "_") %>% 
                    mutate(
                    onset_diff    = first_week_2025 - first_week_2024,
                    end_diff      = last_week_2025 - last_week_2024,
                    duration_diff = duration_2025 - duration_2024)
View(flower.timing.wide)
#Positive values → later start/end or longer duration in 2025.
#Negative values → earlier start/end or shorter duration in 2025.



#plot
flower.timing.long <- flower.timing %>%
        pivot_longer(
          cols = c(first_week, last_week, duration),
          names_to = "metric",
          values_to = "week_value")

ggplot(all.flowercounts,
       aes(x = calendar_week,
           y = flower_number,
           color = plant,
           group = plant)) +
  #scale_fill_manual(values = palette_gradient)+
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  scale_x_continuous(breaks = seq(min(all.flowercounts$calendar_week),
                                  max(all.flowercounts$calendar_week),
                                  by = 1)) +
   scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  facet_wrap(~ year, ncol = 1) +
  labs(
    x = "Calendar week",
    y = "Mean flower number",
    color = "Plant species"
  ) +
  theme_classic() +
  theme(
    panel.grid.minor = element_blank(),       # remove minor grid lines
    panel.grid.major = element_line(color = "lightgrey", size = 0.3)  # keep major grid lines
  )

```

##flower n
Sample size of flower counts in 2024 vs 2025 for each plant species & treatment

n flower counts is the number of times a plant was counted per year (and per treatment below) 
```{r}
#total number of flowers observed per plant × plot × year
n.flowers.years <- all.flowercounts %>%
  group_by(plant, year) %>%
  summarise(total_flowers = sum(flower_number, na.rm = TRUE), .groups = "drop")

#Total number of flowers counted per plant species × transect plot × year × treatment
n.flowers.treat <- all.flowercounts %>%
  group_by(plant, transectplot, year, treatment) %>%
  summarise(total_flowers = sum(flower_number, na.rm = TRUE), .groups = "drop")

#made wide format 
n.flowers.treat.wide <- n.flowers.treat %>%
  pivot_wider(
    names_from = plant,
    values_from = total_flowers,
    values_fill = 0
  )

#heatmap of flower number
library(stringr)

#need to add summarised total flowers to data
flower.number.heatmap.data <- all.flowercounts %>%
  group_by(plant, transect,transectplot,plot, year, treatment) %>%
  summarise(total_flowers = sum(flower_number, na.rm = TRUE), .groups = "drop")

plot_flower_heatmap <- function(data, species_name) {
  
  df <- data %>% filter(plant == species_name)
  
  ggplot(df, aes(x = plot, y = transect, fill = total_flowers)) +
    geom_tile(color = "white") +
    
    facet_grid(year ~ treatment ) +
    
    scale_y_reverse(breaks = 1:10) +
    
    scale_fill_gradient(
      low = "white",
      high = "darkgreen",
      name = "Total flowers"
    ) +
    
    theme_classic() +
    labs(
      x = "Plot",
      y = "Transect",
      title = species_name
    )
}


plot_flower_heatmap(flower.number.heatmap.data, "congesta")

plot_flower_heatmap(flower.number.heatmap.data, "heterotheca")

plot_flower_heatmap(flower.number.heatmap.data, "heliomeris")

plot_flower_heatmap(flower.number.heatmap.data, "flag")

plot_flower_heatmap(flower.number.heatmap.data, "potentilla")


```


# Morohology

##morph year
Match 2024 & 2025 data sets
merge 
model for each trait comparing year x trait of each plant species
data vis 

```{r}

morph.2024 <- read_sheet("1fJuFDOMitdC6bdTm12XWBTZpX4u15uHdkf0y-Ilelw4") %>% 
                    mutate(plot = tolower(as.character(plot))) %>% 
                    mutate(year = 2024) %>%
                    left_join(rainout.treatments) %>% 
                    filter(!plant%in% c("micro white?!", " micro white", "pakera",
                                        "campanula", "collomia", "sedum", "subalpinium",
                                        "speciosus", "microwhite","micro white", 
                                        "tall microwhite", "boechera stricta")) %>% 
#some plants don't have length & width measurements, so this code uses only length if width=NA, otherwise takes the average.  
                     mutate(
                        average_flower_size = if_else(
                        is.na(width),
                        length,
                        (length + width) / 2)) 

morph.2025 <- read_sheet("1grngEjxQ6-uulNl_5yYdbmuajY2DXIGkWfHPBVV-3sM", na=c("")) %>%  
            #remove rows with no tag number
                filter( tag != "") %>% 
                filter(!plant%in% c("clay","delphinium")) %>%
                mutate(year = 2025) %>%
                left_join(rainout.treatments)  %>% 
                #some plants don't have length & width measurements, so this code uses only length if width=NA, otherwise takes the average.  
                     mutate(
                        average_flower_size = if_else(
                        is.na(width),
                        length,
                        (length + width) / 2)) %>% 
                    drop_na(average_flower_size)

unique(morph.2025$plant)


all.years.morph <- bind_rows(morph.2024,morph.2025) %>% 
                    mutate(calendar_week = week(date)) %>% 
                    select(year, date,calendar_week, observer, transect, transectplot,
                           treatment, plant, tag, depth, length, width, 
                           average_flower_size )  
          
#if you want to add in inflo height list above ^


#plot
# Year only - flower size
ggplot(all.years.morph %>% filter(!plant%in% "galium"), aes(x = factor(year), y = average_flower_size)) +
  #scale_fill_manual(values = year_colors)+
  geom_boxplot() +
  facet_wrap(~plant, scales = "free_y") +
  labs(x = "Year", y = "Avg Flower Size", fill = "Year") +
  geom_text(data = avgflowersize.sig.species,
            aes(x = 1.5, y = Inf, label = label),
            inherit.aes = FALSE,
            vjust = 1.5,
            size = 6) +
  theme_classic()


```

## morph avg size

```{r}
#average flower size 2024 vs 2025
#Q: Use anova or Anova???

avg.flower.size.all.years.model <- all.years.morph %>% 
  #galium has a lot of entries where depth, length & width all == NA
  filter(!plant%in% "galium") %>%
  filter(!if_all(c(depth, length, width), is.na)) %>%  
  #filter(!plant%in% "galium") %>% #galium doesn't have enough data
  group_by(plant) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lmer(average_flower_size ~ factor(year)*treatment + (1|observer) + (1|transectplot), data = .x)),
    #ANVOA table for each species:
    anova_table = map(model, ~ {
      broom::tidy(car::Anova(.x)) %>%
        dplyr::rename(F_value = statistic)}),
    # Estimated marginal means for each species
    emmeans_table = map(model, 
            ~ emmeans(.x, ~ factor(year) * treatment) %>%
            as.data.frame()),
    # Pairwise contrasts of treatment within each year
    contrasts_table = map(model, ~ emmeans(.x, pairwise ~ treatment | factor(year))$contrasts %>%
                            as.data.frame() %>%
                            rename(
                              contrast = contrast,
                              estimate = estimate,
                              SE = SE,
                              df = df,
                              t.ratio = t.ratio,
                              p.value = p.value))
    ) %>%
  select(plant, anova_table, emmeans_table, contrasts_table) #%>%
 # unnest(c(anova_table, emmeans_table))# %>%
 # mutate(adjusted.pvalue= p.adjust(p.value, method="fdr")) %>% 
 # mutate(sig = if_else(p.value < 0.05, "*", ""))  # add * if p < 0.05
#view(avg.flower.size.all.years.model )

  # ANOVA table 
anova.flower.size <- avg.flower.size.all.years.model %>%
  select(plant, anova_table) %>%
  unnest(anova_table) %>% 
  mutate(adjusted.pvalue= p.adjust(p.value, method="fdr")) %>%
  mutate(sig = if_else(p.value < 0.05, "*", ""))  # add * if p < 0.05
View(anova.flower.size)

#  emmeans table 
emmeans.flower.size <- avg.flower.size.all.years.model %>%
  select(plant, emmeans_table) %>%
  unnest(emmeans_table)
View(emmeans.flower.size)

#   Pairwise contrasts of treatment within each year 
contrasts.flower.size <- avg.flower.size.all.years.model %>%
  select(plant, contrasts_table) %>%
  unnest(contrasts_table) %>% 
  mutate(sig = if_else(p.value < 0.05, "*", ""))  # add * if p < 0.05
View(contrasts.flower.size)


#   Percent change

#Percent change between treatments within each year
# flower.size.pct.change.treat.year = percent change from control → drought within each year.
flower.size.pct.change.treat.year <- emmeans.flower.size %>%
  select(plant, year, treatment, emmean) %>%
  pivot_wider(
    names_from = treatment,
    values_from = emmean
  ) %>%
  mutate(
    pct_change_drought_vs_control = 100 * (`rainout` - `control`) / `control`
  ) %>%
  select(plant, year, pct_change_drought_vs_control)

View(flower.size.pct.change.treat.year)

#Percent change between years within each treatment
# flower.size.year.pct.change = = percent change from 2024 → 2025 within the same treatment.
flower.size.year.pct.change <- emmeans.flower.size %>%
  select(plant, year, treatment, emmean) %>%
  pivot_wider(names_from = year, values_from = emmean) %>%
  mutate(
    pct_change_year = 100 * (`2025` - `2024`) / `2024`
  ) %>%
  select(plant, treatment, pct_change_year)
View(flower.size.year.pct.change)


```

## Plot - Flower Size

```{r}
#average flower size plot:

avgflowersize.sig.species.year.diffs <- data.frame(
  plant = c("congesta", "heterotheca", "heliomeris", "potentilla","umbellatum"),
  label = "*")

flowersize.sig.treat.diffs.2025 <- data.frame(
  plant = c("potentilla", "flagellaris", "heterotheca", "congesta", "heliomeris"),
  year = 2025,
  x = 2,  # x-position for 2025 (factor(year))
  y = c(23, 25, 32, 16, 36), # adjust y-position above boxplots per species
  #order ^: pot, flag, het, congesta, helio
  label = "*"  # significance asterisk
)

#boxplot for average flower size:

pd <- position_dodge(width = 1) #position of boxplots and jitter points

all.years.morph %>% filter(!plant%in%"galium") %>% 
  ggplot(aes(x = factor(year), y = average_flower_size, color = treatment)) +
  geom_boxplot(fill = NA, linewidth = 0.8,
               position = pd) +
 # geom_jitter(aes(color = treatment),
  #         alpha = 0.4, size = 0.9, position = pd) +
  #scale_color_manual(values = treatment_colors) +
  #scale_y_continuous(limits = c(0, NA)) +
 # facet_wrap(~ plant) +
  facet_wrap(~plant, scales = "free_y") +
  labs(
    x = "Year",
    y = "Average flower size",
    color = "Treatment"
  ) +
  geom_text(
    data = avgflowersize.sig.species.year.diffs,
    aes(x = 1.5, y = Inf, label = label),
    inherit.aes = FALSE,
    vjust = 1.5,
    size = 6
  ) +
  geom_text(
    data = flowersize.sig.treat.diffs.2025,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    size = 6) +
  theme_classic() +
    theme(
    panel.grid.major.x = element_blank(),        # remove vertical lines
    panel.grid.major.y = element_line(color = "grey70", linewidth = 0.3), # horizontal lines
    panel.grid.minor = element_blank())


#Full plot
# ggplot(
#   all.years.morph %>% filter(!plant %in% "galium"),
#   aes(x = factor(year), y = average_flower_size, fill = treatment)
# ) +
#   geom_boxplot(position = position_dodge(width = 0.8)) +
#   facet_wrap(~plant, scales = "free_y") +
#   labs(
#     x = "Year",
#     y = "Average Flower Size",
#     fill = "Treatment"
#   ) +
#   geom_text(
#     data = avgflowersize.sig.species.year.diffs,
#     aes(x = 1.5, y = Inf, label = label),
#     inherit.aes = FALSE,
#     vjust = 1.5,
#     size = 6
#   ) +
#   geom_text(
#     data = flowersize.sig.treat.diffs.2025,
#     aes(x = x, y = y, label = label),
#     inherit.aes = FALSE,
#     size = 6) +
#   theme_classic()

```


##morph depth

```{r}

#Depth: 

depth.all.years.model <- all.years.morph %>% 
  filter(!plant%in% "galium") %>% #galium doesn't have enough data
  group_by(plant) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lmer(depth ~ factor(year) * treatment + (1|observer) +(1|transect/transectplot), data = .x)),
    #ANVOA table for each species:
    anova_table = map(model, ~ {
      broom::tidy(car::Anova(.x)) %>%
        dplyr::rename(F_value = statistic)}),
    # Estimated marginal means for each species
    emmeans_table = map(model, 
            ~ emmeans(.x, ~ factor(year) * treatment) %>%
            as.data.frame()),
    # Pairwise contrasts of treatment within each year
    contrasts_table = map(model, ~ emmeans(.x, pairwise ~ treatment | factor(year))$contrasts %>%
                            as.data.frame() %>%
                            rename(
                              contrast = contrast,
                              estimate = estimate,
                              SE = SE,
                              df = df,
                              t.ratio = t.ratio,
                              p.value = p.value))
    ) %>%
  select(plant, anova_table, emmeans_table, contrasts_table)


# ANOVA table 
anova.flower.depth <- depth.all.years.model %>%
  select(plant, anova_table) %>%
  unnest(anova_table) %>% 
  mutate(adjusted.pvalue= p.adjust(p.value, method="fdr")) %>%
  mutate(sig = if_else(p.value < 0.05, "*", ""))  # add * if p < 0.05
View(anova.flower.depth)

#  emmeans table 
emmeans.flower.depth <- depth.all.years.model %>%
  select(plant, emmeans_table) %>%
  unnest(emmeans_table)
View(emmeans.flower.depth)

#   Pairwise contrasts of treatment within each year 
contrasts.flower.depth <- depth.all.years.model %>%
  select(plant, contrasts_table) %>%
  unnest(contrasts_table) %>% 
  mutate(sig = if_else(p.value < 0.05, "*", ""))  # add * if p < 0.05
View(contrasts.flower.depth)




#flower depth plot:
flowerdepth.sig.species <- data.frame(
  plant = c("congesta", "heterotheca", "flagellaris", "potentilla","umbellatum"),
  label = "*")

#Year only
ggplot(all.years.morph %>%  filter(!plant%in% "galium"), aes(x = factor(year), y = depth)) +
  geom_boxplot() +
  facet_wrap(~plant, scales = "free_y") +
  labs(x = "Year", y = "Flower depth", fill = "Year") +
  geom_text(data = flowerdepth.sig.species,
            aes(x = 1.5, y = Inf, label = label),
            inherit.aes = FALSE,
            vjust = 1.5,
            size = 6) +
  theme_classic()

#Year & Treatment

flower.depth.year.sig.label <- data.frame(
  plant = c("congesta", "heterotheca", "potentilla","umbellatum", "flagellaris"),
  label = "*")

flowerdepth.treat.sig.label <- data.frame(
  plant = c("flagellaris", "heterotheca"),
  year = 2025,
  x = 2,  # x-position for 2025 (factor(year))
  y = c(7, 11), # adjust y-position above boxplots per species
  #order ^: pot, flag, het, congesta, helio
  label = "*"  # significance asterisk
)
pd <- position_dodge(width = 1) #position of boxplots and jitter points

all.years.morph %>% filter(!plant%in%"galium") %>% 
  ggplot(aes(x = factor(year), y = depth, color = treatment)) +
  geom_boxplot(fill = NA, linewidth = 0.8,
               position = pd) +
 # geom_jitter(aes(color = treatment),
  #         alpha = 0.4, size = 0.9, position = pd) +
  #scale_color_manual(values = treatment_colors) +
  #scale_y_continuous(limits = c(0, NA)) +
  #facet_wrap(~ plant) +
  facet_wrap(~plant, scales = "free_y") +
  labs(
    x = "Year",
    y = "Flower Depth",
    color = "Treatment"
  ) +
  geom_text(
    data = flower.depth.year.sig.label,
    aes(x = 1.5, y = Inf, label = label),
    inherit.aes = FALSE,
    vjust = 1.5,
    size = 6
  ) +
  geom_text(
    data = flowerdepth.treat.sig.label,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    size = 6) +
    theme_classic() +
    theme(
    panel.grid.major.x = element_blank(),        # remove vertical lines
    panel.grid.major.y = element_line(color = "grey70", linewidth = 0.3), # horizontal lines
    panel.grid.minor = element_blank())

```


##morph treatment

```{r}
all.morph.traits <- c( "average_flower_size")

#year x treatment morph model for average flower size

morph.treatment.year.models <- all.years.morph %>% 
            filter(!plant%in%"galium") %>% 
            pivot_longer(all_of(all.morph.traits), names_to = "trait") %>% 
            group_by(trait,plant) %>% nest() %>% 
            mutate(morph.trait.lmer=map(data, ~lmer(value~treatment*year + (1|transect), data=.x)), 
            morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%          
            unnest(morph.Anova)%>%
            dplyr::select(-data) %>% #-morph.trait.lmer
            mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
            filter(term!="(Intercept)") %>% 
            mutate(emmeans=map(morph.trait.lmer, ~tidy(emmeans(.x, specs=c("treatment") )) %>% select(-p.value))) %>% 
            select(-morph.trait.lmer)


#only 2024 treatment morph
morph.treatment.models.2024 <- morph.2024 %>% pivot_longer(all_of(all.morph.traits), names_to = "trait") %>% group_by(trait,plant) %>% nest() %>% 
      #filter(!plant%in% c("micro white?!", " micro white", "pakera")) %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~treatment + (1|transect), data=.x)), 
      morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%          
      unnest(morph.Anova)%>%
      dplyr::select(-data) %>% #-morph.trait.lmer
      mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
      filter(term!="(Intercept)") %>% 
       mutate(emmeans=map(morph.trait.lmer, ~tidy(emmeans(.x, specs=c("treatment") )) %>% select(-p.value)))

#kable(morph.models.everything, digits=3) #spits out model into table

#emmeans table 

morph.treatment.emmeans.2024 <- morph.treatment.models.2024 %>% dplyr::select(-morph.trait.lmer, -term, -statistic, -df ) %>% 
      #mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
      unnest(emmeans)  
    
#kable(morph.treatment.emmeans.2024, digits=3) #spits out model into table




```

##morph n
```{r}
#total number of flowers observed per plant × plot × year
n.morph.years <- all.years.morph %>%
  group_by(plant, year) %>%
  summarise(total_flowers_measured = round(sum(average_flower_size, na.rm = TRUE)), .groups = "drop")



```


# Visitation Rate
TODO: how to deal with different number of weeks observed? 2024 has like 6 weeks, 2025 has 12.... downsample? 
Total number of visits different?
Average visits different? 
Visits to plant species different?
Visits by poll groups different? 

After standardizing for flower number, do we see differences in visits by insects? 

```{r}

library(hms)

vrate.cleaned.24 <- read_sheet("1N3HmtcG3Rv4smOS3-SWdpyHAuH3oLvK2KgzWG4JQmNY", na=c("na","","NA"))  %>% 
                mutate(week_num=factor(week_num)) %>% 
                select( -end, -notes, -noyce_data_entry) %>% 
                 mutate(insect_class= fct_collapse(insect_class, 
                       beetle = c("beetle", "beattle"), 
                       lepidoptera= c("butterfly", 
                                "checkerspot butterfly", "moth"), 
                       sol_bee= c( "bee", "carpenter bee"), 
                       fly=c("fly", "hover fly", "hoverfly",
                             "syrphid fly", "beefly", "bee fly"))) %>% 
                      #removed beetles and wasps from dataset
                      filter(!insect_class%in%c("beetle", "wasp")) %>%   
                      mutate(size= fct_collapse(size, 
                            big= c("l", "m", "m/l"), 
                            small = c("t","s", "t/s","s/t", "s/m"))) %>% 
                      #changes NAs in data, assigns NA for sol bee to small size class
                      mutate(size = if_else(is.na(size) & 
                                      insect_class== "sol_bee", "small", size)) %>% 
                      mutate(size = if_else(is.na(size) & 
                                      insect_class== "bombus", "big", size)) %>% 
                      mutate(size = if_else(is.na(size) & 
                                      insect_class== "lepidoptera", "big", size)) %>% 
                      #NOTE: many "fly" insect_class are NA but I don't know the size.... so put placeholder "unknown"
                      mutate(size = if_else(is.na(size) &
                                insect_class == "fly", "unkown", size)) %>% 
                      #assigns NAs (replaces the NaNs) to rows with no size data
                      mutate(insect_size=na_if(paste(size, insect_class), "NA NA")) %>% 
                      #Keep zeros bc you need then for observation.windows but add place holder for plant and insect_class:
                       #mutate(insect_class = if_else(
                        #        is.na(insect_class),
                        #        "no_insects",insect_class)) %>% #if insect_class is NA -> replace with no_insects
                      #mutate(plant = if_else(is.na(plant), "no_visits", plant)) %>% #if plant=NA replace with no_visits
                      #removing plants with little data
                      filter(!plant%in% c("vicia_americana","senecio_integerrimus",
                               "collomia_linearis","erigeron_speciosus",
                               "boechera_stricta","ent","micro white",
                               "campanula_rotundifolia", "eriogonum_subalpinum")) %>% 
                      #rename plant speieces to match flower counts
                      mutate(plant = dplyr::recode(plant,
                                            "eriogonum_umbellatum" = "umbellatum",
                                            "erigeron_flagellaris" = "flag",
                                            "potentilla_hippiana" = "potentilla",
                                            "heterotheca_villosa" = "heterotheca",
                                            "heliomeris_multiflora" = "heliomeris",
                                            "arenaria_congesta" = "congesta",
                                            "galium_boreale" = "galium")) %>% 
                      #remove columns to tidy data set
                      select(-insect_family, -pollinator_group) %>% 
                      #Creates new column flowers.per.plant.visited  
                      #divides # of visits by # of plants visited in that observation bout
                      mutate(flowers.per.plant.visited = case_when(
                            is.na(plants_visited) ~ NA_real_,
                            plants_visited == 0 ~ NA_real_,
                     #if number of plants visited is > number of visits, assign flowers.per.plant.visited ==1
                            plants_visited > visits ~ 1, 
                            TRUE ~ visits / plants_visited)) %>% 
                      rename(vrate24_week_num = week_num) %>% 
                      #fix start column to be time only
                      mutate(start = as_hms(start)) %>% 
                      #add calendar week column: 
                      mutate(calendar_week = week(date))  #adds calendar week based on the date column

#prep for filling non-observed interactions with zeros in the data
#define all possible insect_classes
vrate24.insect.class <- vrate.cleaned.24 %>% 
  distinct(insect_class) %>% 
  filter(!is.na(insect_class))

# define all plants observed
vrate24.plants <- vrate.cleaned.24 %>%
    distinct(plant) %>% 
    filter(!is.na(plant))

# define observation windows
#one row per date × time × plot
#uses calendar_week instead of vrate24_week_num bc some week numbers are wrong
observation.windows.24 <- vrate.cleaned.24 %>%
    distinct(date,calendar_week, start,  
    transect, plot, transectplot)

vrate.24.complete <-  observation.windows.24 %>%
    #create all combinations of time x plants x insects (16,240 total possible combinations)
    #Note: this does not include insect size! 
    crossing(vrate24.plants, vrate24.insect.class) %>%
    left_join(
     vrate.cleaned.24,
      by = c(
        "date", "calendar_week", "start",
        "transect", "plot", "transectplot",
        "plant", "insect_class")) %>%
    mutate(visits = replace_na(visits, 0)) %>%  #makes all NA visits = 0
    mutate(plants_visited = replace_na(plants_visited, 0)) %>% 
    mutate(year = 2024) %>% 
    left_join(rainout.treatments, by = c("transect","transectplot", "plot")) %>% 
    filter(!is.na(treatment)) %>% 
    #Now must filter based on possible interactions where flower_number>0
    left_join(flowercounts.2024) %>% 
    #filter visits to only include an interaction if we recorded flower number for that plant during that week
    #create new column T/F if flowers of that plant are open on that date
    mutate(flower_present = !is.na(flower_number) & flower_number > 0) %>%
    #removes all rows with FALSE (no flowers open) (goes from 16,000 -> 3,500 rows...)  
    filter(flower_present) %>% 
    #remove umbellatum recorded in August because flowers aren't open
    filter(!(plant == "umbellatum" & date == as.Date("2024-08-08"))) %>% 
    # Visits per hour (for 15-min observation windows)
    mutate(visits.per.hr = visits * 4,
   # Visits per flower per hour: only if flower_number > 0
      visitsperflowerperhr = if_else(
      !is.na(flower_number) & flower_number > 0,
      (visits / flower_number) * 4,
      NA_real_)) 
  
  
  
  #remove insect_class = no_insect because isn't an observation
    #filter(!insect_class%in%"no_insects") #3,000 rows now
    #NOTE: is this the best way to do this? maybe shouldn't include option for no_insects above....



#..................... 2025 vrate ........................................
vrate.cleaned.25 <- read_sheet("1Fogbr6sgAxgfnCr5DWvaeCdJNHonhp_mn--mfSsvhCo") %>% 
                      mutate(week_num_vrate25=factor(week_num_vrate)) %>% 
                      select(-week_num_vrate, - end, -data_cleaning_notes, -notes) %>% 
                      #must include NA to keep observation windows with no interaction data
                      filter(plant%in% c(NA, focalspecies.2025)) %>% 
                      filter(!plant%in% "delphinium") %>% #no interaction data for delph
                      #rename plant species to match flower counts
                      mutate(plant = dplyr::recode(plant,
                                            "umb" = "umbellatum",
                                            "pot" = "potentilla",
                                            "het" = "heterotheca",
                                            "helio" = "heliomeris")) %>% 
                     #combine butterfly and moth observations
                      mutate(insect_class= fct_collapse(insect_class, 
                                  lepidoptera= c("butterfly", "moth"))) %>% 
                      filter(!insect_class%in% "wasp") %>%  #remove wasps
                      
                      mutate(start = as_hms(start)) %>% 
                      #adds calendar week based on the date column 
                      mutate(calendar_week = week(date))  %>% 
                      mutate(flowers.per.plant.visited = case_when(
                            is.na(plants_visited) ~ NA_real_,
                            plants_visited == 0 ~ NA_real_,
                     #if number of plants visited is > number of visits, assign flowers.per.plant.visited ==1
                            plants_visited > visits ~ 1, 
                            TRUE ~ visits / plants_visited)) 



#prep for filling non-observed interactions with zeros in the data
#define all possible insect_classes
vrate25.insect.class <- vrate.cleaned.25 %>% 
  distinct(insect_class) %>% 
  filter(!is.na(insect_class))

# define all plants observed
vrate25.plants <- vrate.cleaned.25 %>%
    distinct(plant) %>% 
    filter(!is.na(plant))

# define observation windows
#one row per date × time × plot
#uses calendar_week instead of vrate24_week_num bc some week numbers are wrong
observation.windows.25 <- vrate.cleaned.25 %>%
    distinct(date,calendar_week, start,  
    transect, plot, transectplot)

vrate.25.complete <-  observation.windows.25 %>%
    #create all combinations of time x plants x insects (16,240 total possible combinations)
    #Note: this does not include insect size! 
    crossing(vrate25.plants, vrate25.insect.class) %>%
    left_join(
     vrate.cleaned.25,
      by = c(
        "date", "calendar_week", "start",
        "transect", "plot", "transectplot",
        "plant", "insect_class")) %>%
    mutate(visits = replace_na(visits, 0)) %>%  #makes all NA visits = 0
    mutate(plants_visited = replace_na(plants_visited, 0)) %>% 
    mutate(year = 2025) %>% 
    left_join(rainout.treatments) %>% 
    #Now must filter based on possible interactions where flower_number>0
    left_join(flowercounts.2025 %>% select(-treatment, -year, -transect,-plot), by = c("date","transectplot","plant"),relationship = "many-to-many") %>% 
    #filter visits to only include an interaction if we recorded flower number for that plant during that week
    #create new column T/F if flowers of that plant are open on that date
    mutate(flower_present = !is.na(flower_number) & flower_number > 0) %>%
    #removes all rows with FALSE (no flowers open) (goes from 16,000 -> 3,500 rows...)  
    filter(flower_present) %>% 
    #remove umbellatum recorded in August because flowers aren't open
    #filter(!(plant == "umbellatum" & date == as.Date("2024-08-08"))) %>% 
    # Visits per hour (for 15-min observation windows)
    mutate(visits.per.hr = visits * 4,
   # Visits per flower per hour: only if flower_number > 0
      visitsperflowerperhr = if_else(
      !is.na(flower_number) & flower_number > 0,
      (visits / flower_number) * 4,
      NA_real_))
  
  #combine vrate 2024 and 2025
  #don't I need NAs for visits?? because you want every combination of insect and flower possible (based on open flowers). So there should be a lot of plant x insect with no visits..... both 2024 and 2025 data messed up need plant x insect combination with 0 visits

all.years.vrate <- bind_rows(vrate.24.complete, vrate.25.complete) %>% 
                    mutate(
                      year = factor(year),
                      plant = factor(plant),
                      transectplot = factor(transectplot)) %>% 
                    filter(!plant%in%"galium") #galium does not have enough visitation data for models


all.years.vrate %>% filter(!visits.per.hr== 0) %>% #remove zeros!
  ggplot(aes(x = year,
             y = visits.per.hr,
             color = plant)) +
  geom_boxplot() +
  facet_wrap(~ plant) +
  theme_bw()



ggplot(all.years.vrate,
       aes(x = treatment,
           y = visitsperflowerperhr,
           color = treatment)) +
  geom_point(aes(shape = factor(year)), position = position_dodge(width = 0.5), size = 2) +
  facet_wrap(~ plant) +
  theme_bw() +
  labs(
    x = "Treatment",
    y = "Visits per flower per hr",
    color = "Treatment",
    shape = "Year"  )



```

##Models - vrate

The conditional model asks: given that a plant is being visited, how many visits does it receive? The zero-inflation model asks: what predicts whether a plant receives any visits at all? 
```{r}
#library(glmmTMB)

#glmmTMB with flower number in zi formula
#Zero-inflated negative binomial mixed model with plant-specific dispersion:
fit.year.treat.glmm <- glmmTMB(visits ~ plant*year*treatment + flower_number + (1 | transectplot), 
                ziformula = ~plant*year*treatment + flower_number,
            #need flower number in ziformula bc flower abundance is structurally important for zero probability
                family = nbinom2(), 
                data = all.years.vrate)

AIC(fit.year.treat.glmm)

summary(fit.year.treat.glmm)

car::Anova(fit.year.treat.glmm, type = "III")


fit.year.treat.glmm.dropinteraction <- glmmTMB(visits ~ year + plant + treatment + flower_number + (1|transectplot), 
                            ziformula = ~ year + plant + treatment + flower_number, 
                            family = nbinom2, 
                            data = all.years.vrate)

anova(fit.year.treat.glmm, fit.year.treat.glmm.dropinteraction) #will do a likelihood ratio test of the effect of the interaction.


#emmeans
visits.treat.diff.emms.glmm <-emmeans(fit.year.treat.glmm, pairwise ~ treatment | plant, type="response")
visits.treat.diff.emms.glmm

visits.year.diff.emms.glmm <- emmeans(fit.year.treat.glmm, pairwise ~ year | plant, type="response")
visits.year.diff.emms.glmm


visits.year.pct.decline <- as.data.frame(visits.year.diff.emms.glmm$emmeans) %>% 
   dplyr::select(plant, year, response) %>%
  tidyr::pivot_wider(names_from = year, values_from = response) %>%
  mutate(percent_decline = ((`2024` - `2025`) / `2024`) * 100)
visits.year.pct.decline

```

## visits by bombus

```{r}


bombus.visits <- all.years.vrate %>%
  dplyr::filter(insect_class == "bombus")

bombus.counts <- bombus.visits %>%
  dplyr::count(plant, year, treatment)

# Note: Bombus disappeared from multiple plants in 2025
# so have to take out year*plant interaction in the zi model
fit.bombus.year.treat.glmm <- glmmTMB(
  visits ~ plant+year+treatment + 
     plant:treatment + year:treatment +
    flower_number + (1 | transectplot),
  ziformula = ~ flower_number, 
  #Bombus visits contain extreme zero inflation (80.7% zeros) so must reduce the zi formula
  family = nbinom2(),
  data = bombus.visits
)

AIC(fit.bombus.year.treat.glmm)

summary(fit.bombus.year.treat.glmm)

car::Anova(fit.bombus.year.treat.glmm, type = "III")

#bombus emmeans
bombus.treat.year.emms <- emmeans(fit.bombus.year.treat.glmm,
                                   ~ treatment | plant)
bombus.treat.year.emms

#contrasts
bombus.treat.diff.emms <- emmeans(fit.bombus.year.treat.glmm,
                                  pairwise ~ treatment | plant,
                                  type = "response")

bombus.year.diff.emms <- emmeans(fit.bombus.year.treat.glmm,
                                 pairwise ~ year ,
                                 type = "response")


ggplot(bombus.visits, aes(x = factor(year), y = visits, 
                           color = treatment, shape = treatment)) +
  geom_jitter(size = 2, alpha = 0.6, 
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.5)) +
  
  facet_wrap(~ plant, scales = "free_y") +
  scale_shape_manual(values = c(16, 17, 15)) +
  labs(
    x = "Year",
    y = "Visits",
    color = "Treatment",
    shape = "Treatment",
    title = "Bombus Visit Rates by Treatment and Year"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "grey90"),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "right",
    panel.spacing = unit(0.8, "lines"))

```

## visits by fly
```{r}
fly.visits <- all.years.vrate %>%
  dplyr::filter(insect_class == "fly")

fit.fly.year.treat.glmm <- glmmTMB(
  visits ~ plant*year*treatment + flower_number + (1 | transectplot),
  ziformula = ~ plant*year*treatment + flower_number,
  family = nbinom2(),
  data = fly.visits
)
AIC(fit.fly.year.treat.glmm)
summary(fit.fly.year.treat.glmm)
car::Anova(fit.fly.year.treat.glmm, type = "III")

#Likelihood test:
fit.fly.dropinteraction <- glmmTMB(
  visits ~ plant + year + treatment + flower_number + (1 | transectplot),
  ziformula = ~ plant + year + treatment + flower_number,
  family = nbinom2(),
  data = fly.visits
)

anova(fit.fly.year.treat.glmm, fit.fly.dropinteraction)

#fly emmeans
fly.treat.diff.emms <- emmeans(fit.fly.year.treat.glmm,
                               pairwise ~ treatment | plant,
                               type = "response")
fly.treat.diff.emms

fly.year.diff.emms <- emmeans(fit.fly.year.treat.glmm,
                              pairwise ~ year | plant,
                              type = "response")
fly.year.diff.emms


fly.year.treat.emms <- emmeans(fit.fly.year.treat.glmm,
                               ~ year*treatment | plant)

```

## visits by sol bee

```{r}

solbee.visits <- all.years.vrate %>%
  dplyr::filter(insect_class == "sol_bee")

fit.solbee.year.treat.glmm <- glmmTMB(
  visits ~ plant*year*treatment + flower_number + (1 | transectplot),
  ziformula = ~ plant*year*treatment + flower_number,
  family = nbinom2(),
  data = solbee.visits
)
AIC(fit.solbee.year.treat.glmm)
summary(fit.solbee.year.treat.glmm)
car::Anova(fit.solbee.year.treat.glmm, type = "III")

#liklihood test

fit.solbee.dropinteraction <- glmmTMB(
  visits ~ plant + year + treatment + flower_number + (1 | transectplot),
  ziformula = ~ plant + year + treatment + flower_number,
  family = nbinom2(),
  data = solbee.visits
)

anova(fit.solbee.year.treat.glmm, fit.solbee.dropinteraction)


#emmeans
solbee.treat.diff.emms <- emmeans(fit.solbee.year.treat.glmm,
                                  pairwise ~ treatment | plant,
                                  type = "response")
solbee.treat.diff.emms

solbee.year.diff.emms <- emmeans(fit.solbee.year.treat.glmm,
                                 pairwise ~ year | plant,
                                 type = "response")

solbee.year.diff.emms

solbee.year.treat.emms  <- emmeans(fit.solbee.year.treat.glmm, ~ plant * year * treatment, type = "response")
pairs(solbee.year.treat.emms, by = c("year", "treatment"), adjust = "bonferroni")


#plot
ggplot(solbee.visits, aes(x = factor(year), y = visits, 
                           color = treatment)) +
  geom_jitter(size = 2, alpha = 0.6, 
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.5)) +
  
  facet_wrap(~ plant, scales = "free_y") +
 
  labs(
    x = "Year",
    y = "Visits",
    color = "Treatment",
    
    title = "Solitary bee Visit Rates by Treatment and Year"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "grey90"),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "right",
    panel.spacing = unit(0.8, "lines"))

```

## visits by lep
Lepidoptera data are too sparse for a count model. Must convert the response to presence / absence. This models probability of Lepidoptera visitation rather than visit counts.

report result as: The probability of Lepidoptera visitation was modeled using a binomial GLMM due to extremely low visitation frequencies.
```{r}
#data for count model:
# lep.visits <- all.years.vrate %>%
#   dplyr::filter(insect_class == "lepidoptera")

#data for p/a model: 
lep.visits <- lep.visits %>%
  mutate(presence = ifelse(visits > 0, 1, 0))

#97% of visits = 0 
lep.visits %>%
  summarise(
    total_rows = n(),
    total_visits = sum(visits),
    zeros = sum(visits == 0),
    percent_zero = mean(visits == 0))

lep.visits %>%
  group_by(plant, year, treatment) %>%
  summarise(total_visits = sum(visits), n = n())

fit.lep.presence.glmm <- glmmTMB(
  presence ~ plant + year + treatment +
             plant:treatment + year:treatment +
             flower_number +
             (1 | transectplot),
  family = binomial(),
  data = lep.visits)

AIC(fit.lep.presence.glmm)
summary(fit.lep.presence.glmm)
Anova(fit.lep.presence.glmm)



#emmeans
lep.treat.em <- emmeans(fit.lep.presence.glmm, pairwise ~ treatment | plant, type = "response")

lep.year.em <- emmeans(fit.lep.presence.glmm, pairwise ~ year | plant, type = "response")

#plot 

# Summarize data: proportion of presence per plant × year × treatment
lep.summary <- lep.visits %>%
  group_by(plant, year, treatment) %>%
  summarise(
    n_obs = n(),
    n_present = sum(presence),
    prop_present = n_present / n_obs
  )

# Plot: proportion of visits
ggplot(lep.summary, aes(x = factor(year), y = prop_present, fill = treatment)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  facet_wrap(~ plant, scales = "free_y") +
  labs(
    x = "Year",
    y = "Proportion of observations with Lepidoptera",
    fill = "Treatment",
    title = ""
  ) +
  theme_bw(base_size = 14) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 0, hjust = 0.5)
  )
```


##year only model
```{r}
#lmer
all.years.vrate.lmer <- all.years.vrate %>% 
          group_by(plant) %>%
          nest() %>%
          mutate(model = map(data, ~ lmer(visitsperflowerperhr ~ factor(year) + 
                          (1|calendar_week) +(1|transect/transectplot), data = .x)),
          anova_table = map(model, ~ tidy(anova(.x)))) %>%  # convert ANOVA to tibble
        select(plant, anova_table) %>%
        unnest(anova_table) %>%
        mutate(sig = if_else(p.value < 0.05, "*", ""))  # add * if p < 0.05
  
#glmmTMB with poisson distribution

#library(glmmTMB)

#glmmTMB with flower number in zi formula
#Zero-inflated negative binomial mixed model with plant-specific dispersion:
fit_zipfull <- glmmTMB(visits ~ plant*year + flower_number + (1 | transect), 
                ziformula = ~plant*year + flower_number,
            #need flower number in ziformula bc flower abundance is structurally important for zero probability
                family = nbinom2(), 
                data = all.years.vrate)

AIC(fit_zipfull)

summary(fit_zipfull)


#DON"T NEED TO USE:
#this was used to figure out model structure with initial lowest AIC value
#glmmTMB without flower number in zi formula
#In fit_zip, only plant and year influence the probability of getting zero visits.

# fit_zip <- glmmTMB(visits ~ plant*year + flower_number + (1|transect),
#               ziformula = ~plant*year + flower_number, 
#               family = poisson, 
#               data = all.years.vrate)
# 
# summary(fit_zip)

#find model with lowest AIC:
#AIC(fit_zipfull, fit_zip)

fit_zipdropinteraction <- glmmTMB(visits ~ year + plant + flower_number + (1|transect), 
                            ziformula = ~ year + plant + flower_number, 
                            dispformula = ~plant, 
                            family = nbinom2, 
                            data = all.years.vrate)

anova(fit_zipfull, fit_zipdropinteraction) #will do a likelihood ratio test of the effect of the interaction.


```

##2025 treatment model
```{r}
#2025 visitation rate treatment glmmTMB model:

#glmmTMB with flower number in zi formula
fit_zipfull_treatment_25 <- glmmTMB(visits ~ plant*treatment + flower_number + (1|transect), 
                ziformula = ~plant*treatment + flower_number,
                family = nbinom2, 
                dispformula = ~plant,
                data = all.years.vrate %>% filter(year==2025))
AIC(fit_zipfull_treatment_25)

summary(fit_zipfull_treatment_25)


#DON"T NEED: 
#glmmTMB 
#In fit_zip, only plant and year influence the probability of getting zero visits.
# 
# fit_zip_treatment_25 <- glmmTMB(visits ~ plant*treatment + flower_number + (1|transect),
#               ziformula = ~plant*treatment, 
#               family = poisson, 
#               data = all.years.vrate %>% filter(year==2025))

#find model with lowest AIC:
#AIC(fit_zipfull_treatment_25, fit_zip_treatment_25)

#Model with no interaction
fit_zipdropinteraction_treatment_25 <- glmmTMB(visits ~ treatment + plant + flower_number + (1|transect), 
                            ziformula = ~plant + treatment + flower_number, 
                            family = nbinom2, 
                            dispformula = ~plant,
                            data = all.years.vrate %>% filter(year==2025))
AIC(fit_zipdropinteraction_treatment_25)

anova(fit_zipfull_treatment_25, fit_zipdropinteraction_treatment_25)

```

##2024 treatment model
```{r}
#2024 visitation rate treatment glmmTMB model:

#glmmTMB with flower number in zi formula
fit_zipfull_treatment_24 <- glmmTMB(visits ~ plant*treatment + flower_number + (1|transect), 
                ziformula = ~plant*treatment + flower_number,
                family = nbinom2, 
                dispformula = ~plant, 
                data = all.years.vrate %>% filter(year==2024))
AIC(fit_zipfull_treatment_24)
summary(fit_zipfull_treatment_24)


#DON"T NEED:
#glmmTMB 
#In fit_zip, only plant and year influence the probability of getting zero visits.
# 
# fit_zip_treatment_24 <- glmmTMB(visits ~ plant*treatment + flower_number + (1|transect),
#               ziformula = ~plant*treatment, 
#               family = poisson, 
#               data = all.years.vrate %>% filter(year==2024))

#find model with lowest AIC:
#AIC(fit_zipfull_treatment_24, fit_zip_treatment_24)

fit_zipdropinteraction_treatment_24 <- glmmTMB(visits ~ treatment + plant + flower_number + (1|transect), 
                            ziformula = ~ treatment + plant + flower_number, 
                            family = nbinom2, 
                            dispformula = ~plant, 
                            data = all.years.vrate %>% filter(year==2024))

anova(fit_zipfull_treatment_24, fit_zipdropinteraction_treatment_24)

#Model sample size
nobs(fit_zipfull_treatment_24) #returns the number of observations actually used after missing data were dropped

```
##vrate n
Sample size of number of observed pollinator interactions in 2024 vs 2025 for each plant species

One row = one sampling unit (e.g., one plot × date × plant individual). Therefore n is the number of unique rows included in the analysis. For vrate, one row is one interaction event for plant x insect.
```{r}
#Sample size per plant species per year
#n = number of observations
n.vrate.years <- all.years.vrate %>%
  group_by(plant, year) %>%
  summarise(n = n())

#Sample size per plant species & year by treatment
n.vrate.treat <- all.years.vrate %>%
  group_by(plant, year, treatment) %>%
  summarise(n = n(), .groups = "drop")

```


#plotting vrate
```{r}

library(glmmTMB)
library(ggeffects)
library(emmeans)
library(ggplot2)
library(dplyr)
library(patchwork)

plot_predicted_visitation <- function(model, year_label) {

  preds <- ggpredict(
    model,
    terms = c("flower_number" ,"plant", "treatment"),
    type = "fixed",
    bias_correction = TRUE
  )

  ggplot(preds, aes(x = x, y = predicted, color = group)) + #x=flower number
    geom_point(position = position_dodge(width = 0.3), size = 3) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                  width = 0.2,
                  position = position_dodge(width = 0.3)) +
    labs(
      title = paste("Predicted pollinator visitation —", year_label),
      x = "Flower Number",
      y = "Predicted visitation rate",
      color = "Treatment"
    ) +
    theme_classic()
}

#Predicted visitation panels (2024 and 2025)
predicted_visitation_2024_plot <- plot_predicted_visitation(
  fit_zipfull_treatment_24, "2024"
)

predicted_visitation_2025_plot <- plot_predicted_visitation(
  fit_zipfull_treatment_25, "2025"
)

predicted_visitation_by_flower_abundance_panel <- 
  predicted_visitation_2024_plot + predicted_visitation_2025_plot

```

#glmm by insect

```{r}

vrate.yearxtreat.visits.by.group <- all.years.vrate %>% 
                mutate(plant_insect = paste0(plant, insect_class)) %>% 
                # group by year, plant, treatment, and pollinator group
                group_by(year, plant, plant_insect, treatment, insect_class, transect, transectplot, flower_number) %>%   
                summarize(
                  # total visits for that group
                  visits_by_insect_group = sum(visits, na.rm = TRUE),  
                  n_observations = n(), # number of observations
                  .groups = "drop")


#glmm for visits by insect group (???)
visitsbyinsects_zipfull <- glmmTMB(visits_by_insect_group ~ plant*year + flower_number + (1 | transect), 
                ziformula = ~plant*year + flower_number,
            #need flower number in ziformula bc flower abundance is structurally important for zero probability
                family = nbinom2(), 
                data = vrate.yearxtreat.visits.by.group)

AIC(visitsbyinsects_zipfull)

summary(visitsbyinsects_zipfull)




#glmm for visits per hour (not total visits)
visitswithinsectinmodel_zipfull <- glmmTMB(visits ~ insect_class*year*treatment + flower_number + (1 | transect), 
                ziformula = ~insect_class*year + flower_number,
            #need flower number in ziformula bc flower abundance is structurally important for zero probability
                family = nbinom2(), 
                data = all.years.vrate)

AIC(visitswithinsectinmodel_zipfull)

summary(visitswithinsectinmodel_zipfull)



#glmm for visits per hour (not total visits)
visitsperhr_zipfull <- glmmTMB(visits.per.hr ~ plant*year*treatment + flower_number + (1 | transect), 
                ziformula = ~plant*year + flower_number,
            #need flower number in ziformula bc flower abundance is structurally important for zero probability
                family = nbinom2(), 
                data = all.years.vrate)

AIC(visitsperhr_zipfull)

summary(visitsperhr_zipfull)



```



#Treatment Emmeans for visitation by year (24 v 25)

```{r}
#Extract emmeans from each model
# library(emmeans)
# library(dplyr)
# library(ggplot2)
# 
# emm.2024.visits.treat <- emmeans(
#   fit_zipfull_treatment_24,
#   ~ plant * treatment,
#   type = "response" #response is visits
# ) %>%
#   as.data.frame() %>%
#   mutate(year = "2024")
# 
# emm.2025.visits.treat <- emmeans(
#   fit_zipfull_treatment_25,
#   ~ plant * treatment,
#   type = "response" #response is visits
# ) %>%
#   as.data.frame() %>%
#   mutate(year = "2025")
# 
# emm.combined.visits.treat <- bind_rows(emm.2024.visits.treat, emm.2025.visits.treat )
# 
# 
# # Function to calculate percent change between two years
# year_calculate_percent_change <- function(emm_data, value_col = "response") {
#   # Ensure year is numeric
#   emm_data <- emm_data %>% mutate(year = as.numeric(year))
#   
#   # Separate 2024 and 2025
#   data_2024 <- emm_data %>% filter(year == 2024) %>% select(plant, treatment, !!sym(value_col))
#   data_2025 <- emm_data %>% filter(year == 2025) %>% select(plant, treatment, !!sym(value_col))
#   
#   # Join the two years
#   combined <- data_2024 %>%
#     left_join(data_2025, by = c("plant", "treatment"), suffix = c("_2024", "_2025"))
#   
#   # Calculate percent change
#   combined <- combined %>%
#     mutate(percent_change = 100 * ((!!sym(paste0(value_col, "_2025")) - !!sym(paste0(value_col, "_2024"))) / !!sym(paste0(value_col, "_2024"))))
#   
#   return(combined)
# }
# 
# # Example usage for visitation emmeans
# year_percent_change_visitation <- year_calculate_percent_change(emm.combined.treat, value_col = "response")
# 
# #TODO: doesn't work bc need to do emmeans for flower number...
# # Percent change flower_number emmeans
# year_percent_change_flower <- year_calculate_percent_change(emm.combined.treat, value_col = "flower_number")
# 
# 
# ####################################################
# 
# #plot raw data & emmeans
# 
# library(ggplot2)
# 
# ggplot(all.years.vrate, aes(x = factor(year), y = visits, color = treatment)) +
#   geom_boxplot(outlier.shape = NA, alpha = 0.5, position = position_dodge(width = 0.75)) +
#   geom_jitter(width = 0.1, alpha = 0.3) +  # raw points overlaid
#   # geom_point(
#   #   data = emm.combined.treat,
#   #   aes(x = factor(year), y = response, color = treatment),
#   #   position = position_dodge(width = 0.75),
#   #   size = 3,
#   #   inherit.aes = FALSE
#   # ) +
#   geom_errorbar(
#     data = emm.combined.treat,
#     aes(x = factor(year), ymin = asymp.LCL, ymax = asymp.UCL, color = treatment),
#     width = 0.2,
#     position = position_dodge(width = 0.75),
#     inherit.aes = FALSE
#   ) +
#   facet_wrap(~plant) +
#   theme_classic() +
#   labs(x = "Year", y = "Pollinator visitation", color = "Treatment")
# 
# # Violin plot with raw data and emmeans
# violin_visitation_plot <- ggplot(all.years.vrate, aes(x = factor(year), y = visits, color = treatment)) +
#   
#   # Violin to show distribution per year/treatment
#   geom_violin(alpha = 0.3, position = position_dodge(width = 0.8)) +
#   
#   # Raw data points overlaid with jitter + dodge
#   # geom_jitter(
#   #   position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8),
#   #   alpha = 0.3
#   # ) +
#   
#   # Add emmeans points
#   # geom_point(
#   #   data = emm.combined.treat,
#   #   aes(x = factor(year), y = response, color = treatment),
#   #   size = 3,
#   #   position = position_dodge(width = 0.8),
#   #   inherit.aes = FALSE
#   # ) +
#   
#   # Add emmeans confidence intervals
#   geom_errorbar(
#     data = emm.combined.treat,
#     aes(x = factor(year), ymin = asymp.LCL, ymax = asymp.UCL, color = treatment),
#     width = 0.2,
#     position = position_dodge(width = 0.8),
#     inherit.aes = FALSE
#   ) +
#   
#   # Facet by plant species
#   facet_wrap(~plant, scales ="free_y") +
#   
#   theme_bw() +
#   labs(
#     x = "Year",
#     y = "Pollinator visitation (visits)",
#     color = "Treatment"
#   )
# 
# violin_visitation_plot

```


#Raw data with model predictions overlay (both years)
```{r}

# predictions_24 <- ggpredict(fit_zipfull_treatment_24,
#                            terms = c("flower_number [all]", "plant", "treatment"))
# predictions_24$year <- "2024"
# 
# predictions_25 <- ggpredict(fit_zipfull_treatment_25,
#                            terms = c("flower_number [all]", "plant", "treatment"))
# predictions_25$year <- "2025"
# 
# prediction_data_combined <- bind_rows(predictions_24, predictions_25)
# 
# raw_data_with_model_predictions_plot <- ggplot(
#   all.years.vrate,
#   aes(flower_number, visits, color = plant)
# ) +
#   geom_point(
#     alpha = 0.4,
#     position = position_jitter(width = 0.1, height = 0)
#   ) +
#   geom_line(
#     data = prediction_data_combined,
#     aes(
#       x = x,
#       y = predicted,
#       color = group,
#       linetype = facet
#     ),
#     linewidth = 1,
#     inherit.aes = FALSE
#   ) +
#   geom_ribbon(
#     data = prediction_data_combined,
#     aes(
#       x = x,
#       ymin = conf.low,
#       ymax = conf.high,
#       fill = group
#     ),
#     alpha = 0.15,
#     color = NA,
#     inherit.aes = FALSE
#   ) +
#   facet_wrap(~year) +
#   theme_classic() +
#   labs(
#     x = "Flower number",
#     y = "Observed visitation",
#     color = "Plant species",
#     fill = "Plant species",
#     linetype = "Treatment"
#   )
# 
# raw_data_with_model_predictions_plot

```

#Predicted model - Estimated marginal means (species × treatment)
```{r}

# plot_emmeans_species_treatment <- function(model, year_label) {
#   
#   preds <- ggeffects::ggpredict(
#     model,
#     terms = c("plant", "treatment"),
#     bias_correction = TRUE
#   )
#   
#   p <- ggplot(preds, aes(x = x, y = predicted, color = group)) +
#     geom_point(position = position_dodge(width = 0.3), size = 3) +
#     geom_errorbar(
#       aes(ymin = conf.low, ymax = conf.high),
#       position = position_dodge(width = 0.3),
#       width = 0.2
#     ) +
#     theme_classic() +
#     labs(
#       title = paste("Estimated visitation:", year_label),
#       x = "Plant species",
#       y = "Predicted visitation",
#       color = "Treatment"
#     )
#   
#   return(p)
# }
# 
# emmeans_plot_2024 <- plot_emmeans_species_treatment(fit_zipfull_treatment_24, "2024")
# emmeans_plot_2024
# 
# emmeans_plot_2025 <- plot_emmeans_species_treatment(fit_zipfull_treatment_25, "2025")
# emmeans_plot_2025

```


#difference in observed insect visits (relative amount of visits in each year)

```{r}
#2024 relative amounts of observed insect visits
relative.insects.24 <- vrate.24.complete %>%   count(insect_class) %>%
  mutate(relative_proportion = n / sum(n))

#2025
relative.insects.25 <- vrate.25.complete %>%   count(insect_class) %>%
  mutate(relative_proportion = n / sum(n))

#NOTE: proportion of visits per insect_class is not different between years. 

```

#relative insect visits per plant species
```{r}

#Relative proportion of all insect visits to plant species (NOT STADARDIZED by flower number)
#proportion of visits by an insect per year out of that year's total
relative.global.by.year <- all.years.vrate %>% 
  group_by(year, plant, insect_class) %>% 
  summarise(total_visits = sum(visits), .groups = "drop") %>% 
  group_by(year) %>% 
  mutate(relative_proportion = round(total_visits / sum(total_visits), 3)) %>% 
  ungroup()

ggplot(relative.global.by.year,
       aes(x = factor(year),
           y = relative_proportion,
           fill = insect_class)) +
  geom_col() +
  #scale_fill_manual(values = palette_gradient) +
  facet_wrap(~ plant) +
  labs(
    x = "Year",
    y = "Relative proportion of all visits",
    fill = "Insect class"
  ) +
  theme_bw()

#Standardize relative proportion of visits to plants by flower number
#Standardize within plant first, then global to remove plant abundance effects but still compare insects
relative.by.plant.flower <- all.years.vrate %>% 
  group_by(year, plant, insect_class) %>% 
  summarise(
    total_visits = sum(visits),
    flowers = mean(flower_number),
    .groups = "drop"
  ) %>% 
  mutate(visits_per_flower = total_visits / flowers) %>% 
  group_by(year, plant) %>% 
  mutate(relative_within_plant = visits_per_flower / sum(visits_per_flower)) %>% 
  ungroup()

#DON'T USE! only for visualization purposes - use relative.global.plant.flower.by.year below
ggplot(relative.by.plant.flower,
       aes(x = factor(year),
           y = relative_within_plant,
           fill = insect_class)) +
  geom_col() +
  facet_wrap(~ plant) +
  labs(
    x = "Year",
    y = "Relative visits per flower",
    fill = "Insect class"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


#USE THIS: 
#For a single year, what fraction of all flower-standardized visits went to each plant × insect combination, relative to the total for that year?
#Insect visits to plants standardized by flower number
#global proportion of insect visits to each plant species within a single year that is standardized by flower number
##Sum the visits_per_flower across all plants for each insect class within a year.
##Divide by the total visits_per_flower for that year to get the proportion.

relative.global.plant.flower.by.year <- relative.by.plant.flower %>%
                      group_by(year, plant, insect_class) %>% #all standardized visits per plant × insect are summed
                      summarise(total_visits_per_flower = sum(visits_per_flower),
                                  .groups = "drop") %>%
                      group_by(year) %>% #the denominator is all visits per flower in that year, across all plants and insects
                      mutate(relative_proportion = round(total_visits_per_flower / sum(total_visits_per_flower), 3)) %>%
                      ungroup()


ggplot(relative.global.plant.flower.by.year,
       aes(x = factor(year),
           y = relative_proportion,
           fill = insect_class)) +
  geom_col() +
  facet_wrap(~ plant) +
  labs(
    x = "Year",
    y = "Relative visits per flower",
    fill = "Insect class"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(
    panel.grid.major.x = element_blank(),        # remove vertical lines
    panel.grid.major.y = element_line(color = "gray80", linewidth = 0.3), # horizontal lines
    panel.grid.minor = element_blank())

ggplot(relative.global.plant.flower.by.year,
       aes(x = factor(year),
           y = relative_proportion,
           color = insect_class,
           group = insect_class)) +
  geom_point(size = 2) +
  geom_line() +
  facet_wrap(~ plant) +
  labs(
    x = "Year",
    y = "Relative visits per flower",
    color = "Insect class"
  ) +
  theme_bw()


```


#network
##2024
```{r}

net.2024 <- vrate.24.complete %>% 
            group_by(plant, insect_class) %>% 
            summarise(visits=sum(visits)) %>% 
            drop_na(plant) %>% 
            pivot_wider(names_from=insect_class, values_from = visits, values_fill = 0) %>%  
            arrange(plant) %>% #orders rows (plant names) alphabetically
            column_to_rownames("plant") %>% 
            relocate(sort(tidyselect::peek_vars()))  #orders columns alphabetically

#bipartite network plot:                          
pp.network.24 <- plotweb(sqrt(net.2024), empty=F, text.rot=90, method="normal")

#network statistics
net.24.stats <- networklevel(net.2024) 



```

##2025
```{r}
net.2025 <- vrate.25.complete %>% 
            group_by(plant, insect_class) %>% 
            summarise(visits=sum(visits)) %>% 
            drop_na(plant) %>% 
            pivot_wider(names_from=insect_class, values_from = visits, values_fill = 0) %>%  
            arrange(plant) %>% #orders rows (plant names) alphabetically
            column_to_rownames("plant") %>% 
            relocate(sort(tidyselect::peek_vars()))  #orders columns alphabetically

#bipartite network plot:                          
pp.network.25 <- plotweb(
  sqrt(net.2025),
  empty = FALSE,
  text.rot = 90,
  method = "normal",
)

#network statistics
net.25.stats <- networklevel(net.2025) 
```

#Network rareification
downsample networks to 1000 rows for each treatment in each year

-determine smallest network size
-rarefy within year × treatment networks
-generate 500 replicate networks
-convert to matrices
-calculate network metrics
-analyze distributions of metrics across replicates

```{r}

#Creates networks for each treatment in each year 
#NOTE: creating week level networks doesn't generate enough interaction data to run networklevel() on.

#Define species list
plants <- sort(unique(all.years.vrate$plant))
pollinators <- sort(unique(all.years.vrate$insect_class))

#Rarefaction depth:
rarefaction.depth <- all.years.vrate %>%
  filter(year %in% c(2024, 2025),
         treatment %in% c("control","rainout")) %>%
  count(year, treatment)

min.rows.for.rarefaction <- min(rarefaction.depth$n)
min.rows.for.rarefaction #1349 rows

#target rows need to be lower than min.rows.for.rarefaction
target.rows <- 1000


#Rarefaction funtion (within each network)
#Rarefies within each year × treatment network before aggregation

rarefy.network.function <- function(df, sample_size, rep_id) {

  df %>%
    filter(year %in% c(2024, 2025),
           treatment %in% c("control","rainout")) %>%
    
    group_by(year, treatment) %>%
    
    group_modify(~{

      sampled <- .x %>%
        slice_sample(n = sample_size, replace = FALSE)

      aggregated <- sampled %>%
        group_by(plant, insect_class) %>%
        summarise(visits = sum(visits), .groups = "drop")

      full_grid <- expand_grid(
        plant = plants,
        insect_class = pollinators
      )

      full_grid %>%
        left_join(aggregated,
                  by = c("plant","insect_class")) %>%
        mutate(
          visits = replace_na(visits,0),
          replicate = rep_id
        )
    }) %>%
    
    ungroup()
}

#generate rarefied networks
# with n number of replicate networks
#output structure: year, treatment, plant, insect_class, visits, replicate number

n_reps <- 500

rarefied.networks <- map_dfr(
  1:n_reps,
  ~rarefy.network.function(
      df = all.years.vrate,
      sample_size = target.rows,
      rep_id = .x)
)

#Convert to plant × pollinator network matrices (stats)
pp.matrix.function <- function(df){

  df %>%
    select(plant, insect_class, visits) %>%
    pivot_wider(names_from = insect_class,
                values_from = visits,
                values_fill = 0) %>%
    column_to_rownames("plant") %>%
    as.matrix()
}


#Build matrix list for each network replicate
#Each element = one network matrix
#500 replicates × 4 networks = 2000 matrices

matrices.list <- rarefied.networks %>%
  group_by(year, treatment, replicate) %>%
  group_split() %>%
  map(pp.matrix.function)

#Calculate network metrics (stats) using bipartite package

#library(bipartite)

network.metrics.all <- matrices.list %>%
  map_df(~{

    mat <- .x

    mat <- mat[rowSums(mat) > 0, colSums(mat) > 0]

    if(nrow(mat) < 2 | ncol(mat) < 2 | sum(mat) < 3){
      return(tibble(
        connectance = NA,
        NODF = NA,
        H2 = NA,
        links_per_species = NA
      ))
    }

    metrics <- networklevel(
      mat,
      index = c("connectance","NODF","H2","links per species")
    )

    as_tibble(as.list(metrics))
  })

#Attach metadata back to metrics
network.ids <- rarefied.networks %>%
  distinct(year, treatment, replicate) %>%
  arrange(year, treatment, replicate)

network.metrics.all.final <- bind_cols(network.ids, network.metrics.all) %>% 
                              rename(links_per_species = `links per species`)

#Mean network metrics per treatment:
mean.network.matrics.per.treatment <- network.metrics.all.final %>%
  group_by(year, treatment) %>%
  summarise(across(connectance:links_per_species,
            #alternative if you add more network metrics replace above line with:
            #across(where(is.numeric) & !replicate, \(x) mean(x, na.rm = TRUE))
           \(x) mean(x, na.rm = TRUE)),
            .groups = "drop")

```

##LMM network metrics

statistical comparison of network metrics across treatments and years  
handle the replicate rarefactions to avoid pseudoreplication)

500 rarefied networks per year × treatment, so each metric has a distribution of values

treats replicate rarefactions as repeated measurements of the same underlying network, not independent ecological replicates.--> use mixed-effects model with fix & random effects to control for repeated measures

```{r}

connectance.model <- lmer(
  connectance ~ year * treatment + (1 | replicate), #replicate does not contribute to the model boundary is singlular warning, try t-test for nonparametric analysis
  data = network.metrics.all.final)

#p-value is very small since there are many replicates, doesn't mean much.
Anova(connectance.model, type = "III")

nodf.mod <- lmer(
  NODF ~ year * treatment + (1 | replicate),
  data = network.metrics.all.final)
Anova(nodf.mod, type = "III")

h2.mod <- lmer(
  H2 ~ year * treatment + (1 | replicate),
  data = network.metrics.all.final)
Anova(h2.mod, type = "III")

links.mod <- lmer(
  links_per_species ~ year * treatment + (1 | replicate),
  data = network.metrics.all.final)
Anova(links.mod, type = "III")

#plot 
#change out connectance with different metrics 
#shows the uncertainty introduced by rarefaction
ggplot(network.metrics.all.final,
       aes(treatment, connectance, fill = treatment)) +
  geom_violin() +
  geom_boxplot(width = 0.2) +
  facet_wrap(~year)

#emmeans
library(emmeans)

#treatment differences within each year
#connectance, H2 emmeans contrasts significant for treatment in both years.  
#Only links per species in 2024 (not 2025), nodf emmeans contrasts no significant
emmeans(connectance.model, pairwise ~ treatment | year)
emmeans(nodf.mod, pairwise ~ treatment | year)
emmeans(h2.mod, pairwise ~ treatment | year)
emmeans(links.mod, pairwise ~ treatment | year)

#year differences within each treatment:
#Summary: all metrics shows year difference P<0.001
emmeans(connectance.model, pairwise ~ year | treatment)
emmeans(nodf.mod, pairwise ~ year | treatment)
emmeans(h2.mod, pairwise ~ year | treatment)
emmeans(links.mod, pairwise ~ year | treatment)

```

##glmmTMB for rarefied networks
```{r}

glmm.full.downsample.net <- glmmTMB(
  visits ~ year * treatment * plant + (1|insect_class),
  ziformula = ~1,
  family = nbinom2,
  data = downsampled.network)

summary(glmm.full.downsample.net)

Anova(glmm.full.downsample.net, type = "III")

#post-hoc comparisons using emmeans

#Treatment effect within each year
emmeans(glmm.full.downsample.net, pairwise ~ treatment | year)

#Year effect within each treatment
emmeans(glmm.full.downsample.net, pairwise ~ year | treatment)

```


###2024
```{r}

#down sampling allows you to have same number of interactions in both control and drought so you can compare the two networks more easily
#take net.2024 and net.2025 and take random rows

#number to down sample to
#figure out number to downsample to:
vrate.24.complete %>% count(treatment)
vrate.25.complete %>% count(treatment)

downsample.rows <- 1000 #number of rows in down sampled networks
#down sample number based on number of interactions in 2024 control 

#make mini networks 

downsample.function.2024 <- function(manipulation, downsample.rows){
                          vrate.24.complete %>%  
                          #left_join(rainout.treatments) %>% 
                          filter(treatment== manipulation) %>% 
                          slice_sample(n=downsample.rows) %>% 
                          group_by(plant, insect_class) %>% summarise(visits=sum(visits)) %>% drop_na(plant) %>% 
                          pivot_wider(names_from=insect_class, values_from = visits, values_fill = 0) %>%  
                          arrange(plant) %>% #orders rows (plant names) alphabetically
                          column_to_rownames("plant") %>% 
                          relocate(sort(tidyselect::peek_vars())) %>%  #orders columns alphabetically
                          networklevel(index=c("connectance", "NODF", "H2", "links per species")) %>% 
                          t() %>% data.frame()
                        }

stats.24.downsample <- tibble(permutation=rep(1:99, times=2), treatment=rep(c("control", "rainout"), each=99)) %>% 
              mutate(stats= map(treatment, ~downsample.function.2024(.x, downsample.rows))) %>% unnest(stats)

# Links per species plot (species = insect class and size)
dnet.color.palette <- c(control="cornflowerblue", rainout="orange")
year.color.palette <- c("2024" = "darkgreen", "2025"= "purple")


ggplot(stats.24.downsample, aes(x=links.per.species, fill= treatment))+ geom_histogram(binwidth = 0.05) + scale_fill_manual(values=dnet.color.palette) + facet_wrap(vars(treatment), ncol = 1)

ggplot(stats.24.downsample, aes(x=connectance, fill= treatment))+ geom_histogram(binwidth = 0.05) + scale_fill_manual(values=dnet.color.palette) + facet_wrap(vars(treatment), ncol = 1)



#t test
#don't use pvalue bc we are just choosing how much data to create with number of permutations
anova(lm(links.per.species~treatment, data=stats.24.downsample))



```

###2025
```{r}
downsample.function.2025 <- function(manipulation, downsample.rows){
                          vrate.25.complete %>%  
                          #left_join(rainout.treatments) %>% 
                          filter(treatment== manipulation) %>% 
                          #randomly selects a subset of rows
                          slice_sample(n=downsample.rows) %>% 
                          #aggregate visits 
                          group_by(plant, insect_class) %>%
                          summarise(visits=sum(visits)) %>% 
                          #removes missing plant names
                          drop_na(plant) %>% 
                          #convert to an interaction matrix
                          pivot_wider(names_from=insect_class, values_from = visits, values_fill = 0) %>%  
                          arrange(plant) %>% #orders rows (plant names) alphabetically
                          #column_to_rownames("plant") %>% 
                          #relocate(sort(tidyselect::peek_vars())) %>%  #orders columns alphabetically
                          #networklevel(index=c("connectance", "NODF", "H2", "links per species")) %>% 
                          #t() %>% data.frame()
                          print(head(downsampled_data))  # Debugging line
                        }

    test_output <- downsample.function.2025("control", downsample.rows)
    print(test_output)  # Check the structure and columns of the output

downsample.25 <- tibble(permutation = rep(1:99, times = 2), 
                              treatment = rep(c("control", "rainout"), each = 99)) %>% 
  mutate(stats = map(treatment, ~downsample.function.2025(.x, downsample.rows))) %>% 
  unnest(stats, keep_empty = TRUE) %>%  # Keep empty rows if any
  select(plant, insect_class, visits, permutation, treatment)  # Select relevant columns

#####
#999 permutations for each treatment
stats.25.downsample <- tibble(permutation=rep(1:99, times=2), treatment=rep(c("control", "rainout"), each=99)) %>% 
                        mutate(stats= map(treatment, ~downsample.function.2025(.x, downsample.rows))) %>% 
                        unnest(stats) #Expands the list column so each permutation becomes a normal row with network metrics

# Links per species plot (species = insect class and size - TODO: does it? might need to double check)
ggplot(stats.25.downsample, aes(x=links.per.species, fill= treatment))+ geom_histogram(binwidth = 0.05) + scale_fill_manual(values=dnet.color.palette) + facet_wrap(vars(treatment), ncol = 1)

ggplot(stats.25.downsample, aes(x=connectance, fill= treatment))+ geom_histogram(binwidth = 0.05) + scale_fill_manual(values=dnet.color.palette) + facet_wrap(vars(treatment), ncol = 1)

```

