---
title: "drought-net-2025"
author: "Janelle Bohey"
date: "2025-06-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Packages

```{r}
library(bipartite)
library(vegan)
library(phytools)
library(taxize)
library(brranching)
library(rotl)
library(dendextend)
library(tidyverse)
library(heatmaply)
library(googlesheets4)
library(broom)
library(dplyr)


library(bipartite)
library(vegan)
library(phytools)
library(taxize)
library(rotl)
library(dendextend)
library(tidyverse)
library(heatmaply)
library(googlesheets4)
library(tibble)
library(lme4)
library(lmerTest)
library(broom)
library(tidyverse)
library(tidyr)
library(vegan)
library(dplyr)
library(broom.mixed)
library(car)
library(emmeans)
library(knitr)
```


# Metadata 
```{r}

#treatments 
#loads in plot and transect codes paired with treatments (rainout, control)
rainout.treatments <- read_sheet("1YiOm8el0IGCvUHm8xDRxf8U9IWiwI8uNT6KGCzZP2sY") %>% 
                  mutate(transectplot = as.character(transectplot))

```

# Tagged Plants
```{r}


#TODO: combine 2024 and 2025; filter so no repeat tag numbers; count species per plot and number of tags per species per plot

# #code to figure out how to randomly sample for scent - choose 1 tagged plant per plot 
tagged_plants_2025 <- read_sheet("1xWH7fVRm352wHamgUgpa1hkvj_DPWLRnzoZ9cCY6gpk")

tagged_plants_2024 <- read_sheet("1CQpUDOA6vYTmdWZ6qwkOehrOVDI3ZZbw2jYZquq4IwU", sheet = "Sheet1")

all_tagged_plants <- tagged_plants_2024 %>% full_join(tagged_plants_2025, by = c("date"))



```

# Soil Moisture 

```{r}

soil_moisture <- read_sheet("1WE6tQZkowFBW0A5h5wrFIRIXZ3p4bs5qPBXu1PPG5XU") %>% 
                  drop_na(vwc) %>% 
                  pivot_longer(vwc, names_to="position", values_to="VWC") %>%      
                  group_by(plot, transect, transectplot, date) %>% 
 
                  summarize(VWC = mean(VWC, na.rm=TRUE)) %>% 
                  left_join(rainout.treatments)
                    
vwc.lmer.2025 <- lmer(VWC~treatment + (1|date), data = soil_moisture)


anova(vwc.lmer.2025) #p=
summary(vwc.lmer.2025)

plot_soil_moisture <- ggplot(soil_moisture, aes(x=factor(date), y=VWC, color=treatment))+geom_boxplot()+ 

   
    labs(y="VWC (%)", x="Date", color="Treatment") +
  scale_color_manual(values = c("skyblue","orange"))+
  theme_minimal()+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(panel.grid.major.x = element_blank(), legend.position = "right") +
  #scale_y_continuous(limits=c(0, NA))
  theme(strip.text.x = element_text(size = 11))
  
print(plot_soil_moisture)

#ggsave(plot.total.emissions.field, file='plots-nitrogen/total_emissions_field_boxplot.jpg', width=150, height=140, units= 'mm', dpi=600)


#install.packages("effsize")  # only need to run this once
library(effsize)  
soil_moisture_cohend <- cohen.d(
  soil_moisture$VWC,
  soil_moisture$treatment,
  pooled = TRUE # pooled standard deviation
)

soil_moisture_cohend

```

# Flower Number 

```{r}
#flower counts
#loads in flower counts for each transect for each day of pollinator observations
visrate.flowercounts.25<- read_sheet("1ObUirLHnFpHzpepx_tiXLPFxXfSVSJ6s9IAa4_Kd25o", na="")%>% mutate(flower_number = as.numeric(flower_number)) %>%  left_join(rainout.treatments) %>% 
  #remove species with little or no data:
  filter(!plant%in% c( "campanula", "dandelion","dandilion", "hydro", "orgrape", "or grape", "senecio", "senecio (small)", "lath", "metensia"))

focalspecies.2025 <- c("clay","delphinium", "umb", "flag", "pot", "congesta", "het", "helio", "galium" )
  
#Getting to know the data
#give number of flowers per species in each week number
visrate.flowercounts.25 %>%
  count(plant, week_num_flowcounts) %>%
  arrange(plant, week_num_flowcounts) %>% print(n=20)

#fill zeros into the data
#records what species are found in which plot
#Makes sure species absent from a plot are never zero-filled
species.present.by.plot.df <- visrate.flowercounts.25 %>%
  distinct(plant, treatment, transect, plot, transectplot) %>%
  group_by(plant, treatment, transect, plot, transectplot) %>%
  summarise(present = TRUE, .groups = "drop")

#Create the full week × plot × species grid
datexweek <- visrate.flowercounts.25 %>% distinct(date, week_num_flowcounts)

phenology.structural.grid.df <- species.present.by.plot.df %>%
  crossing(datexweek)

#Fill structural zeros
flower.number.with.zeros.df <- phenology.structural.grid.df %>% filter(plant %in% focalspecies.2025) %>% 
  left_join(
    visrate.flowercounts.25,
    by = c("plant", "treatment", "transect", "plot", "transectplot" ,"date", "week_num_flowcounts")
  ) %>%
  mutate(
    flower_number = replace_na(flower_number, 0)) %>% 
  mutate(week_num_flowcounts = as.numeric(week_num_flowcounts)) %>% 
  select(-scribe, -notes)


#plot flower number:

#Mean flower number in rainout vs control for each plant species
#(not used for anything but interesting to see raw means)
flower.number.means.2025 <- visrate.flowercounts.25 %>% 
            group_by(treatment, plant) %>% 
            summarise( mean.flower.number = mean(flower_number))
        

plot.flower.num.2025 <- 
  visrate.flowercounts.25 %>% filter(plant %in% focalspecies.2025) %>% 
  ggplot(aes(
    x = str_to_title(plant),
    y = flower_number,
    color = treatment
  )) +
  geom_boxplot() +
  labs(
    x = "",
    y = "Flower number",
    color = "Treatment"
  ) +
  scale_color_manual(values = c("skyblue", "orange")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom"
  )
  plot.flower.num.2025
  
  

#-------make model with only 1 species: -----------
congesta.flwrs <- flower.number.with.zeros.df %>%
  filter(plant == "congesta")

congesta.flwrs.lm <- lm(flower_number ~treatment*week_num_flowcounts, data = congesta.flwrs)
anova(congesta.flwrs.lm)

  library(ggplot2)
  #plot congesta flower number 
  ggplot(congesta.flwrs, aes(x = week_num_flowcounts, y = flower_number, color = treatment)) +
    stat_summary(fun = mean, geom = "line") +
    stat_summary(fun = mean, geom = "point") +
    theme_bw()

#-------- All plants flower number model ----------------
library(purrr)
library(broom)

  
#  “Does treatment change flowering phenology through time?”
  flower.models.25 <- flower.number.with.zeros.df %>%  filter(plant %in% focalspecies.2025) %>% 
    group_by(plant) %>%
    nest() %>%
    mutate(
      model = map(data, ~ lm(flower_number ~ treatment + (1|week_num_flowcounts), data = .x)),
      anova = map(model, ~ tidy(anova(.x))),
      #emmeans = map(model, ~ tidy(emmeans(.x, specs = "treatment")))
      ) %>%
   unnest(anova) %>% 
      filter(term %in% c("treatment", "week_num_flowcounts", "treatment:week_num_flowcounts")) %>%
  mutate(sig = ifelse(p.value < 0.05, "*", ""))

View(flower.models.25)
#congesta, delphinium and heliomeris are significant
  

#....................................

#glmm

#library(glmmTMB)
# flower.number.glmm <- flower.number.with.zeros.df %>% 
#   group_by(plant) %>%
#   filter(n() > 20) %>% 
#   nest() %>% 
#   mutate(
#     model = map(data, ~ glmmTMB(
#       flower_number ~ treatment * factor(week_num_flowcounts) + (1 | transect),
#       family = nbinom2,
#       data = .x
#     ))
#   ) %>% 
# mutate(
#     anova = map(model, ~ car::Anova(.x, type = 2))
#   ) %>%  
#   mutate(
#     emmeans = map(model, ~ emmeans(.x, ~ treatment | factor(week_num_flowcounts)))
#   )
# 
# Anova(
#   flower.number.glmm,
#   type = 2   # or type = 3 if you have strong reasons
# )
# 
#   # flower.emmeans <- flower.models %>% dplyr::select(emmeans) %>% unnest(emmeans)
# 
# #emmeans:
# vrate.flower.emmeans <- visrate.flowercounts.25 %>%
#   group_by(plant) %>%
#   nest() %>%
#   mutate(
#     model = map(data, ~ lm(flower_number ~ treatment * week_num_flowcounts, data = .x)),
#     emmeans = map(model, ~ tidy(emmeans(.x, specs = "treatment")))
#   ) %>%
#   unnest(emmeans)

  
#...............................................................

#CAP preparation: (why are you doing a cap? jb 1/13/26)
#turns flowercount into wide format
# flowercount.wide <- visrate.flowercounts.25 %>% pivot_wider(names_from = plant, values_from = flower_number, values_fill=0) %>% left_join(rainout.treatments)
# 
# #from 2024 code??
# # focalspecies <- c("clay", "delphinium","hydro", "or grape", "pansy","dandilion","scenecio","scenecio (small)", "scenecio (big)", "mertensia", "bochera" )
# # flowercount.numbers <- flowercount.wide [,focalspecies]
# 
# #CAP of flower number for each treatment (???? Nov 22, 2025)
# cap.flowernumber <- capscale(flowercount.numbers~treatment+week_num, data = flowercount.wide)
# 
# #anova on the CAP 
# anova.cap.flowernumber <- anova(cap.flowernumber, by="term")
# print(anova.cap.flowernumber)
#week number significant p=0.001; treatment not significant


```



#vrate phenology
Phenological timing was quantified using proportional flowering. Plot-level proportions were used for mixed-effects models, while species × treatment–level phenology curves were generated by summing flower counts across plots prior to proportional scaling

TODO: change week_num to week_num_flowcounts
```{r}
#Need to run flower.number.with.zeros.df in flower number section before running this code.
flower.number.with.zeros.df 

#Plot-level seasonal totals (needed for proportional phenology)
#Adds up flower number per species per plot
plot.seasonal.flower.totals.df <- flower.number.with.zeros.df %>%
  group_by(plant, treatment, transect, plot, transectplot) %>%
  summarise(
    total_flowers_season = sum(flower_number),
    .groups = "drop"
  )

#Weekly proportional flowering
plot.level.phenology.df <- flower.number.with.zeros.df %>%
  left_join(
    plot.seasonal.flower.totals.df,
    by = c("plant", "treatment", "transect", "plot", "transectplot")
  ) %>%
  mutate(
    prop_flowers_week = if_else(
      total_flowers_season > 0,
      flower_number / total_flowers_season,
      NA_real_)) 
#Cumulative flowering through time

#to determine natural variation in abundance
plot.level.cumulative.phenology.df <- plot.level.phenology.df %>%
  arrange(plant, treatment, transect, plot, transectplot, week_num_flowcounts) %>%
  group_by(plant, treatment, transect, plot, transectplot) %>%
  mutate(
    cumulative_prop_flowers = cumsum(prop_flowers_week)
  ) %>%
  ungroup()

#phenological timing metrics (plot level)
plot.level.phenology.metrics.df <- plot.level.cumulative.phenology.df %>%
  group_by(plant, treatment, transect, plot, transectplot) %>%
  summarise(
    onset_week = min(week_num_flowcounts[cumulative_prop_flowers >= 0.05]),
    peak_week  = week_num_flowcounts[which.max(prop_flowers_week)],
    end_week   = min(week_num_flowcounts[cumulative_prop_flowers >= 0.95]),
    duration_weeks = end_week - onset_week,
    .groups = "drop"
  )

phenological.timing.lmer <- lmer(onset_week ~ treatment * plant + (1|transect/plot), data = plot.level.phenology.metrics.df) 

anova(phenological.timing.lmer)

#.......................................................
#Mean proportional flowering across plots
species.treatment.phenology.df <- plot.level.phenology.df %>%
  group_by(plant, treatment, week_num_flowcounts) %>%
  summarise(
    mean_prop_flowers = mean(prop_flowers_week, na.rm = TRUE),
    se_prop_flowers   = sd(prop_flowers_week, na.rm = TRUE) /
                        sqrt(n()),
    .groups = "drop"
  )

#How does the timing and shape of flowering differ between treatments for each species?
mean.proportional.flowering.plot <- ggplot(
  species.treatment.phenology.df,
  aes(
    x = week_num_flowcounts,
    y = mean_prop_flowers,
    color = treatment,
    fill  = treatment
  )
) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(
      ymin = mean_prop_flowers - se_prop_flowers,
      ymax = mean_prop_flowers + se_prop_flowers
    ),
    alpha = 0.25,
    color = NA
  ) +
  facet_wrap(~ plant, scales = "free_y") +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    x = "Week of season",
    y = "Mean proportional flowering",
    color = "Treatment",
    fill  = "Treatment"
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "italic"),
    legend.position = "top"
  )

print(mean.proportional.flowering.plot)


species.treatment.phenology.df %>%
  group_by(plant, treatment) %>%
  summarise(total = sum(mean_prop_flowers))

#.................................................


#Use this for Species × treatment phenology:  
#Flowering phenology calculated after aggregating flower counts across all plots within each species and treatment.


#Sum flower numbers across plots within each week
species.treatment.weekly.flowers.df <- flower.number.with.zeros.df %>%
  group_by(plant, treatment, week_num_flowcounts) %>%
  summarise(
    total_flowers_week = sum(flower_number),
    .groups = "drop"
  )

#seasonal totals per species × treatment
species.treatment.seasonal.totals.df <- species.treatment.weekly.flowers.df %>%
  group_by(plant, treatment) %>%
  summarise(
    total_flowers_season = sum(total_flowers_week),
    .groups = "drop"
  )

#proportional flowering
species.treatment.phenology.df <- species.treatment.weekly.flowers.df %>%
  filter(plant %in% focalspecies.2025) %>%
  left_join(
    species.treatment.seasonal.totals.df,
    by = c("plant", "treatment")
  ) %>%
  mutate(
    prop_flowers_week = total_flowers_week / total_flowers_season
  ) 

# Flowering curve by species × treatment

#library(scales)
species.treatment.flowering.curve.plot <- ggplot(
  species.treatment.phenology.df,
  aes(
    x = week_num_flowcounts,
    y = prop_flowers_week,
    color = treatment
  )
) +
  geom_line(linewidth = 1.2) +                  # lines for treatment
  #geom_point(size = 1.5) +                      # optional points at weeks
  facet_wrap(~ plant, scales = "fixed") +      # separate panel per species
  scale_y_continuous(
    #labels = percent_format(accuracy = 1),
    limits = c(0, 1)                           # 0 → auto max
  ) +
  scale_x_continuous(breaks = seq(0, 13, by = 2)) +
  labs(
    x = "Week of season",
    y = "Proportional flowering",
    color = "Treatment",
    title = "Species × Treatment Flowering Phenology"
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "italic", size = 11),
    legend.position = "top",
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

# Display the plot
species.treatment.flowering.curve.plot

#.................................................

#Cumulative flowering (aggregated)
#What proportion of total seasonal flowering has occurred by week t?
species.treatment.cumulative.phenology.df <- species.treatment.phenology.df %>%
  arrange(plant, treatment, week_num_flowcounts) %>%
  group_by(plant, treatment) %>%
  mutate(
    cumulative_mean_prop_flowers = cumsum(prop_flowers_week)
  ) %>%
  ungroup()

#plot 
cumulative.flowering.plot <- ggplot(
  species.treatment.cumulative.phenology.df,
  aes(
    x = week_num_flowcounts,
    y = cumulative_mean_prop_flowers,
    color = treatment
  )
) +
  geom_line(linewidth = 1) +
  facet_wrap(~ plant) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent
  ) +
  labs(
    x = "Week of season",
    y = "Cumulative proportional flowering",
    color = "Treatment"
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "italic"),
    legend.position = "top"
  )
print(cumulative.flowering.plot)

#.................................................

#modeling
#make an lmer for each species?
phenology.lmer <- lmer(
  onset_week ~ treatment + (1 | transect/plot),
  data = plot.level.phenology.metrics.df
)

anova(phenology.lmer, by="term")

```



# Nectar - Delphinium 
```{r}

nectar.bud <- read_sheet("1agBw8n2KTjn9A6fB7-xlnjAmuUPiIZbBdVruBfp7EzU")

nectar.old.flw <- read_sheet("1tcEVfHzXJY1YS1TI1_vL7zz6vYmDgV8MdyJeODz5mLM")



```

#Morphology


## Morph metadata
```{r}

#TODO: Inflo height in cm is returning TRUE or 1 - need to fix

morph_2025 <- read_sheet("1grngEjxQ6-uulNl_5yYdbmuajY2DXIGkWfHPBVV-3sM", na=c("")) %>%  
            #remove rows with no tag number
                filter( tag != "") %>% 
                left_join(rainout.treatments)  %>% 
                mutate(average_flower_size = (length+width)/2)

          
#Number of morph measurements per tagged plant 
morph.counts.2025 <-  morph_2025 %>% 
          #filter(plant == "delphinium") %>% 
      group_by(plant, tag, transectplot) %>%
    summarise(
    total_entries = n(),
    depth_measured = sum(!is.na(depth)),
    width_measured = sum(!is.na(width)),
    length_measured = sum(!is.na(length)), 
    .groups = "drop"
  )
View(morph.counts.2025)
#write.csv(morph.counts.2025, "morph.counts.2025.csv", row.names = FALSE)



#.....................................................................

#Making new vector with means of each trait for each transect
flower.size.means.2025 <- morph_2025 %>%
    group_by(transectplot, transect, plant, treatment) %>% #takes mean for each plant 
    summarise(across(c(depth, length, width),
              ~mean(.x, na.rm=TRUE))) %>% 
 #For missing values, changes NaNs to NAs
   mutate(across(
    c(depth, length, width), #TODO: add average_flower_size later (broken for now)
    ~ifelse(is.nan(.x), NA, .x)
  )) #%>%
 # drop_na()

print(flower.size.means.2025)
View(flower.size.means.2025)


all.morph.traits.2025 <- c("length", "width", "depth") #TODO: ADD  "average_flower_size"

morph.models.everything.2025 <- flower.size.means.2025 %>% filter(plant!="galium") %>% 
  filter(!is.na(treatment)) %>% 
  pivot_longer(all_of(all.morph.traits.2025), names_to = "trait") %>% group_by(trait,plant) %>% nest() %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~treatment + (1|transect) , data=.x)),  #TODO: add (1 | transectplot) later
      morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%          
      unnest(morph.Anova)%>%
      dplyr::select(-data) %>% #-morph.trait.lmer
      mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
      filter(term!="(Intercept)") %>% 
       mutate(emmeans=map(morph.trait.lmer, ~tidy(emmeans(.x, specs=c("treatment") )) %>% select(-p.value)))

View(morph.models.everything.2025)
#kable(morph.models.everything.2025, digits=3) #spits out model into table


#emmeans table 

morph.emmeans.2025<- morph.models.everything.2025 %>% dplyr::select(-morph.trait.lmer, -term, -statistic, -df ) %>% 
      #mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
      unnest(emmeans)  
    
kable(morph.emmeans.2025, digits=3) #spits out model into table

#plot morph traits

plot.morph.traits.2025 <- flower.size.means.2025 %>% filter(!is.na(treatment)) %>% 
    pivot_longer(all_of(all.morph.traits.2025), names_to = "trait") %>% 
    ggplot(aes(color=treatment, x=str_to_title(plant), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(depth="Nectar Depth", length= "Flower Length", width="Flower Width")))+
    labs(y="", x="", color="Treatment") +
  scale_color_manual(values = c("skyblue","orange"))+
  theme_minimal()+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(panel.grid.major.x = element_blank(), legend.position = "bottom") +
  #scale_y_continuous(limits=c(0, NA))
  theme(strip.text.x = element_text(size = 11))
                                     
print(plot.morph.traits.2025)

```

#inflo height

```{r}
#TODO: need to fix, made changes just so code would print but need to fix! 

inflo_height_2025 <- morph_2025 %>% 
      #Adds 4mm to all mm inflo measurements   
           mutate(inflo_height_mm = ifelse(!is.na(inflo_height_add4mm_in_mm), inflo_height_add4mm_in_mm + 4, NA)) %>% 
        #convert mm to cm
        mutate(inflo_height_mm_to_cm = inflo_height_mm /10) %>% 
    #clugie way of getting cm data into R
      #For some reason the inflo_height_cm column is only showing up as TRUE or blank value in R. In the google sheet I made a separate column which has both cm and mm inflo height data. The follow code is saying, if the value is TRUE in the inflo_height_cm column then take the value from the inflo_height_ALL_mm_cm column and put it in a new column called height_cm.
        mutate(
    inflo_height_cm_logical = inflo_height_cm == "TRUE",  # creates a logical column
    height_cm = case_when(
      inflo_height_cm_logical ~ as.numeric(inflo_height_ALL_mm_cm),
      TRUE ~ NA_real_
    )
  ) %>% filter(inflo_height_ALL_mm_cm != 0) %>% 
            mutate(inflo_height_ALL_mm_cm = as.numeric(inflo_height_ALL_mm_cm))
  
  
    #combine columns with inflo height converted to cm
  #         mutate(inflo_height_cm_all_final = coalesce(height_cm, inflo_height_mm_to_cm)) %>% 
  # drop_na()


#Overall model 

inflo.lmer.2025 <- lmer(inflo_height_ALL_mm_cm ~ treatment + (1 | transect), data = inflo_height_2025)
summary(inflo.lmer.2025)
Anova(inflo.lmer.2025, type=3)            
  
#model for each species

inflo_models_by_species <- inflo_height_2025 %>%
  group_by(plant) %>%
  nest() %>%
  mutate(model = map(data, ~ lmer(inflo_height_ALL_mm_cm ~ treatment + (1 | transect) + (1 | transect:plot) +(1|tag), data = .x))) %>% 
  mutate(summary = map(model, ~ tidy(.x, effects = "fixed"))) %>%
  unnest(summary) %>%
  mutate(
    signif = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ ""
    )
  ) %>%
  select(plant, term, estimate, std.error, statistic, p.value, signif)

View(inflo_models_by_species)
```

#.......................................

# Visitation Rate 

```{r}
visitation.rate.25 <- read_sheet("1Fogbr6sgAxgfnCr5DWvaeCdJNHonhp_mn--mfSsvhCo") %>% 
       mutate(week_num_vrate=factor(week_num_vrate))  
        
              
```

#Getting to know the vrate data
```{r}

#How many interactions each plant had:
#Q: How many total insect visits did each plant receive?

plant.interactions.2025 <- visitation.rate.25 %>%
  group_by(plant) %>%
  summarise(
    total_visits = sum(visits, na.rm = TRUE),
    n_records = n(), #how many observation rows involved that plant; NA = no visits for 15 min obs period
    n_insect_classes = n_distinct(insect_class)
  ) %>%
  arrange(desc(total_visits))

#How many visits each insect_class made
#Q: Which insect groups are doing the visiting?
insect.visits.2025 <- visitation.rate.25 %>%
  group_by(insect_class) %>%
  summarise(
    total_visits = sum(visits, na.rm = TRUE),
    n_records = n(),
    n_plants_visited = n_distinct(plant)
  ) %>%
  arrange(desc(total_visits))

#How many visits per transectplot
transectplot.visits.2025 <- visitation.rate.25 %>%
  group_by(transectplot) %>%
  summarise(
    total_visits = sum(visits, na.rm = TRUE),
    n_records = n(),
    n_plants = n_distinct(plant),
    n_insect_classes = n_distinct(insect_class)
  ) %>%
  arrange(desc(total_visits))

#visits are not normally distributed
hist(visitation.rate$visits, breaks = 30)

```

#vistitation rate cleaning
```{r}
#TODO: Separate presence/absence vs intensity of visitation; Calculate visitation rates (per time, per flower, per plot); Build plant × insect interaction matrices; Diagnose zero inflation before GLMMs

#define insect_classes
vrate.insect.class.2025 <- visitation.rate.25 %>% 
  mutate(
    insect_class = if_else(
      is.na(insect_class),
      "no_insects",
      insect_class
    )
  ) %>% 
  distinct(insect_class)

# define plants observed
vrate.plants.2025 <- visitation.rate.25 %>%
  distinct(plant)

# define observation windows
#raw visitation data contains multiple rows per observation window (one per insect class, per visit)
#distinct() answers: “When and where did we observe, regardless of what we saw?”
#extracts unique 15-minute observation windows and makes each row an observation event
#one row per date × time × plot; no insect or plant information yet
observation.windows.2025.df <- visitation.rate.25 %>%
  distinct(
    date, week_num_vrate, start, 
    transect, plot, transectplot)

# create zero-filled visitation structure
visitation.rate.complete.2025.df <- observation.windows.2025.df %>%
  crossing(vrate.plants.2025, vrate.insect.class.2025) %>%
  left_join(
    visitation.rate.25,
    by = c(
      "date", "week_num_vrate", "start", "end",
      "transect", "plot", "transectplot",
      "plant", "insect_class"
    )
  ) %>%
  mutate(visits = replace_na(visits, 0)) %>% 
  mutate(plant = if_else(is.na(plant), "no_visits", plant))

#visits == 0 == TRUE → you watched and did not see that insect class
#visits == 0 == FALSE → you did see that insect class at least once in that observation window
visitation.rate.complete.2025.df %>%
  count(insect_class, visits == 0)


```

```{r}
#combine visitation rate and flower counts for each plot
#how many total visits did each plant species get for each plot each week. 

# Need to group flower number df to remove many-to-many warning later: 

#filter vrate based on phenology: 
flower.availability.week.df <- plot.level.cumulative.phenology.df %>%
  filter(!plant%in% c( "campanula", "dandelion","dandilion", "hydro", "orgrape", "senecio", "senecio (small)")) %>% 
  group_by(
    date,
    week_num_flowcounts,
    transectplot,
    transect,
    plot,
    plant
  ) %>%
  mutate(flowers_present = flower_number > 0) %>% 
  filter(!is.na(plant), plant!="clay") %>% 
  filter(!is.na(plant), plant!="bochera") %>% 
  filter(!is.na(plant), plant!="or grape")

# Aggregate total visits per plant per plot per week
visitation.rate.by.plant.week.2025.df <- visitation.rate.complete.2025.df %>%
  left_join(rainout.treatments, by = c("transect", "plot", "transectplot")) %>% 
  mutate(week_num_vrate = as.numeric(as.character(week_num_vrate))) %>% 
  filter(!is.na(insect_class), plant!="clay") %>% 
  group_by(
    date,
    #week_num_vrate, #I think id be better to make a total season week count and add that
    transect,
    plot,
    transectplot,
    plant, insect_class, visits, plants_visisted
  ) %>%
  summarise(
    total.visits = sum(visits),
    .groups = "drop")

#USE THIS: 
#filters visitation rate by plant x week to only include if flower number is >0
vrate.2025.filtered <-
  visitation.rate.by.plant.week.2025.df %>%
  filter(insect_class != "wasp") %>% 
  filter(insect_class != "no_insects") %>% 
  filter(insect_class != "beetle") %>% 
  filter(insect_class != "moth") %>% 
  left_join(
    flower.availability.week.df,
    by = c(
      "date",
      "transect",
      "plot",
      "transectplot",
      "plant")) %>%
  filter(flowers_present) %>% 
   mutate(
    # Visits per hour (for 15-min observation windows)
    visits.per.hr = total.visits * 4,

   # Visits per flower per hour: only if flower_number > 0
    visitsperflowerperhr = if_else(
      !is.na(flower_number) & flower_number > 0,
      (total.visits / flower_number) * 4,
      NA_real_)) %>% 
  mutate(plant_insect_class = paste0(plant, insect_class))

############################


# flower.number.week.2025.df <- flower.number.with.zeros.df %>%
#   group_by(
#     date,
#     week_num_flowcounts,
#     transectplot,
#     transect,
#     plot,
#     plant
#   ) %>% filter(!is.na(plant), plant!="clay") %>% 
#   summarise(
#     flower_number = mean(flower_number, na.rm = TRUE),
#     .groups = "drop"
#   )
# 
# unique(flower.number.week.2025.df$plant)
# unique(flower.number.week.2025.df$date)
# 
# # Aggregate total visits per plant per plot per week
# visitation.rate.by.plant.week.2025.df <- visitation.rate.complete.2025.df %>%
#   left_join(rainout.treatments, by = c("transect", "plot", "transectplot")) %>% 
#   mutate(week_num_vrate = as.numeric(as.character(week_num_vrate))) %>% 
#   filter(!is.na(insect_class), plant!="clay") %>% 
#   group_by(
#     date,
#     #week_num_vrate, #I think id be better to make a total season week count and add that
#     transect,
#     plot,
#     transectplot,
#     plant, insect_class, visits, plants_visisted
#   ) %>%
#   summarise(
#     total.visits = sum(visits),
#     .groups = "drop")
# 
# #What plant species are in this dataframe:
# unique(visitation.rate.by.plant.week.2025.df$plant)
# 
# setdiff(visitation.rate.by.plant.week.2025.df$plant, flower.number.week.2025.df$plant)
# 
# #No interaction data for ther following plantds:
# setdiff( flower.number.week.2025.df$plant, visitation.rate.by.plant.week.2025.df$plant)
# 
# 
# # Join flower counts
# visitation.rate.with.flowers.2025.df <- visitation.rate.by.plant.week.2025.df %>%
#   
#   #add flower counts to vrate data
#   left_join(
#     flower.number.week.2025.df, by = c("date", "transect", "plot", "transectplot", "plant")) %>% 
#   mutate(
#     # Visits per hour (for 15-min observation windows)
#     visits.per.hr = total.visits * 4,
# 
#    # Visits per flower per hour: only if flower_number > 0
#     visitsperflowerperhr = if_else(
#       !is.na(flower_number) & flower_number > 0,
#       (total.visits / flower_number) * 4,
#       NA_real_)) %>% 
#       group_by(transect) %>%
#   #adds treatment to newly added possible insect interactions
#   mutate(
#     treatment = first(na.omit(treatment))) %>%
#   ungroup()  %>% 
#   mutate(plant_insect_class = paste0(plant, insect_class))
# 




ggplot(visitation.rate.with.flowers.2025.df, aes(x = treatment, y = visitsperflowerperhr)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.3) +
  scale_y_continuous(trans = "log1p") # optional log-transform for visualization

#average visits for each insect x plant interaction across whole season (based on total number of interactions to that plant by x insect)
nonzero.average.vrate.by.treatment <- vrate.2025.filtered %>% 
              group_by(plant, insect_class, treatment) %>% 
                        summarise(mean.vpfphr = mean(visitsperflowerperhr, na.rm = TRUE),
                                   .groups = "drop") %>% filter(mean.vpfphr>0) %>%
                        mutate(plant_insect_class = paste0(plant, insect_class)) %>% ungroup()

#number of transects that you have both plant and pollinator data for in control and treatment 
weekly.average.plant.insect.vrate <- vrate.2025.filtered %>% 
                      filter(plant_insect_class%in%nonzero.average.vrate.by.treatment$plant_insect_class) %>% 
                      #ungroup() %>% 
  #only includes columns with plant_insect_class, treatment, week_num_flowcounts and number of transects that plant insect combination are found in (ranges 1-4 transects)
                      count(plant_insect_class, treatment, week_num_flowcounts) 

#creat column with paste0(plant_insect_week) so you can add to vrate.2025.filtered
plant.insect.week.25 <- weekly.average.plant.insect.vrate %>% 
  #makes treatment column into two separate columns one for control, another for rainout:                    
                      pivot_wider(names_from = treatment, values_from = n, values_fill = 0) %>%  
                      filter(control>0, rainout>0) %>% #filters out plant insect combinations that are not found in any of the transects (must occur in at least one transect)
                      mutate(plant_insect_week= paste0(plant_insect_class, week_num_flowcounts)) #%>% #creates a new column with plant, insect_class and week_num pasted together
                      pull(plant_insect_week) #only includes the column plant_insect_week so that we can add it to vrate.summary.treatment.transect


#plot





```


#vrate analysis
```{r}
#anova

vrate.resid.var <- vrate.2025.filtered %>%
  group_by(plant_insect_class) %>%
  summarise(resid_var = var(visitsperflowerperhr), .groups = "drop") %>% 
  filter(resid_var > 0) %>%
  pull(plant_insect_class)

#creates a model for each combination of plant and insect group
vrate.anova.model.25 <- vrate.2025.filtered %>% 
                      filter(plant_insect_class %in% vrate.resid.var) %>% 
                      filter(plant_insect_class!="dandilionbeetle") %>% 
                      mutate(plant_insect_week = paste0(plant_insect_class, week_num_flowcounts)) %>% 
                      #filter(plant_insect_week%in%plant.insect.week.25) %>% #adds visitsperflowerperhr column
                      mutate(week_num_flowcounts=factor(week_num_flowcounts)) %>% #make week_num a factor so anova doesn't break
                        group_by(plant_insect_class) %>%
                        filter(n_distinct(treatment) > 1, n_distinct(week_num_flowcounts) > 1) %>%
                        ungroup() %>%
  
                     # nest by plant_insect_class
                      group_by(plant_insect_class) %>%
                      nest() %>%
                      mutate(model=map(data, ~ lm(visitsperflowerperhr~ treatment*week_num_flowcounts, data=.x))) %>% #makes a model for each plant species x pollinator combination
                      mutate(summary = map(model,~ tidy(anova(.x)))) %>% #summarizes each model made in the line before, tidy makes summary into table
                      select(plant_insect_class, summary) %>%  unnest(cols = c(summary)) #expands dataframe
#WARNING - ignore warning message, just means there's not enough data
#NaN means there is not enough data for the model, you can ignore warning in console this is also the reason (not enough data for some plant insect combos)

View(vrate.anova.model.25)



#test overall differences in visitation patterns across treatments while accounting for multivariate structure, instead of just looking at one plant × insect combination at a time.

#TODO: set up cap, run cap and anova

#do I want visits per flower per hour for each speices or treat. MEan vpfph for plant_insect_ class combination

```

# multivariate analysis on visitation rate
```{r}
#cap by week (6 different caps)
#we did cap by week (instead of cap on whole summer vrate) bc you'd see week effect on whole summer vrate cap but this by week cap says is there a difference by week visitation rate. Maybe there's differences when a certain plant or pollinator is present in the network.

vrate.cap.byweek.25 <- vrate.2025.filtered %>% 
             #filter(visits.per.hr >0) %>% #filters column plant_insect_class but the list of plant x pollinator combinations with visits per flower per hr greater than 0 (which means there is data for them)
             # aggregate visits per week x transect x treatment x plant
              group_by(week_num_flowcounts, transect, treatment, plant_insect_class) %>%
              summarise(visitsperflowerperhr = mean(visitsperflowerperhr), .groups = "drop") %>%          
  
              # pivot to wide format
              pivot_wider(
                names_from = plant_insect_class,
                values_from = visitsperflowerperhr,
                values_fill = 0) %>%
  
            group_by(week_num_flowcounts) %>%  
            nest() %>% 
            mutate(model=map(data, ~ capscale(.x[,-(1:3)] ~ treatment, data=.x))) %>%   #cap minus the first two cols which are transect and treatment 
            mutate(test= map(model, ~tidy(anova(.x, by="term")))) %>%  unnest(cols = c(test))

View(vrate.cap.byweek.25)
anova(vrate.cap.byweek.25, by="term")

```



#For each pollinator group did they make different amounts of visits (to any plant) between treatments? Are there more bees in the drought treatment?
```{r}

poll.vrate.plot.25 <- ggplot(vrate.2025.filtered %>% #filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class) %>% 
                          group_by(insect_class, week_num_flowcounts, treatment, transect) %>% 
                          summarise(visitsperflowerperhr=sum(visitsperflowerperhr)), 
                          aes(x=week_num_flowcounts, y=visitsperflowerperhr, color=treatment, group = paste0(treatment, transect))) + 
                          facet_wrap(vars(insect_class)) + geom_point() + geom_line()+ scale_color_manual(values=dnet.color.palette) + theme_minimal() + ylab("Visits per flower per hour")+ xlab("Week Number") +  theme(strip.text.x = element_text(size = 11)) +  theme(axis.text.x=element_text(size=10), axis.title.x = element_text(size=11), axis.text.y = element_text(size=10), axis.title.y = element_text(size=12))  

print(poll.vrate.plot.25)

#ggsave(poll.vrate.3.plot, file='poll.vrate.3.plot.jpg', width=140, height=70, units= 'mm', dpi=600)


#...........................................
# poll.vrate.3.plot <- ggplot(vrate.summary.treatment.transect %>% 
# filter(insect_class %in% c("big bumblebee", "big lepidoptera", "big solitary bee"))  %>%  
#   mutate(insect_class = dplyr::recode(insect_class, 
#     "big bumblebee" = "Bumblebee",
#     "big lepidoptera" = "Lepidoptera",
#     "big solitary bee" = "Solitary bee")) %>% 
#   filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class) %>% 
#                           group_by(insect_class, week_num, treatment, transect) %>% 
#                           summarise(visitsperflowerperhr=sum(visitsperflowerperhr)), 
#                           aes(x=week_num, y=visitsperflowerperhr, color=treatment, group = paste0(treatment, transect))) + 
#                           facet_wrap(vars(insect_class)) + geom_point() + geom_line()+ scale_color_manual(values=dnet.color.palette) + theme_minimal() + ylab("Visits per flower per hour")+ xlab("Week Number") +  theme(strip.text.x = element_text(size = 13)) +  theme(axis.text.x=element_text(size=12), axis.title.x = element_text(size=11), axis.text.y = element_text(size=12), axis.title.y = element_text(size=14))  
# 
# print(poll.vrate.3.plot)
# 
# ggsave(poll.vrate.plot, file='poll.vrate.plot.jpg', width=140, height=70, units= 'mm', dpi=600)

#.......................................
#anova
insect.class.anova.25 <- vrate.2025.filtered %>% 
                    
                      mutate(week_num_flowcounts=factor(week_num_flowcounts)) %>% #make week_num a factor so anova doesn't break
                      group_by(insect_class, week_num_flowcounts, treatment, transect) %>% 
                      summarise(visitsperflowerperhr=sum(visitsperflowerperhr)) %>% 
                      group_by(insect_class) %>% nest() %>% 
                      mutate(model=map(data, ~ lm(visitsperflowerperhr~ treatment*week_num_flowcounts, data=.x))) %>% #makes a model for each plant species x pollinator combination
                      mutate(summary = map(model,~ tidy(anova(.x)))) %>% #summarizes each model made in the line before, tidy makes summary into table
                      select(insect_class, summary) %>%  unnest(summary)

View(insect.class.anova.25)

```


# For each plant species, is there a difference in pollinator visitation between treatments?

```{r}

ggplot(vrate.2025.filtered %>% filter(!plant%in% c( "campanula", "dandelion","dandilion", "hydro", "orgrape", "senecio", "senecio (small)")) %>% 
                          group_by(plant, week_num_flowcounts, treatment, transect) %>% 
                          summarise(visitsperflowerperhr=sum(visitsperflowerperhr)), 
                          aes(x=week_num_flowcounts, y=visitsperflowerperhr, color=treatment, group = paste0(treatment, transect))) + 
                          facet_wrap(vars(plant)) + geom_point() + geom_line()+ scale_color_manual(values=dnet.color.palette) + theme_minimal() + ylab("Visits per flower per hour")+ xlab("Week Number") +  theme(strip.text.x = element_text(size = 11)) + theme(axis.text.x=element_text(size=12))

#anova
plant.anova.25 <- vrate.2025.filtered %>% 
                      filter(!plant%in% c("galium","sedum", "sub", "rosa", "speciosus")) %>% #these species don't have enough interaction data
                      mutate(week_num_flowcounts=factor(week_num_flowcounts)) %>% #make week_num a factor so anova doesn't break
                      group_by(plant, week_num_flowcounts, treatment, transect) %>% 
                      summarise(visitsperflowerperhr=sum(visitsperflowerperhr)) %>% 
                      group_by(plant) %>% nest() %>% 
                      mutate(model=map(data, ~ lm(visitsperflowerperhr~ treatment*week_num_flowcounts, data=.x))) %>% #makes a model for each plant species x pollinator combination
                      mutate(summary = map(model,~ tidy(anova(.x)))) %>% #summarizes each model made in the line before, tidy makes summary into table
                      select(plant, summary) %>%  unnest(summary)

View(plant.anova.25)

```

#....................
# Compare 2024 & 2025


read in dataframes: 
```{r}





```


#....................
# Transect walks visitation by insect class
## do more insects visit control vs drought plots?

```{r}
#load metadata: 
transect_walks <- read_sheet("1YujI3IvVgb2Z4d-srCpmBw2GzixHL3PYhDe2Wlt4eGQ", na=c("na","","NA") ) %>% 
                   mutate(transectplot = as.character(transectplot), Transects = as.character(Transects)) %>% 
                  left_join(rainout.treatments) %>% 
                  filter(!is.na(Caught_time)) %>% 
                  filter(!is.na(treatment))

#Do more insects visit control vs drought plots? 
# Summarize to counts per transect walk per week
insects_per_transect_tw <- transect_walks %>%
  group_by(Week, Transects,treatment) %>%
  summarise(total_count = n(), .groups = "drop") 
#total count is the number of insects caught in each transect in each week
  
View(insects_per_transect_tw)

# Mixed-effects model
# Random effect for transect ID (since each is sampled repeatedly) and week
transect_walks_model <- lmer(total_count ~ treatment + 
                      (1|Transects) + (1|Week),
                    data = insects_per_transect_tw)
summary(transect_walks_model)
#p= 0.256

# Get model-predicted means and SEs
#emmeans.transect.walks <- emmeans(transect_walks_model, ~ treatment) %>%
 # as.data.frame()

#plot total insects per treatment: 
# Plot: Total insects per transect walk, by treatment


# Plot: Total insects per transect walk, by treatment
plot.insects <- insects_per_transect_tw %>%
  ggplot(aes(color = treatment,
             x = str_to_title(Transects),  # Capitalize transect labels
             y = total_count)) +
  geom_boxplot() +
  labs(
    y = "Number of Insects",
    x = "Transect Pairs",
    color = "Treatment"
  ) +
  scale_color_manual(values = c("skyblue", "orange")) +  # match your color palette
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    strip.text.x = element_text(size = 11)
  ) +
  expand_limits(y = 0)

plot.insects

#Weekly counts
# Summarize insects per week & treatment
weekly_counts <- insects_per_transect_tw %>%
  filter(!is.na(total_count), !is.na(treatment)) %>%
  group_by(Week, treatment) %>%
  summarise(
    mean_count = mean(total_count),
    se = sd(total_count) / sqrt(n()),
    .groups = "drop"
  ) %>% 
    mutate(Week = as.numeric(Week))

#plot by week number of insects in rainout vs control 
# Plot by week
ggplot(weekly_counts, aes(x = Week, y = mean_count, color = treatment, group = treatment)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = mean_count - se, ymax = mean_count + se),
    width = 0.2
  ) +
   scale_x_continuous(
    breaks = unique(weekly_counts$Week),
    labels = paste0("Week ", unique(weekly_counts$Week))
  ) +
  scale_color_manual(values = c("control" = "skyblue", "rainout" = "orange")) +
  labs(
    x = "",
    y = "Average Insects",
    color = "Treatment"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )

#####
# Weekly counts (sum across all transects for that treatment/week)
insects_per_week <- transect_walks %>%
  filter(!is.na(Caught_time), !is.na(treatment)) %>%
  group_by(Week, treatment) %>%
  summarise(weekly_count = n(), .groups = "drop")


weekly_model <- lmer(
  weekly_count ~ treatment + (1 | Week),
  data = insects_per_week
)
summary(weekly_model)


#######

# Filter for week 6
week6_data <- insects_per_transect_tw %>% 
  filter(Week == 6)

# Poisson GLM
week6_glm <- glm(total_count ~ treatment, data = week6_data, family = poisson)
summary(week6_glm)
# p = 0.00698 **

# Filter for week 2
week2_data <- insects_per_transect_tw %>% 
  filter(Week == 2)

# Poisson GLM
week2_glm <- glm(total_count ~ treatment, data = week2_data, family = poisson)
summary(week2_glm)
# p = 0.00698 **

##Results text: To test whether the number of insect visits differed between rainout and control treatments, we analyzed weekly insect counts per transect using generalized linear models (GLMs) with a Poisson error distribution. Each GLM included treatment as a fixed effect and total_count (number of insects per transect) as the response variable.  

# Didn't do this but: Analyses were performed separately for each week to assess temporal differences in treatment effects. Model assumptions were checked by examining residual plots and overdispersion statistics.

#All analyses were conducted in R (version X.X.X) using the glm() function for Poisson models and glm.nb() from the MASS package for negative binomial models.


###################
# Sum all weeks together
insects_per_transect <- insects_per_transect_tw %>%
  group_by(Transects, treatment) %>%
  summarise(total_count_allweeks = sum(total_count), .groups = "drop")

#Now total_count_allweeks is the total insects observed per transect over the whole sampling period.

#This model tests if rainout vs control differs within transects, accounting for repeated sampling of the same transect.
# GLMM with transect as random effect
transect_glmm <- glmer(total_count ~ treatment + (1|Transects),
                        data = insects_per_transect_tw,
                        family = poisson)
summary(transect_glmm)
#p=0.0486 *

#plot
ggplot(insects_per_transect, aes(x = treatment, y = total_count_allweeks)) +
  geom_boxplot(aes(fill = treatment), alpha = 0.5) +
  geom_jitter(width = 0.2, size = 2, alpha = 0.8) +
  labs(
    x = "Treatment",
    y = "Total insects per transect (all weeks)",
    title = "Insect visits per transect by treatment"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

summary_plot <- insects_per_transect_tw %>%
  group_by(treatment) %>%
  summarise(
    mean_count = mean(total_count),
    se_count = sd(total_count)/sqrt(n()),
    .groups = "drop"
  )
##########################

#..................
ggplot(summary_plot, aes(x = treatment, y = mean_count, fill = treatment)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  geom_errorbar(aes(ymin = mean_count - se_count, ymax = mean_count + se_count), width = 0.2) +
  labs(
    x = "Treatment",
    y = "Mean Number of Insects"
    #title = "Average insect visits per transect by treatment"
  ) +
  scale_fill_manual(values = c("control" = "skyblue", "rainout" = "orange")) +
  theme_minimal() +
  theme(legend.position = "none")

################
ggplot(insects_per_transect, aes(x = treatment, y = total_count_allweeks, group = Transects)) +
  geom_line(color = "gray", alpha = 0.5) +
  geom_point(aes(color = treatment), size = 3) +
  labs(
    x = "Treatment",
    y = "Total insects per transect",
    title = "Paired transect comparison: rainout vs control"
  ) +
  theme_minimal()
```
## do specific insect classes visit drought or control more?

```{r}
#Do specific insect classes visit drought or control more?
# 1. Count insects per class, per transect walk.
# 2. Fit a mixed model with Transect_plot × Insect_class interaction.
# 3. Run post-hoc tests to see which classes differ.

# Count insects per class per transect walk
counts_by_class <- transect_walks %>%
  group_by(Week, Transects, transectplot, Insect_class, treatment) %>%
  summarise(count = n(), .groups = "drop") %>% 
  filter(!is.na(treatment)) %>% 
  complete(Week, Transects, Insect_class, fill = list(count = 0)) %>% 
    # Ensure grouping variables are factors
  mutate(
    Transects = as.character(Transects),
    Week = as.character(Week),
    transectplot = as.character(transectplot),
    Insect_class = as.factor(Insect_class)
  )
View(counts_by_class)

#mixed effects model 

library(glmmTMB)
# Model: effect of treatment for each insect class
# Poisson GLMM (or try nbinom2 if overdispersed)
model_by_class <- glmmTMB(
  count ~ treatment * Insect_class + 
    (1|Transects) + (1|Week),
  family = poisson,
  data = counts_by_class
)

summary(model_by_class)
# no sig p-values
```

#---------------
#REU Project


# delphinium morphology metadata & find plant averags for morph traits
```{r}

delph.flower.size <- read_sheet("1grngEjxQ6-uulNl_5yYdbmuajY2DXIGkWfHPBVV-3sM",  na=c("na","","NA")) %>% 
      left_join(rainout.treatments) %>% 
    filter(plant%in%c("delphinium","flagellaris","congesta")) %>% 
    filter( tag != "") %>% 
      mutate(width = as.character(width)) %>% 
  mutate(width = na_if(width, "NULL"))  %>% 
         mutate(width = as.numeric(width), length = as.numeric(length), inflo_height_add4mm_in_mm = as.numeric(inflo_height_add4mm_in_mm)) %>% 
          mutate(average_flower_size = if_else(
          plant %in% c("delph", "congesta"),
            (length + width) / 2,
             length ))

#TODO: Add average per tag!!
  
#table of # of plant species in each treatment
with(delph.flower.size, table(plant, treatment))

#Making new vector with means of each trait for each transect
delph.flower.size.means <- delph.flower.size %>%
    group_by(transect, plant) %>% #takes mean for each plant 
    summarise(across(c(depth, length, width, inflo_height_add4mm_in_mm),
              ~mean(.x, na.rm=TRUE))) %>% 
  mutate(across(
    c(depth, length, width, inflo_height_add4mm_in_mm),
    ~ifelse(is.nan(.x), NA, .x)
  )) #%>% 
  #drop_na()

print(delph.flower.size.means)
View(delph.flower.size.means)

all.morph.traits.2025 <- c("length", "width", "depth", "inflo_height_add4mm_in_mm", "average_flower_size")

#TODO: do we need tag means to run in model?
delph.morph.models.everything.2025 <- delph.flower.size %>% pivot_longer(all_of(all.morph.traits.2025), names_to = "trait") %>% group_by(trait,plant) %>% nest() %>% 
      mutate(morph.trait.lmer=map(data, ~lmer(value~treatment + (1|transect) + (1 | transect:plot) + (1|tag), data=.x)), 
      morph.Anova= map(morph.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%          
      unnest(morph.Anova)%>%
      dplyr::select(-data) %>% #-morph.trait.lmer
    ungroup(trait) %>%   #but want to keep group by plant so might want to change 
    mutate(
    p.adj = p.adjust(p.value, method = "fdr"),
    significant = ifelse(p.adj < 0.05, "*", "")
  )  %>% 
      filter(term!="(Intercept)") %>% 
       mutate(emmeans=map(morph.trait.lmer, ~tidy(emmeans(.x, specs=c("treatment") )) %>% select(-p.value)))

View(delph.morph.models.everything.2025)
#kable(morph.models.everything.2025, digits=3) #spits out model into table


#emmeans table 

delph.morph.emmeans<- delph.morph.models.everything.2025 %>% dplyr::select(-morph.trait.lmer, -term, -statistic, -df ) %>% 
      #mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
      unnest(emmeans)  
    
kable(delph.morph.emmeans, digits=3) #spits out model into table

#plot morph traits

figure.traits <- c("depth", "inflo_height_add4mm_in_mm", "average_flower_size")


plot.delph.traits <- delph.flower.size %>% 
    pivot_longer(all_of(figure.traits), names_to = "trait") %>% 
      mutate(
    plant = fct_recode(plant,
       "D. nuttallianum"= "delphinium" ,
       "E. flagellaris"= "flagellaris",
       "E. congesta" = "congesta"
      # Add more recoding here if needed
    ),
    plant = fct_relevel(plant, 
      "D. nuttallianum", 
      "E. flagellaris",
      "E. congesta"
      # Put them in the desired order here
    )
  ) %>% 
  
    ggplot(aes(color=treatment, x=str_to_title(plant), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(depth="Nectar Depth", length= "Flower Length", width="Flower Width", inflo_height_add4mm_in_mm="Inflorescence Height", average_flower_size ="Average Flower Size")))+
    labs(y="", x="", color="Treatment") +
  expand_limits(y = 0)+
  scale_color_manual(values = c("skyblue","orange"))+
  theme_minimal()+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(panel.grid.major.x = element_blank(), legend.position = "bottom") +
  #scale_y_continuous(limits=c(0, NA))
  theme(strip.text.x = element_text(size = 11))
                                     
print(plot.delph.traits)
```


#Seeds - delphinium only
```{r}

seed_counts <- read_sheet("1z5DuogRE8C-wbcRLjBXV9OTpDxfRxh1GF4WDKeh8w4o", sheet = "delph",  na=c("")) %>% 
                   left_join(rainout.treatments) %>% 
                    select(species, tag, transect, plot, transectplot, weight, fruit_number, aborted_fruits,	seed_number, treatment) %>% 
                  mutate(tag = as.character(tag)) %>%   
                mutate(tag = na_if(tag, "NULL")) %>% 
                filter(!is.na(tag))

all.seed.traits.2025 <- c("weight", "fruit_number", "aborted_fruits", "seed_number")

seed.models.2025 <- seed_counts %>% pivot_longer(all_of(all.seed.traits.2025), names_to = "trait") %>% group_by(trait,species) %>% nest() %>% 
      mutate(seed.trait.lmer=map(data, ~lmer(value~treatment + (1|transect)+(1 | transect:plot), data=.x)), 
      seed.Anova= map(seed.trait.lmer, ~tidy(Anova(.x, type=3)))) %>% 
      unnest(seed.Anova)%>%
      dplyr::select(-data) %>% #-morph.trait.lmer
      mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
      filter(term!="(Intercept)") %>% 
       mutate(emmeans=map(seed.trait.lmer, ~tidy(emmeans(.x, specs=c("treatment") )) %>% select(-p.value)))

View(seed.models.2025)

###
#emmeans table 

seed.emmeans<- seed.models.2025 %>% dplyr::select(-seed.trait.lmer, -term, -statistic, -df ) %>% 
      #mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
      unnest(emmeans)  
    
kable(seed.emmeans, digits=3) #spits out model into table


#plot morph traits

plot.seed.traits <- seed_counts %>% 
    pivot_longer(all_of(all.seed.traits.2025), names_to = "trait") %>% 
    ggplot(aes(color=treatment, x=str_to_title(species), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(weight="Weight (g)", fruit_number= "Number of Fruits", seed_number="Seed Set", aborted_fruits="Aborted Fruits")))+
    labs(y="", x="", color="Treatment") +
  scale_y_continuous(limits = c(0.06, NA))+
  expand_limits(y = 0)+
  scale_color_manual(values = c("skyblue","orange"))+
  theme_minimal()+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(panel.grid.major.x = element_blank(), legend.position = "bottom") +
  #scale_y_continuous(limits=c(0, NA))
  theme(strip.text.x = element_text(size = 11)) 
                                     
print(plot.seed.traits)

```

#Flag seed analyses
```{r}

seed_counts_flag <- read_sheet("1z5DuogRE8C-wbcRLjBXV9OTpDxfRxh1GF4WDKeh8w4o", sheet = "flag",  na="") %>% 
                   left_join(rainout.treatments) %>%  
                    filter(!is.na(flower_number) & !is.na(seed_number)) %>% 
                    
                    select(species, tag,transect, plot, transectplot, flower_number, seed_number, treatment)

flag.seed.traits.2025 <- c("flower_number", "seed_number")

flag.seed.models.2025 <- seed_counts_flag %>% pivot_longer(all_of(flag.seed.traits.2025), names_to = "trait") %>% group_by(trait,species) %>% nest() %>% 
      mutate(flag.seed.trait.lmer=map(data, ~lmer(value~treatment + (1|transect)+(1 | transect:plot), data=.x)), 
      flag.seed.Anova= map(flag.seed.trait.lmer, ~tidy(Anova(.x, type=3)))) %>%          
      unnest(flag.seed.Anova)%>%
      dplyr::select(-data) %>% #-morph.trait.lmer
      mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
      filter(term!="(Intercept)") %>% 
       mutate(emmeans=map(flag.seed.trait.lmer, ~tidy(emmeans(.x, specs=c("treatment") )) %>% select(-p.value)))

View(flag.seed.models.2025)

###
#emmeans table 

flag.seed.emmeans<- flag.seed.models.2025 %>% dplyr::select(-flag.seed.trait.lmer, -term, -statistic, -df ) %>% 
      #mutate(significant = ifelse(p.value<0.05, "*","") ) %>% 
      unnest(emmeans)  
    
kable(flag.seed.emmeans, digits=3) #spits out model into table

#plot morph traits

plot.flag.seed.traits <- seed_counts_flag %>% 
    pivot_longer(all_of(flag.seed.traits.2025), names_to = "trait") %>% drop_na() %>% 
    ggplot(aes(color=treatment, x=str_to_title(species), y=value))+geom_boxplot()+
    facet_wrap(vars(trait),scales = "free_y", labeller=as_labeller(c(flower_number="Flower Number", seed_number= "Seed Set")))+
    labs(y="", x="", color="Treatment") +
  expand_limits(y = 0)+
  scale_color_manual(values = c("skyblue","orange"))+
  theme_minimal()+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(panel.grid.major.x = element_blank(), legend.position = "bottom") +
  #scale_y_continuous(limits=c(0, NA))
  theme(strip.text.x = element_text(size = 11))
                                     
print(plot.flag.seed.traits)

```

# seed set X flower size 
#correlation

#want seeds per flower for correlation (right now have total seed number)
```{r}

## Step 1: Remove any seed entries without a tag
#tag needs to be a character not list
seed_counts_flag$tag <- as.character(seed_counts_flag$tag)

delph.flower.size <- delph.flower.size %>%
  mutate(tag = as.character(tag))


#create new dataframe that averages measurements per tag so only 1 flower size value per tag
avg_flower_size_per_tag <- delph.flower.size %>%
  group_by(tag,plant) %>%
  summarize(average_flower_size = mean(average_flower_size, na.rm = TRUE)) %>% 
  # Replace NaN with NA first (optional)
  mutate(average_flower_size = ifelse(is.nan(average_flower_size), NA, average_flower_size)) %>%
  # Then drop rows with NA average_flower_size
  filter(!is.na(average_flower_size)) %>% 
  mutate(tag = as.character(tag))

#bind rows from delph seed_counts and seed_counts_flag, join average flower size
seedxflowersize <-  bind_rows(seed_counts, seed_counts_flag) %>% 
                    mutate(seed_number_sqrt = sqrt(seed_number)) %>% 
                    filter(!tag %in% c("598")) %>% 
                    #filter(!average_flower_size %in% c("598")) %>% 
                    filter(tag %in% avg_flower_size_per_tag$tag) %>%  #only keep tagged plants that are measured
                    left_join(avg_flower_size_per_tag ,
                     by = "tag")

# Duplicate number species delph = 4988 but plant =congesta
#filter(!tag %in% c("4988", "4687"))

#counting how many tags per species
tag_counts_per_species <- seedxflowersize %>%
  group_by(species) %>%              # group by species
  summarize(tag_count = n())          # count rows per species

#gives sample size
print(tag_counts_per_species)

#check for normal distribution to determine what method to use for correlation test
hist(seedxflowersize$average_flower_size, main = "Histogram of Flower Size")
hist(seedxflowersize$seed_number, main = "Histogram of Seed Number")
hist(seedxflowersize$seed_number_sqrt, main = "Histogram of Seed Number Sqrt transformed")
#not normally distributed so use spearman method below


#Correlation Test: 
#(for all species - need to do one test per species)
#need to test for average seeds per flower not seed number
cor.test(seedxflowersize$average_flower_size, seedxflowersize$seed_number, method = "spearman")


#plot - species combined

ggplot(seedxflowersize, aes(x = average_flower_size, y = seed_number)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  theme_minimal()


#Run separate model for each plant species:
# Model for delphinium
delph_model <- lmer(
  seed_number ~ average_flower_size + treatment + (1 | transect:plot) + (1 | transect),
  data = seedxflowersize %>% filter(species == "delphinium")
)

(1|transect/plot)
summary(delph_model)

# Model for flagellaris
flag_model <- lmer(
  seed_number ~ average_flower_size + (1|treatment) + (1 | transect:plot) + (1 | transect),
  data = seedxflowersize %>% filter(species == "flagellaris")
)
summary(flag_model)

#plot

ggplot(seedxflowersize, aes(x = average_flower_size, y = seed_number, color = treatment)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +  # add trend lines per treatment
  facet_wrap(~ species, scales = "free_y") +
  theme_minimal() +
  labs(
    x = "Average Flower Size",
    y = "Seed Number",
    color = "Treatment"
  ) +
  theme(strip.text = element_text(face = "bold"))

#remove treatments: 
ggplot(seedxflowersize, aes(x = average_flower_size, y = seed_number)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +  # add trend lines per treatment
  facet_wrap(~ species, scales = "free_y") +
  theme_minimal() +
  labs(
    x = "Average Flower Size",
    y = "Seed Number"
  ) +
  theme(strip.text = element_text(face = "bold"))

################ other models #################
#Model by Plant Species
#does the effect of flower size on seed number vary by species?

#first need to set species column as.factor
seedxflowersize$species <- as.factor(seedxflowersize$species)

#model:
seedxflower.model <- lmer(
  seed_number ~ average_flower_size * species + (1 | transect:plot) + (1 | transect),
  data = seedxflowersize
)
summary(seedxflower.model)

#model includes:
#(Intercept): mean seed number for delphinium when average_flower_size = 0.
#average_flower_size: slope of the relationship between flower size and seed number for delphinium.
#speciesflagellaris: difference in intercept between flagellaris and delphinium.
#average_flower_size:speciesflagellaris: difference in slope between species.
```

#Tag plant sample size -- wrong but idk why???

```{r}
morph_tags_samplesize <- delph.flower.size %>%
  #filter(!is.na(length)) %>%
  distinct(plant, tag) %>% 
  count(plant, name = "n_tags")
print(morph_tags_samplesize)

seed_tags_samplesize <- seedxflowersize  %>%
  #filter(!is.na(seed_number)) %>%
  #distinct(species, tag) %>% 
  count(species, name = "n_tags")

print(seed_tags_samplesize)
```

#----------------

