---
title: "JB-Scent-Net"
author: "Janelle Bohey"
date: "2023-11-28"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="  ", cache=TRUE, fig.width=14, fig.asp=.678)
library(openxlsx)
library(data.table)
library(RColorBrewer)
pal <- brewer.pal(5,"Set2")
#library(heatmaply)
library(bipartite)
data(mosquin1967)
library(vegan)
library(phytools)
library(taxize)
#remotes::install_github("ropensci/brranching")
library(brranching)
library(rotl)
library(dendextend)
library(tidyverse)
library(heatmaply)

```

#load network
```{r data}
#load network
setwd("~/Desktop/Scent-Net")
net.long <- read_sheet("1plhxbVh5IQLNVMazqO4OvfM3X6vbll_Y6BTMJ94bkI4", na=c("","na", "NA")) %>%    
      mutate(pollinator_group= if_else(pollinator=="papilio_gothica", "butterfly", pollinator_group)) %>% #fix papilio_gothica from bee -> butterfly
      mutate(plant= recode(plant,senecio_tall_int = "senecio_integerrimus", agoceris_glauca = "agoseris_glauca", agoceris_aurantiaca="agoseris_aurantiaca", hydrophyllum_fendeleri="hydrophyllum_fendleri"))

net <- net.long %>% select(plant, pollinator, interactions) %>% pivot_wider(names_from = plant, values_from = interactions, values_fn= sum, values_fill= 0) %>%  #pivots data set longer so plants as columns and pollinators as rows, replaces NAs with zeros
    column_to_rownames("pollinator") #pollinators are in first column, makes them into rownames; #deletes the first pollinators column


net <- as.matrix(t(net)) # turns it from dataframe to matrix bc packages like matrix; t transposes (turns dataset sideways - rows -> columns; columns -> rows) 
polls <- colnames(net) #now pollinators are column names
plants<-rownames(net)
snet <- sortweb(net) #sort by row & column totals; sorted rows by rowsums and columns by column sums. Biggest numbers in top left of matrix



```

#load scent 
```{r}
#Go to comm_vols for code to make scent.net
#Loads in vol, filters with bouquet, takes mean of each voc for each species 
scent.net <- read_csv("data/Volatiles_by_species_mean.csv") %>%     column_to_rownames("species") %>% #species are now rownames
    as.matrix() #network analysis likes matrix 

```

#match paul's plants with the plants we have vocs for
```{r}
scentplants<-rownames(scent.net) %>% tolower() %>%  str_replace(" ", "_")

paulsplants <- plants

#plants not in both datasets 
setdiff(scentplants, paulsplants) #gives us bonus plants
setdiff(paulsplants, scentplants) 

```


# Plant phylogeny
```{r pphylo}
#plants.class <- classification(str_replace(plants,"_"," "), db="gbif")
#save(plants.class, file="data/plantsclass.Rdata")
load("data/plantsclass.Rdata")
#phylogeny based on plant class
plants.classtree <- class2tree(plants.class) #taxonomic tree
plot(plants.classtree)
#make phylogeny from Zanne 2014 (giant plant phylogeny)
plants.tree <-phylomatic(taxa=names(plants.class), get = 'POST', storedtree="zanne2014") 
plants.tree$tip.label[order(plants.tree$tip.label)] <- plants
par(mar=c(0,0,0,0))
cp <- cophylo(plants.tree, plants.classtree$phylo)
plot(cp)
```


# Pollinator groups and family
```{r }
polls.count <-count(net.long, pollinator_group,pollinator_family, pollinator) #counting number of rows for each pollinator
polls.class <- tibble(pollinator=polls) %>%  left_join(polls.count) %>% pull(pollinator_group)
polls.family <- tibble(pollinator=polls) %>%  left_join(polls.count) %>% pull(pollinator_family)
polls.order.all <- polls.class

net.grp <- aggregate(.~ polls.class, as.data.frame(t(net)), sum) #lump network by pollinator class; plants columns, pollinators rows
rownames(net.grp) <- net.grp[,1] #make plants rows, pollinators column names
net.grp[,1] <- NULL
net.grp <- as.matrix(t(net.grp)) #transposes matrix

```


# Plant-pollinator network
##Cophylogeny
```{r cophylo}
mnet.lump <- melt(net.lump)
mnet.lump <- mnet.lump[mnet.lump$value>0,]
cophyloplot(plants.tree, polls.classtree$phylo, mnet.lump[,1:2], lwd=sqrt(mnet.lump[,3])/4, show.tip.label=F,  col='steelblue', length.line=0, gap=-20, space=60)
```

##Pollinator groups
```{r pollgroups}
#Tells you plants major pollinator
plants.majorpoll <- setNames(factor(colnames(net.grp)[apply(net.grp, 1, which.max)]), rownames(net.grp))

#sets colors for ALL 5 pollinator groups
pcols.all <- setNames(brewer.pal(5,"Set2"), colnames(net.grp))

#Four colors for each major pollinator group
pcols <- pcols.all[ levels(plants.majorpoll)]

```


##Heatmap
```{r heatmap, fig.asp=1}

heatmaply(net, scale="none", showticklabels = c(F,T), margins= c(0,NA,0,0),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradient(low = "white", high = "#000050", trans="log1p"),
          distfun=vegdist, 
          hclustfun=function(x) hclust(x, method="mcquitty"), 
          row_side_colors = data.frame(P=plants.majorpoll[rowSums(net)>0]),
          row_side_palette=function(n) pcols, 
          col_side_colors=data.frame(PollinatorGroup=polls.class), 
          col_side_palette=function(n) pcols.all)
```

##Matrix plot
```{r matrixplot, fig.asp=.2}
#interaction matrix plot
visweb(net, type="nested", circles=T, boxes=F, circle.max=1.2)
```

##Web plot
```{r webplot}
#interaction web plot
plotweb(sqrt(net), empty=F, text.rot=90, method="cca", col.high=pcols.all[polls.order.all], col.low=pcols[plants.majorpoll], col.interaction=pcols.all[polls.order.all], bor.col.interaction=pcols.all[polls.order.all], bor.col.high=pcols.all[polls.order.all], bor.col.low=pcols[plants.majorpoll])
```

#Network topography
```{r topo}
#computes network statistics (like nestedness)
networklevel(net)
```

# Scent network
##Heatmap
```{r heatmap.s, fig.asp=1}
#get heatmap code from NEXP 
```

##Web plot VOCs-Plants
```{r webplot.s}
#interaction web plot
#plant-voc bipartite
plotweb(sqrt(scent.net), text.rot=90, method="cca", col.high=pal[2], bor.col.high=pal[2], col.low=pcols[plants.majorpoll], bor.col.low=pcols[plants.majorpoll], col.interaction=pal[5], bor.col.interaction=pal[5])
```

#Plant-scent-pollinator network
##NMDS
```{r nmds}
scent.mds <- metaMDS(sqrt(scent.net), autotransform = F, trace=F)
scent.op <- ordiplot(scent.mds, type="n")
points(scent.op, what="species", pch=3, col="grey50")
points(scent.op, what="sites", cex=1.5, pch=19, col=pcols[plants.majorpoll])
text(scent.op, what="sites", cex=0.8, col=pcols[plants.majorpoll])

```

##Phylochemospace
```{r phylochemo, cache=FALSE}
scent.mds.taxa <- scent.mds$points
rownames(scent.mds.taxa) <- sort(plants.tree$tip.label)
rownames(scent.mds.taxa) <- gsub(" ", "_", rownames(scent.mds.taxa), fixed=T)
plants.tree$tip.label <- gsub(" ", "_", plants.tree$tip.label, fixed=T)

#plants.tree$phylo$edge.length <- plants.tree$phylo$edge.length+0.00001

#Do generalists cluster?
plants.numlinks  <- rowSums(decostand(net.lump, "pa"))
tip.pcols <- setNames(colorRampPalette(c("white", "darkblue"))(67)[plants.numlinks+1], 
                      sort(plants.tree$tip.label))
node.pcols<- setNames(c(tip.pcols[plants.tree$tip.label], rep("black",plants.tree$Nnode)),
                      1:(length(plants.tree$tip)+plants.tree$Nnode))

phylomorphospace(plants.tree, scent.mds.taxa, control=list(col.node=node.pcols), node.size=c(0,2), label="off", lwd=1)
text(scent.mds.taxa[,1], scent.mds.taxa[,2], rownames(scent.mds.taxa), cex=0.8, offset=0.5, xpd=T, col=tip.pcols)

#Do plant major pollinators cluster?
tip.pcols <- setNames(pcols[plants.majorpoll], 
                      sort(plants.tree$tip.label))
node.pcols<- setNames(c(tip.pcols[plants.tree$tip.label], rep("black",plants.tree$Nnode)),
                      1:(length(plants.tree$tip)+plants.tree$Nnode))

phylomorphospace(plants.tree, scent.mds.taxa, control=list(col.node=node.pcols), node.size=c(0,2), label="off", lwd=1)
text(scent.mds.taxa[,1], scent.mds.taxa[,2], rownames(scent.mds.taxa), cex=0.8, offset=0.5, xpd=T, col=tip.pcols)

library(vegan3d)
orditree3d(scent.mds.taxa, as.hclust(force.ultrametric(multi2di(plants.tree))), lwd=2, pch=19, col=tip.pcols)

phenogram(plants.tree, setNames(plants.numlinks,sort(plants.tree$tip.label)))
```


```{r phylochemoanimate, eval=FALSE}
png(file="ppm-%04d.png",width=600,height=600,res=120)
par(mar=c(2.1,2.1,1.1,1.1))
project.phylomorphospace(tree=plants.tree, X=scent.mds.taxa, xlab="", ylab="", node.size=c(0,0), lwd=1, direction="to", nsteps=100, fsize=0.6)
dev.off()
system("convert -delay 10 -loop 2 *.png phylochemo.gif")
file.remove(list.files(pattern=".png"))
```


##Phylogeny scent bar chart
```{r barchart}
ccol = sample(rainbow(ncol(scent.net)))
par(mfrow=c(1,2))
par(mar=c(0.8,0,0.8,0))
plot(plants.tree, cex=1, tip.color = tip.pcols)
par(mar=c(0,0,0,0))
barplot(t(scent.net), col=ccol, names.arg=rep("",nrow(scent.net)), horiz=TRUE) 
```

```{r barchart2}
par(mfrow=c(1,2))
par(mar=c(0.8,0,0.8,0))
plot(plants.tree, cex=1, tip.color = tip.pcols)
par(mar=c(0,0,0,0))
barplot(t(decostand(scent.net, method="tot")), col=ccol, names.arg=rep("",nrow(scent.net)), horiz=TRUE) 
```


##Indicator compounds
```{r indicators}
library(indicspecies)
scent.net.cut <- scent.net[rownames(scent.net)%in% paulsplants, ]

plants.majorpoll.cut <- plants.majorpoll[rownames(scent.net.cut)]
  
mp <- multipatt(as.data.frame(scent.net[rownames(scent.net)%in% paulsplants],), plants.majorpoll[rownames(scent.net ], control=how(nperm=999))
summary(mp)
```


```{r paco}
library(paco)

# start the paco procedure
poll_phy <- polls.classtree$phylo
pla_phy <- as.phylo(hclust(vegdist(scent.net), "mcquitty"))
int <- net.lump
H <- cophenetic(polls.classtree$phylo)
P <- cophenetic(as.phylo(hclust(vegdist(scent.net), "mcquitty")))
D <- prepare_paco_data(H, P, net.lump)
D <- add_pcoord(D, correction='cailliez')
# now we are ready for cophylogenetic analysis
#D <- PACo(D, nperm=1000, seed=13, method='quasiswap', symmetric=TRUE)
#save(D, file="D.Rdata")
load("D.Rdata")
# and to investigate the contribution of individual links
res <- residuals_paco(D$proc)

# to visualise this we use the ape function cophyloplot weighted by interaction contribution
# first we must make a list out of the interaction matrix
assoc <- data.frame(pol=rownames(net.lump)[which(net.lump>0, arr.ind=TRUE)[,'row']], pla=colnames(net.lump)[which(net.lump>0, arr.ind=TRUE)[,'col']])
# to weight the interactions we use the cophylogenetic contribution transformed to best show
# the differences graphically
weight <- (res^-2)/50

mnet.lump <- melt(net.lump) #replaces assoc
mnet.lump <- mnet.lump[mnet.lump$value>0,]
#cophylo with interaction strength
cophyloplot(pla_phy, poll_phy, mnet.lump[,1:2], lwd=sqrt(mnet.lump[,3])/4, show.tip.label=F,  col='steelblue', length.line=0, gap=-20, space=60)
#cophylo with residuals
cophyloplot(pla_phy, poll_phy, mnet.lump[,1:2], lwd=weight/8, show.tip.label=F, col='steelblue', length.line=0, gap=-20, space=60)
```


#Do generalist plants have higher diversity of scent compounds?
```{r generalists}
plants.numlinks  <- rowSums(decostand(net.lump, "pa"))
polls.classtree <- compute.brlen(polls.classtree$phylo,1)
library(picante)
plants.pdlinks   <- pd(net.lump, polls.classtree)$PD
plants.numcompounds <- rowSums(decostand(scent.net, "pa"))
plants.sumcompounds <- rowSums(scent.net)
plants.shannoncompounds <- diversity(scent.net, index="shannon")
plants.shannonlinks <- diversity(net.lump, index="shannon")

plot(plants.numcompounds~plants.numlinks)

plot(plants.numcompounds~plants.pdlinks)

plot(plants.shannoncompounds~plants.pdlinks)

plot(plants.shannoncompounds~plants.shannonlinks)

plot(sqrt(plants.sumcompounds)~plants.pdlinks)

plot(plants.shannoncompounds~plants.majorpoll)

plot(plants.numcompounds~plants.majorpoll)

plot(sqrt(plants.sumcompounds)~plants.majorpoll)

qplot(plants.pdlinks, plants.numcompounds) + geom_smooth(method="lm")

summary(lm(plants.numcompounds[-9]~plants.pdlinks[-9]))

data(phylocom)
pd(phylocom$sample, phylocom$phylo)

```
