---
title: "JB-Scent-Net"
author: "Janelle Bohey"
date: "2023-11-28"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="  ", cache=TRUE, fig.width=14, fig.asp=.678)
library(openxlsx)
library(data.table)
library(RColorBrewer)
pal <- brewer.pal(5,"Set2")
#library(heatmaply)
library(bipartite)
data(mosquin1967)
library(vegan)
library(phytools)
library(taxize)
#remotes::install_github("ropensci/brranching")
library(brranching)
library(rotl)
library(dendextend)
library(tidyverse)
library(heatmaply)
library(googlesheets4)

```

#load network
```{r data}
#load network
setwd("~/Desktop/Scent-Net")
gs4_auth(email=TRUE)
net.long <- read_sheet("1plhxbVh5IQLNVMazqO4OvfM3X6vbll_Y6BTMJ94bkI4", na=c("","na", "NA")) %>%    
      mutate(pollinator_group= if_else(pollinator=="papilio_gothica", "butterfly", pollinator_group)) %>% #fix papilio_gothica from bee -> butterfly
      mutate(plant= recode(plant,senecio_tall_int = "senecio_integerrimus", agoceris_glauca = "agoseris_glauca", agoceris_aurantiaca="agoseris_aurantiaca", hydrophyllum_fendeleri="hydrophyllum_fendleri"))

net <- net.long %>% select(plant, pollinator, interactions) %>% pivot_wider(names_from = plant, values_from = interactions, values_fn= sum, values_fill= 0) %>%  #pivots data set longer so plants as columns and pollinators as rows, replaces NAs with zeros
    column_to_rownames("pollinator") #pollinators are in first column, makes them into rownames; #deletes the first pollinators column


net <- as.matrix(t(net)) # turns it from dataframe to matrix bc packages like matrix; t transposes (turns dataset sideways - rows -> columns; columns -> rows) 
polls <- colnames(net) #now pollinators are column names
plants<-rownames(net)
snet <- sortweb(net) #sort by row & column totals; sorted rows by rowsums and columns by column sums. Biggest numbers in top left of matrix
```

#load scent 
```{r}

#Go to comm_vols for code to make scent.net
#Loads in vol, filters with bouquet, takes mean of each voc for each species 
#From comm_vols.Rmd
scent.net <- read_csv("data/Volatiles_by_species_mean.csv") %>% column_to_rownames("species") %>% #species are now rownames
    as.matrix() #network analysis likes matrix 

#loads in chemical class and gets rid of ?
compound.class <-read_sheet("1BlDIOs4STe5ZTTPYE6qez1Nsey_ovEpe5m-SN8Q55gg", sheet= "chemical_class") %>% mutate(major_class= factor(major_class,levels=c("aliphatic", "monoterpene", "sesquiterpene", "nitrogenous", "sulfur", "benzenoid")) %>% fct_na_value_to_level("other")) %>% filter(compound_long!="Cyclohexane, 1,3,5-trimethyl-2-octadecyl-") #filtering this compound bc it has 0 interactions or emissions... 355 vs 354

```

##match paul's plants with the plants we have vocs for
```{r}
scentplants<-rownames(scent.net) #%>% tolower() %>%  str_replace(" ", "_")

paulsplants <- plants

#plants not in both datasets 
setdiff(scentplants, paulsplants) #gives us bonus plants
setdiff(paulsplants, scentplants) 

bothplants <- intersect(scentplants, paulsplants) #plants found in amanda and paul data
```


# Pollinator groups and family
```{r }
polls.count <-count(net.long, pollinator_group,pollinator_family, pollinator) #counting number of rows for each pollinator
polls.class <- tibble(pollinator=polls) %>%  left_join(polls.count) %>% pull(pollinator_group)
polls.class.dataframe <- tibble(pollinator=polls) %>%  left_join(polls.count)
polls.family <- tibble(pollinator=polls) %>%  left_join(polls.count) %>% pull(pollinator_family)
polls.order.all <- polls.class

#net.grp.all has all of pauls 45 plant species 
#net.grp only has plant species found in Amanda's and paul's datasets
net.grp.all <- aggregate(.~ polls.class, as.data.frame(t(net)), sum) #lump network by pollinator class; plants columns, pollinators rows
rownames(net.grp.all) <- net.grp.all[,1] #make plants rows, pollinators column names
net.grp.all[,1] <- NULL
net.grp.all <- as.matrix(t(net.grp.all)) #transposes matrix
net.grp<- net.grp.all[bothplants,] 

net.cut <- net[bothplants,]

```


# Plant-pollinator network

##Pollinator groups
```{r pollgroups}
#Tells you plants major pollinator
plants.majorpoll <- setNames(factor(colnames(net.grp)[apply(net.grp, 1, which.max)]), rownames(net.grp))

#sets colors for ALL 5 pollinator groups
pcols.all <- setNames(brewer.pal(5,"Set2"), colnames(net.grp))

#Four colors for each major pollinator group
pcols <- pcols.all[ levels(plants.majorpoll)]

```

# Proportion of visits by pollinator group
```{r}
#gives % bee pollination for each plant species 
plants.bees <- net.grp[,"bee"]/rowSums(net.grp)
plot(sort(plants.bees))

# Cap of plants.bees
plant.bee.cap <- capscale(sqrt(scent.net.cut)~plants.bees)
anova(plant.bee.cap)
#p=0.348
plot(plant.bee.cap, type="text")

#gives % fly pollination for each plant species 
plants.fly <- net.grp[,"fly"]/rowSums(net.grp)
plot(sort(plants.fly))
sort(plants.fly)
```

```{r}
VOClist <- 
```


##Heatmap
```{r heatmap, fig.asp=1}

#pp_heatmap
heatmaply(net, scale="none", showticklabels = c(F,T), margins= c(0,NA,0,0),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradient(low = "white", high = "#000050", trans="log1p"),
          distfun=vegdist, 
          hclustfun=function(x) hclust(x, method="mcquitty"), 
          row_side_colors = data.frame(P=plants.majorpoll[rowSums(net)>0]),
          row_side_palette=function(n) pcols, 
          col_side_colors=data.frame(PollinatorGroup=polls.class), 
          col_side_palette=function(n) pcols.all)
```

##Matrix plot
```{r matrixplot, fig.asp=.2}
#interaction matrix plot
visweb(net, type="nested", circles=T, boxes=F, circle.max=1.2)
```

##Web plot
```{r webplot, fig.height=15, fig.width=18}

#fix names of species (plants rows, polls cols)
net.nice.labels <- net
rownames(net.nice.labels) <- str_to_sentence(str_replace_all(rownames(net), "_"," "))
colnames(net.nice.labels) <- str_to_sentence(str_replace_all(colnames(net), "_"," "))

#interaction web plot
par(mar=c(0,0,0,0))
plant.pollinator.network <- plotweb(sqrt(net.nice.labels), empty=F, text.rot=90, method="cca", col.high=pcols.all[polls.order.all], col.low=pcols[plants.majorpoll], col.interaction=pcols.all[polls.order.all], bor.col.interaction=pcols.all[polls.order.all], bor.col.high=pcols.all[polls.order.all], bor.col.low=pcols[plants.majorpoll])

library(ggplot2)
ggsave(plant.pollinator.network, file='plant.pollinator.networkplot.jpg', width=130, height=105, units= 'mm', dpi=300)

```

#Network topography
```{r topo}
#computes network statistics (like nestedness)
#networkstats <- networklevel(net)
#write_csv(enframe(networkstats), "data/p-p network stats.csv")

#computes network statistics for voc-pollinator network
#scentnetworkstats<- networklevel(scentpollnet)
#write_csv(enframe(scentnetworkstats), "data/scent network stats.csv")

#the network statistics null model: TODO- broken
#nullmmodel<-nullmodel(scentpollnet)
#nullmodel.stats <- networklevel(nullmodel)
#write_csv(enframe(nullmodel.stats), "data/scent network null model.csv")
```

# Scent network
##Heatmap
```{r heatmap.s, fig.asp=1}
scent.net.cut <- scent.net[rownames(scent.net)%in% paulsplants, ] #cuts scent to match what paul has pollinator data for
plants.majorpoll.cut <- plants.majorpoll[rownames(scent.net.cut)] #cuts Pauls data to match Amanda's

plants.bees.cut<- plants.bees[rownames(scent.net.cut)]

heatmaply(sqrt(scent.net.cut), scale="none", showticklabels = c(F,T), margins= c(0,NA,0,0),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradient(low = "white", high = "#000050", trans="log1p"),
          distfun=vegdist, 
          hclustfun=function(x) hclust(x, method="mcquitty"), 
          row_side_colors = data.frame(major.polls=plants.majorpoll.cut),
          row_side_palette=function(n) pcols )
          #col_side_colors=data.frame(PollinatorGroup=scent.net), #put compound class here
        #  col_side_palette=function(n) pcols.all)

```

##Web plot VOCs-Plants
```{r webplot.s}
#interaction web plot
#plant-voc bipartite
plotweb(sqrt(scent.net), text.rot=90, method="cca", col.high=pal[2], bor.col.high=pal[2], col.low=pcols[plants.majorpoll], bor.col.low=pcols[plants.majorpoll], col.interaction=pal[5], bor.col.interaction=pal[5])


```


#Plant-scent-pollinator network
##NMDS
```{r nmds}
scent.net.cut <- scent.net[rownames(scent.net)%in% paulsplants, ]
scent.mds <- metaMDS(sqrt(scent.net.cut), autotransform = F, trace=F)
scent.op <- ordiplot(scent.mds, type="n")
points(scent.op, what="species", pch=3, col="grey50")
points(scent.op, what="sites", cex=1.5, pch=19, col=pcols[plants.majorpoll])
text(scent.op, what="sites", cex=0.8, col=pcols[plants.majorpoll])

#are scents of plants with different major pollinators different? no... 
#do phylogeny, plants in same genus will cluster together need to account for phylogeny
```

#cap
```{r}
scent.cap <- capscale(sqrt(scent.net.cut)~plants.majorpoll.cut)
anova(scent.cap)

plot(scent.cap, type="text")
```


##Indicator compounds
```{r indicators}
library(indicspecies)

plants.majorpoll.cut <- plants.majorpoll[rownames(scent.net.cut)]
  
mp <- multipatt(as.data.frame(scent.net.cut), plants.majorpoll.cut, control=how(nperm=999))
summary(mp)
```


#Do generalist plants have higher diversity of scent compounds?
```{r generalists}
plants.numlinks  <- rowSums(decostand(net.cut, "pa"))
#polls.classtree <- compute.brlen(polls.classtree$phylo,1)
#library(picante)
#plants.pdlinks   <- pd(net.lump, polls.classtree)$PD
plants.numcompounds <- rowSums(decostand(scent.net.cut, "pa"))
plants.sumcompounds <- rowSums(scent.net.cut)
plants.shannoncompounds <- diversity(scent.net.cut, index="shannon")
plants.shannonlinks <- diversity(net.grp, index="shannon")

scent.diversity <- tibble(numlinks=plants.numlinks, numcompounds=plants.numcompounds, sumcompounds=plants.sumcompounds, shannoncompounds=plants.shannoncompounds, shannonlinks=plants.shannonlinks)

ggplot(scent.diversity, aes(y=numcompounds, x=numlinks))+
    geom_point()+ geom_smooth()

lm(numcompounds ~ numlinks, data=scent.diversity)
anova(lm(numcompounds ~ numlinks, data=scent.diversity))

#plot(plants.numcompounds~plants.pdlinks)

#plot(plants.shannoncompounds~plants.pdlinks)

ggplot(scent.diversity, aes(y=shannoncompounds, x=shannonlinks))+
    geom_point()+ geom_smooth()


ggplot(scent.diversity, aes(y=shannoncompounds, x=plants.majorpoll.cut))+
    geom_point()+ geom_smooth()

ggplot(scent.diversity, aes(y=numcompounds, x=plants.majorpoll.cut))+
    geom_point()+ geom_smooth()

ggplot(scent.diversity, aes(y=sumcompounds, x=plants.majorpoll.cut))+
    geom_point()+ geom_smooth()


```

#VOC-pollinator network
```{r}
#interaction web plot
#voc-pollinator bipartite

#make dataframe with vocs and pollinators
##change scent.net to long format
plantcompoundemission1 <- scent.net.cut %>% as.data.frame() %>% rownames_to_column("plant") %>% #average peak area
  pivot_longer(-plant, names_to = "compound", values_to = "emissions") %>% filter(emissions!=0)

#write_csv(plantcompoundemission, "plantcompoundemissions.csv")
 
#total number of visits to plant by each pollinator species
plantpollinatorvisits <- net.cut %>% as.data.frame() %>% rownames_to_column("plant") %>% pivot_longer(-plant, names_to = "pollinator", values_to = "visits")

#TODO: 
#total number of visits to each compound. Pollinators rows, compounds cols, values = number of visits
scentpollnet <- left_join(plantcompoundemission, plantpollinatorvisits,relationship = "many-to-many") %>%  group_by(pollinator, compound) %>% summarise(visits=sum(visits)) %>%  pivot_wider(names_from = "compound", values_from = "visits") %>% column_to_rownames("pollinator") %>% as.matrix()

#relative amount of each compound
plantcompoundrelative <- scent.net.cut %>% as.data.frame() %>% decostand(method="tot") %>% rownames_to_column("plant") %>% #average peak area
  pivot_longer(-plant, names_to = "compound", values_to = "emissions") %>% filter(emissions!=0)

#multiply number of visits by relative peak area
# weights visit to compounds. Proportions visits based on how many VOCs there are.
scentpollnet.area <- left_join(plantcompoundrelative, plantpollinatorvisits,relationship = "many-to-many") %>%  group_by(pollinator, compound) %>% mutate(visits.area=visits*emissions) %>%  summarise(visits.area=sum(visits.area)) %>%  pivot_wider(names_from = "compound", values_from = "visits.area") %>% column_to_rownames("pollinator") %>% as.matrix()

scentpollnet.class <- tibble(pollinator=rownames(scentpollnet)) %>% left_join(polls.class.dataframe) %>% pull(pollinator_group)


#plot bipartite network
plotweb(sqrt(scentpollnet), text.rot=90, method="cca", col.high=pal[2], bor.col.high=pal[2], col.low=pcols[scentpollnet.class], bor.col.low=pcols[scentpollnet.class], col.interaction=pal[5], bor.col.interaction=pal[5])
```

##Heatmap scent-net
```{r heatmap, fig.asp=1}

#value = number of 
heatmaply(scentpollnet, scale="none", showticklabels = c(F,T), margins= c(0,NA,0,0),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradient(low = "white", high = "#000050", trans="log1p"),
          distfun=vegdist, 
          hclustfun=function(x) hclust(x, method="mcquitty"), 
          #row_side_colors = data.frame(P=plants.majorpoll[rowSums(net)>0]),
          #row_side_palette=function(n) pcols, 
          row_side_colors=data.frame(PollinatorGroup=scentpollnet.class), 
          row_side_palette=function(n) pcols.all)

voc.pollgrp.cap <- capscale(sqrt(scentpollnet.area)~scentpollnet.class)
anova(voc.pollgrp.cap)

plot(voc.pollgrp.cap, type="text")

```

#relative peak area x visits
```{r}

#value = number of 
heatmaply(scentpollnet.area, scale="none", showticklabels = c(F,T), margins= c(0,NA,0,0),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradient(low = "white", high = "#000050", trans="log1p"),
          distfun=vegdist, 
          hclustfun=function(x) hclust(x, method="mcquitty"), 
          #row_side_colors = data.frame(P=plants.majorpoll[rowSums(net)>0]),
          #row_side_palette=function(n) pcols, 
          row_side_colors=data.frame(PollinatorGroup=scentpollnet.class), 
          row_side_palette=function(n) pcols.all)

```


#Rarifaction
```{r}
#species accumulation curves
#figure out what percent Pauls data catpured the whole pollinator community
#pull out 1 random 1 hr sampling period and say how many pollinators observered, then do this again and again randomly - variation in samples = standard error y axis= number of species, x= number of pollinator observations 

#class= bee, fly, butterfly, hummingbird
#xaxis is number of plant species observed (if you did the whole summer but only looked at 1 species and then next summer looked at 2 species, etc.)
#

#simplify net.long to only have numbers in it so that accumulation curve will work 
#rows are observation "units" and columns are pollinators
net.wide <- net.long %>% select(pollinator, interactions) %>% mutate(index=row_number()) %>% 
    pivot_wider(names_from=pollinator, values_from = interactions) %>% mutate(across(everything(),~replace_na(.x, 0))) %>% select(-index) %>% as.matrix()

for(class in unique(polls.class)) { #could replace class with family
  class_subset <- net[,polls.class==class] #making accumulation curve for each pollinator class
  class_sa <- vegan::specaccum(class_subset, permutations = 999) 
  plot(class_sa, col = as.integer(factor(class))+1, add = class != "bee", lwd=2, 
       xlab="Pollinator Observations", ylab="Pollinator Species", main="Pollinator Species accumulation curve", sub=class)
}
legend("bottomright", unique(polls.class), 
       col = 2:5, pch = 19, title = "Class")

sa<- vegan::specaccum(net.wide,permutations=100, method="exact")
plot(sa) 

#Pauls total pollinator observations 8hrs/wk * 42wks = 336 hours total

visitationrate <- nrow(net.long)/336 #15.33 unique p-p observations per hr (observation can include multiple visits)
#observation=within a 15min period how many p-p interactions did we see, total number of links observed
2000/visitationrate # 130 hrs per treatment ( 2 treatments = 260 hrs per year)

#then based on my observation hours, what percent of the pollinator community could be captured


```

#modularity
```{r}



```

#normalized degree for p-p
```{r}


library(rstatix)
ND_poll<-ND(net, normalised = T)$higher 

#creates dataframe with normalized degree for each pollinator and pollinator class (same as data.frame)
ND.poll.class <- tibble(ND=ND_poll, polls.class=polls.class) 
  
kruskal_test(ND.poll.class, ND~polls.class)

ggplot(ND.poll.class, aes(x= polls.class, y=ND)) +geom_boxplot()



```

#normalized degree for VOC-pollinator
```{r}
library(rstatix)

#ND for pollinators in scent network
ND_poll_scent<-ND(scentpollnet, normalised = T)$lower 

#creates dataframe with normalized degree for each pollinator and pollinator class (same as data.frame)
ND.poll <- tibble(ND=ND_poll_scent, polls.class=polls.class) 
  
kruskal_test(ND.poll, ND~polls.class)

ggplot(ND.poll, aes(x= polls.class, y=ND)) +geom_boxplot()


#ND of VOC chemical class
ND_scent_poll<-ND(scentpollnet, normalised = T)$higher 

#creates dataframe with normalized degree for each VOC and chemical class (same as data.frame)
ND.scent <- tibble(ND=ND_scent_poll, vols.class=compound.class$major_class) 
  
kruskal_test(ND.scent, ND~vols.class)

ggplot(ND.scent, aes(x= vols.class, y=ND)) +geom_boxplot()

```

#normalized visits - home cooked
```{r}
#created a function similar to normalized degree(ND) but ND looks at presence/absence in the network where normalized visits (NV) looks at the normalized weighted visits/interactions. This accounts for variation in the amount each plant or compound is visited and give those with higher interaction frequency greater weight or contribution to the network  

normalized.visits <- function (web, normalised = TRUE) 
{

  websum <- sum(web) #takes sum of all interactions in the web
  dlower <- rowSums(web)
  dhigher <- colSums(web)
  Nlow <- Nhigh <- 2
  if (normalised) {
    Nlow <- websum 
    Nhigh <- websum
  }
  low <- dlower/Nlow #weighting the interactions by dividing rowsums by the sum of all links in the network
  names(low) <- rownames(web) #dividing colsums by the sum of all links in the network
  high <- dhigher/Nhigh
  names(high) <- colnames(web)
  list(lower = low, higher = high)
}
```


#weighted normalized  for p-p (pollinators only)- home cooked
```{r}
#Which pollinators have the most visits. of all the visits recorded, what percentage of them were by that pollinator (for a given plant what percentage of visits does it get out of the whole meadow, for a pollinator what percentage of visits does it give out of all visits)

#normalized visits for pollinators in scent net
NV_poll<-normalized.visits(net, normalised = T)$higher 

#creates dataframe with normalized degree for each pollinator and pollinator class (same as data.frame)
NV.poll.class <- tibble(NV=NV_poll, polls.class=polls.class) 
  
kruskal_test(NV.poll.class, NV~polls.class)

ggplot(NV.poll.class, aes(x= polls.class, y=NV)) +geom_boxplot()
#only a few species making most of the visits
#1 bee responsible for 40% of visits


```

#normalized visits for scent network
```{r}

#normalized visits for pollinators in scent net
#which pollinator class visits the most amount of compounds weighted by the number of visits
NV_poll_scentnet<-normalized.visits(scentpollnet, normalised = T)$lower 

#creates dataframe with normalized degree for each pollinator and pollinator class (same as data.frame)
NV.poll.class.scentnet <- tibble(NV=NV_poll_scentnet, polls.class=polls.class) 
  
kruskal_test(NV.poll.class.scentnet, NV~polls.class)

ggplot(NV.poll.class.scentnet, aes(x= polls.class, y=NV)) +geom_boxplot()
#super generalist bees

#normalized visits for VOCs in scent net
#which compound class are most visited
NV_voc<-normalized.visits(scentpollnet, normalised = T)$higher 

#creates dataframe with normalized degree for each pollinator and pollinator class (same as data.frame)
NV.vols.class <- tibble(NV=NV_voc, vols.class=compound.class$major_class) 
  
kruskal_test(NV.vols.class, NV~vols.class)

ggplot(NV.vols.class, aes(x= vols.class, y=NV)) +geom_boxplot()
#tells you what vocs get the most visits
#no difference in visitation rate to compound classes

```

