---
title: "Plant-scent-pollinator network"
author: '[John Powers](http://sites.uci.edu/powers/)'
date: '`r Sys.Date()`'
output:
  html_document:
    toc: yes
    toc_float: TRUE
---
<style type="text/css">
.main-container { max-width: 1600px; margin-left: 0; margin-right: auto; }
img{ max-width:200%; height: auto; }
td, th { padding : 6px }
</style>

#Load network and scent data
From Tables S1 and S2 in:
Kantsa, A., R.A. Raguso, A.G. Dyer, J.M. Olesen, T. Tscheulin, and T. Petanidou. 2018. Disentangling the role of floral sensory stimuli in pollination networks. Nature Communications 9: 1041. doi:[10.1038/s41467-018-03448-w](https://doi.org/10.1038/s41467-018-03448-w)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="  ", cache=TRUE, fig.width=14, fig.asp=.678)
library(openxlsx)
library(data.table)
library(RColorBrewer)
pal <- brewer.pal(5,"Set2")
library(heatmaply)
library(bipartite)
data(mosquin1967)
library(vegan)
library(phytools)
library(taxize)
library(brranching)
library(rotl)
```

```{r data}
#load network
net <- read.delim("Kantsa_etal_Supp_Data1.csv", skip = 2)
rownames(net) <- net[,1]
net[,1] <- NULL
net <- as.matrix(t(net))
polls <- colnames(net)
snet <- sortweb(net) #sort by row & column totals

#load floral scents
scent.file <- "Kantsa_2018_Supp_Data2.xlsx"
plants <- getSheetNames(scent.file)
scentlist <- lapply(2:length(plants), read.xlsx, xlsxFile=scent.file)
plants <- plants[-1]
scent <- rbindlist(scentlist, fill=T, idcol="plant")[,-4]
colnames(scent)[2:3] <- c("cmpd","mean")
scent$plant <- factor(plants[scent$plant])
scent$cmpd <- factor(tolower(scent$cmpd))
cmpds <- levels(scent$cmpd)

scent.net <- dcast(scent, plant ~ cmpd, value.var="mean", fun.aggregate=mean, fill=0) #long to wide
scent.net <- as.data.frame(scent.net)
rownames(scent.net) <- scent.net[,1]
scent.net[,c(1,380)] <- NULL #weird NA column has one entry
scent.net <- as.matrix(scent.net)
```


# Plant phylogeny
```{r pphylo}
#plants.class <- classification(plants, db="gbif")
#save(plants.class, file="plantsclass.Rdata")
load("plantsclass.Rdata")
plants.classtree <- class2tree(plants.class) #taxonomic tree
plants.tree <-phylomatic(taxa=names(plants.class), get = 'POST', storedtree="zanne2014")
plants.tree$tip.label[order(plants.tree$tip.label)] <- plants
par(mar=c(0,0,0,0))
cp <- cophylo(plants.tree, plants.classtree$phylo)
plot(cp)
```

# Pollinator phylogeny
```{r pollphylo}
polls <- gsub("Bug", "Hemiptera", polls, fixed=T)
polls <- gsub("Auchenorrhyncha", "Hemiptera", polls, fixed=T)
polls <- gsub("Fly", "Diptera", polls, fixed=T)
polls <- gsub("Beetle", "Coleoptera", polls, fixed=T)
#polls.class <- classification(polls, db="gbif")
#save(polls.class, file="pollsclass.Rdata")
load("pollsclass.Rdata")
polls.gbif <- factor(sapply(polls.class, function(x) tail(x$name, 1)))
net.lump <- aggregate(.~ polls.gbif, as.data.frame(t(net)), sum)
rownames(net.lump) <- net.lump[,1]
net.lump[,1] <- NULL
net.lump <- as.matrix(t(net.lump))
rownames(net.lump) <- sort(plants.tree$tip.label)
polls.classtree <- class2tree(unique(polls.class)) #taxonomic tree
polls.matched <- tnrs_match_names(lapply(polls.class, function(x) tail(x$name, 1)), context_name = "Animals")
polls.ids <- ott_id(polls.matched)
polls.ids <- polls.ids[!polls.ids %in% c(968359, 707236, 707875, 63881, 458953, 42699, 742352, 340559, 672497, 34898, 332969,356987)]
polls.tree <- tol_induced_subtree(ott_ids=polls.ids)
par(mar=c(0,0,0,0))
plot(polls.classtree, cex=0.5)
plot(polls.tree, cex=0.5)
```


# Plant-pollinator network
##Cophylogeny
```{r cophylo}
mnet.lump <- melt(net.lump)
mnet.lump <- mnet.lump[mnet.lump$value>0,]
cophyloplot(plants.tree, polls.classtree$phylo, mnet.lump[,1:2], lwd=sqrt(mnet.lump[,3])/4, show.tip.label=F,  col='steelblue', length.line=0, gap=-20, space=60)
```

##Pollinator groups
```{r pollgroups}
polls.cut <- dendextend::cutree(polls.classtree$phylo,5)
polls.order <- setNames(factor(sapply(unique(polls.class), function(x) x$name[4])), names(polls.class[!duplicated(polls.class)]))
polls.order.all <- setNames(factor(sapply(polls.class, function(x) x$name[4])),names(polls.class))
#table(polls.cut, polls.order)

net.grp <- aggregate(.~ polls.order.all, as.data.frame(t(net)), sum)
rownames(net.grp) <- net.grp[,1]
net.grp[,1] <- NULL
net.grp <- as.matrix(t(net.grp))
rownames(net.grp) <- sort(plants.tree$tip.label)

plants.majorpoll <- setNames(factor(colnames(net.grp)[apply(net.grp, 1, which.max)]), rownames(net.grp))

pcols <- setNames(brewer.pal(4,"Set1"), levels(plants.majorpoll))
plants.polltrees<-make.simmap(plants.tree,plants.majorpoll,model="ARD",nsim=100)
plants.polltree<-summary(plants.polltrees,plot=FALSE)
plot(plants.polltree,colors=pcols,fsize=0.8,cex=c(0,0.3))
add.simmap.legend(colors=pcols,x=0.9*par()$usr[1],
    y=0.2*par()$usr[4],prompt=FALSE,fsize=0.9)

```


##Heatmap
```{r heatmap, fig.asp=1}
heatmaply(net, scale="column", row_side_colors = data.frame(plants.majorpoll), row_side_palette=function(n) pcols)
```

##Matrix plot
```{r matrixplot, fig.asp=.2}
#interaction matrix plot
visweb(net, type="nested", circles=T, boxes=F, circle.max=1.2)
```

##Web plot
```{r webplot}
#interaction web plot
pcols.all <- setNames(c(pcols[1:2],brewer.pal(5,"Set1")[5],pcols[3:4]), levels(polls.order.all))
plotweb(sqrt(net), empty=F, text.rot=90, method="cca", col.high=pcols.all[polls.order.all], col.low=pcols[plants.majorpoll], col.interaction=pcols.all[polls.order.all], bor.col.interaction=pcols.all[polls.order.all], bor.col.high=pcols.all[polls.order.all], bor.col.low=pcols[plants.majorpoll])
```

#Network topography
```{r topo}
networklevel(net)
```

# Scent network
##Heatmap
```{r heatmap.s, fig.asp=1}
heatmaply(scent.net, scale="column", row_side_colors = data.frame(plants.majorpoll), row_side_palette=function(n) pcols)
```

##Web plot
```{r webplot.s}
#interaction web plot
plotweb(sqrt(scent.net), text.rot=90, method="cca", col.high=pal[2], bor.col.high=pal[2], col.low=pcols[plants.majorpoll], bor.col.low=pcols[plants.majorpoll], col.interaction=pal[5], bor.col.interaction=pal[5])
```

##NMDS
```{r nmds}
scent.mds <- metaMDS(sqrt(scent.net), autotransform = F, trace=F)
#scent.op <- ordiplot(scent.mds, type="n")
#points(scent.op, what="species", pch=3, col="grey50")
#points(scent.op, what="sites", cex=1.5, pch=19, col=pcols[plants.majorpoll])
#text(scent.op, what="sites", cex=0.8, col=pcols[plants.majorpoll])
```

##Phylochemospace
```{r phylochemo, cache=FALSE}
scent.mds.taxa <- scent.mds$points
rownames(scent.mds.taxa) <- sort(plants.tree$tip.label)
rownames(scent.mds.taxa) <- gsub(" ", "_", rownames(scent.mds.taxa), fixed=T)
plants.tree$tip.label <- gsub(" ", "_", plants.tree$tip.label, fixed=T)

#plants.tree$phylo$edge.length <- plants.tree$phylo$edge.length+0.00001

tip.pcols <- setNames(pcols[plants.majorpoll], 
                      sort(plants.tree$tip.label))
node.pcols<- setNames(c(tip.pcols[plants.tree$tip.label], rep("black",plants.tree$Nnode)),
                      1:(length(plants.tree$tip)+plants.tree$Nnode))

phylomorphospace(plants.tree, scent.mds.taxa, control=list(col.node=node.pcols), node.size=c(0,1), label="off", lwd=1)
text(scent.mds.taxa[,1], scent.mds.taxa[,2], rownames(scent.mds.taxa), cex=0.8, offset=0.5, xpd=T, col=tip.pcols)

library(vegan3d)
orditree3d(scent.mds.taxa, as.hclust(force.ultrametric(multi2di(plants.tree))), lwd=2, pch=19, col=tip.pcols)
```


```{r phylochemoanimate, eval=FALSE}
png(file="ppm-%04d.png",width=600,height=600,res=120)
par(mar=c(2.1,2.1,1.1,1.1))
project.phylomorphospace(tree=plants.tree, X=scent.mds.taxa, xlab="", ylab="", node.size=c(0,0), lwd=1, direction="to", nsteps=100, fsize=0.6)
dev.off()
system("convert -delay 10 -loop 2 *.png phylochemo.gif")
file.remove(list.files(pattern=".png"))
```


##Phylo Bar chart
```{r barchart}
ccol = sample(rainbow(ncol(scent.net)))
par(mfrow=c(1,2))
par(mar=c(0.8,0,0.8,0))
plot(plants.tree, cex=1, tip.color = tip.pcols)
par(mar=c(0,0,0,0))
barplot(t(scent.net), col=ccol, names.arg=rep("",nrow(scent.net)), horiz=TRUE) 
```

```{r barchart2}
par(mfrow=c(1,2))
par(mar=c(0.8,0,0.8,0))
plot(plants.tree, cex=1, tip.color = tip.pcols)
par(mar=c(0,0,0,0))
barplot(t(decostand(scent.net, method="tot")), col=ccol, names.arg=rep("",nrow(scent.net)), horiz=TRUE) 
```

