---
title: "Plant-scent-pollinator network"
author: '[John Powers](http://sites.uci.edu/powers/)'
date: '`r Sys.Date()`'
output:
  html_document:
    toc: yes
    toc_float: TRUE
---
<style type="text/css">
.main-container { max-width: 1600px; margin-left: 0; margin-right: auto; }
img{ max-width:200%; height: auto; }
td, th { padding : 6px }
</style>

#Load network and scent data
From Tables S1 and S2 in:
Kantsa, A., R.A. Raguso, A.G. Dyer, J.M. Olesen, T. Tscheulin, and T. Petanidou. 2018. Disentangling the role of floral sensory stimuli in pollination networks. Nature Communications 9: 1041. doi:[10.1038/s41467-018-03448-w](https://doi.org/10.1038/s41467-018-03448-w)

```{r setup}
knitr::opts_chunk$set(echo = TRUE, comment="  ", cache=TRUE, fig.width=14, fig.asp=.678)
library(openxlsx)
library(data.table)
library(RColorBrewer)
pal <- brewer.pal(5,"Set2")
library(heatmaply)
library(bipartite)
data(mosquin1967)
library(vegan)
library(phytools)
library(taxize)

#load network
net <- read.delim("Kantsa_etal_Supp_Data1.csv", skip = 2)
rownames(net) <- net[,1]
net[,1] <- NULL
net <- as.matrix(t(net))
snet <- sortweb(net) #sort by row & column totals

#load floral scents
scent.file <- "Kantsa_2018_Supp_Data2.xlsx"
plants <- getSheetNames(scent.file)
scentlist <- lapply(2:length(plants), read.xlsx, xlsxFile=scent.file)
plants <- plants[-1]
scent <- rbindlist(scentlist, fill=T, idcol="plant")[,-4]
colnames(scent)[2:3] <- c("cmpd","mean")
scent$plant <- factor(plants[scent$plant])
scent$cmpd <- factor(tolower(scent$cmpd))
cmpds <- levels(scent$cmpd)

scent.net <- dcast(scent, plant ~ cmpd, value.var="mean", fun.aggregate=mean, fill=0) #long to wide
scent.net <- as.data.frame(scent.net)
rownames(scent.net) <- scent.net[,1]
scent.net[,c(1,380)] <- NULL #weird NA column has one entry
scent.net <- as.matrix(scent.net)
```

# Plant phylogeny
```{r phylo}
#plants.class <- classification(plants, db="gbif")
#save(plants.class, file="plantsclass.Rdata")
load("plantsclass.Rdata")
plants.tree <- class2tree(plants.class)
plot(plants.tree)
```


# Plant-pollinator network
##Heatmap
```{r heatmap, fig.asp=1}
heatmaply(net, scale="column")
```

##Matrix plot
```{r matrixplot, fig.asp=.2}
#interaction matrix plot
visweb(net, type="nested", circles=T, boxes=F, circle.max=1.2)
```

##Web plot
```{r webplot}
#interaction web plot
plotweb(net, text.rot=90, method="cca", col.high=pal[2], col.low=pal[1], col.interaction=pal[5])
```

#Network topography
```{r topo}
networklevel(net)
```

# Scent network
##Heatmap
```{r heatmap.s, fig.asp=1}
heatmaply(scent.net, scale="column")
```

##Web plot
```{r webplot.s}
#interaction web plot
plotweb(scent.net, text.rot=90, method="cca", col.high=pal[2], col.low=pal[1], col.interaction=pal[5])
```

##NMDS
```{r nmds}
scent.mds <- metaMDS(sqrt(scent.net), autotransform = F)
scent.op <- ordiplot(scent.mds, type="n")
points(scent.op, what="species", pch=3, col=pal[3])
text(scent.op, what="sites", cex=0.8)
```

##Phylochemospace
```{r phylochemo}
scent.mds.taxa <- scent.mds$points
rownames(scent.mds.taxa) <- plants.tree$phylo$tip.label

plants.tree$phylo$edge.length <- plants.tree$phylo$edge.length+0.00001

phylomorphospace(plants.tree$phylo, scent.mds.taxa, control=list(col.node=0), label="none", lwd=1)
text(scent.mds.taxa[,1], scent.mds.taxa[,2], rownames(scent.mds.taxa), cex=0.8, offset=0.5, xpd=T)
```


##Phylo Bar chart
```{r barchart}
par(mfrow=c(1,2))
par(mar=c(0.8,0,0.8,0))
plot(plants.tree, cex=1)
par(mar=c(0,0,0,0))
barplot(t(scent.net), col=sample(rainbow(ncol(scent.net))), names.arg=rep("",nrow(scent.net)), horiz=TRUE) 
```

```{r barchart2}
par(mfrow=c(1,2))
par(mar=c(0.8,0,0.8,0))
plot(plants.tree, cex=1)
par(mar=c(0,0,0,0))
barplot(t(decostand(scent.net, method="tot")), col=sample(rainbow(ncol(scent.net))), names.arg=rep("",nrow(scent.net)), horiz=TRUE) 
```

