---
title: "drought-net"
author: "Janelle and John"
date: "2024-07-12"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(bipartite)
library(vegan)
library(phytools)
library(taxize)
library(brranching)
library(rotl)
library(dendextend)
library(tidyverse)
library(heatmaply)
library(googlesheets4)
library(broom)
```

# Visitation Rate
```{r}
#load metadata

#visitation rate
#loads in pollinator observation google sheet
vrate <- read_sheet("1N3HmtcG3Rv4smOS3-SWdpyHAuH3oLvK2KgzWG4JQmNY", na=c("na","","NA"))  %>% mutate(week_num=factor(week_num)) %>% 
                mutate(insect_class= fct_collapse(insect_class, beetle = c("beetle", "beattle"), lepidoptera= c("butterfly", "checkerspot butterfly", "swallowtail butterfly", "moth"), "solitary bee"= c("solitary bee", "bee", "carpenter bee", fly=c("fly", "hover fly", "hoverfly","mosquito", "syrphid fly", "bee_fly", "beefly", "bee fly")))) %>% 
                      filter(!insect_class%in%c("beetle", "wasp")) %>%   #removed beetles and wasps from dataset
                      mutate(size= fct_collapse(size, big= c("l", "m", "m/l"), small = c("t","s", "t/s","s/t", "s/m"))) %>% 
                      mutate(size = if_else(is.na(size) & insect_class== "solitary bee", "small", size)) %>% #changes NAs in data, assigns NA for sol bee to small size class
                      mutate(size = if_else(is.na(size) & insect_class== "bumblebee", "big", size)) %>% 
                      mutate(size = if_else(is.na(size) & insect_class== "lepidoptera", "big", size)) %>% 
                      mutate(insect_class=na_if(paste(size, insect_class), "NA NA")) #assigns NAs (replaces the NaNs) to rows with no size data
                      
  
                

count(vrate, insect_class, size)

#flower counts
#loads in flower counts for each transect for each day of pollinator observations
flowercounts<- read_sheet("1YR7lBNf5s9Drw9Ix61J8wUtKQO-XuZmXrRMhgktsc5k", sheet="2024") %>% filter(plant!="na") %>%  #2024-07-02, 4f - no flowers open wrote na
        mutate(transectplot=paste0(transect,plot)) %>%
        mutate(week_num = factor(week(date)-26)) %>% 
  group_by(week_num, transect, plot, plant, transectplot) %>% 
  filter(! (week_num== "1"& scribe =="sb")) %>% #removes sb's duplicate flower counts in week 1
  summarize(flower_number=mean(flower_number)) #takes mean of flower counts if there are two flower counts in a transect in the same week

#checks for duplicate flower counts per week 
#in order for this code to work, need to comment out last line that summarize( flower_number = ...) when you read in flowercounts 
#flowercounts %>% count(week_num, transect, plot, plant, date) %>% filter(n>1) %>% left_join(flowercounts) %>% View()
#TODO: check data sheets for sb duplicates on 7/29 and 8/5

#loads in plot and transect codes paired with treatments (rainout, control)
rainout.treatments <- read_sheet("13J7QB2i9mxBsIlbdRu0zL7YKLHL-Wgd-bN3LngkiZww")%>% 
  mutate(plot = tolower(as.character(plot))) #makes plot letter lowercase

```

#theme
```{r}
#sets a theme for ggplot where points that are blue are control, orange is rainout
dnet.color.palette <- c(control="cornflowerblue", rainout="orange")

#TODO: make custom color palette for each plant species
species_colors <- c(
  "erigeron_flagellaris" = "#0072B2",  # Blue
  "eriogonum_subalpinum" = "#D55E00",  # Vermilion (Orange-like but distinct)
  "potentilla_hippiana" = "#F0E442",  # Yellow
  "erigeron_speciosus" = "#CC79A7",  # Reddish Purple
  "eremogone_congesta" = "#56B4E9",  # Sky Blue
  "eriogonum_umbellatum" = "#E69F00",  # Orange (high contrast)
  "collomia_linearis" = "#882255",  # Dark Pink (instead of Black)
  "galium_boreale" = "#999933",  # Olive Green (distinct from other greens)
  "heterotheca_villosa" = "#117733" , # Dark Teal (instead of Brown), 
  "heliomeris_multiflora" = "darkred"
)


```

```{r}
#fills in all non-observed interactions with zeros - makes a matrix of all combinations of insect class and size and then adds zeros to ones with no interaction for the plot/time
#this is because in the data there are NAs when we observed 0 interactions during the 15 min observation period. But those NAs are important becuase it means there was no interaction. Additionally, For each observation window, not seeing a particular insect visitor is just as important as noting which ones we did not see any visits from. Note - this does not account for pollinator phenology. 
vrate.complete<- vrate %>% complete(nesting(date, week_num, start, end, transect, plot, transectplot), plant, nesting(insect_class,size)) %>% 
  #finds all unique combinations of plant-insect visits
                    mutate(visits=replace_na(visits,0)) %>% #replaces all NAs with a zero-> 0 visits
                    drop_na(plant, insect_class) #drops NAs in plant & insect_class - NAs mean we did not get any pollinator visits during observation window

#combine visitation rate and flower counts for each plot
#how many total visits did each plant species get for each plot each week. Last line divides visits by flower number
#TODO: there are some weeks that have poll obs but not flower counts?

vrate.flowercount<- vrate.complete %>% 
                    group_by(week_num, start, end, transect, plot, transectplot, plant, insect_class) %>% #selects columns to keep in dataframe
                    summarise(visits=sum(visits)) %>% #sum of visits for each insect_class to each plant in each plot at a specific date & time
                    left_join(flowercounts) %>% #adds flower counts of each plot for each of these dates (and week_num)
                    drop_na(flower_number) %>% #drops NAs in flower_number, this means that for that species there were no open flowers at that time
                    mutate(visitsperflowerperhr = (visits/flower_number)*4) %>%  #Divides visits by flower number to get visits per flower per hour                
                    mutate(visitsperhr = visits*4) %>%  #makes column with visits per hour
                    mutate(plant_insect_class = paste0(plant, insect_class)) %>% ungroup()

has <- vrate.flowercount %>%  mutate(hascount=!is.na(flower_number)) %>% group_by(week_num, transect, hascount) %>% tally() %>% pivot_wider(names_from= hascount, values_from=n)
#TODO: the following weeks and transects don't have any flower number data, go to flower count data sheet and check to see why theres no flower counts (week 4: transect 3, 6, 7 (can't find these... not entered? or no data?); week 5: transect 5,8 -- WEEK 5 is -25!!!!)


#checks to see if there are visits without flower number for each week
vrate.flowercount.check<- vrate.complete %>% 
                    group_by(week_num, start, end, transect, plot, transectplot, plant, insect_class) %>% #selects columns to keep in dataframe
                    summarise(visits=sum(visits)) %>% #sum of visits for each insect_class to each plant in each plot at a specific date & time
                    left_join(flowercounts) %>% 
                    filter(visits>0, is.na(flower_number)) 
#View(vrate.flowercount.check)

#makes list of all plant species
plantlist <- (unique(flowercounts$plant))

#makes a list of plant species that are in more than 2 transects
majorplant <- flowercounts %>% left_join(rainout.treatments) %>%  count(plant, treatment, transect) %>% #n tells you how many weeks that plant was seen
                    count(plant, treatment) %>%  #n tells you how many transects the plant is seen in 
                      pivot_wider(names_from = treatment, values_from = n, values_fill = 0) %>% #wide format where columns are the treatment names (rainout, contorl), values are the number of transects that plant species is found in, values_fill replaces NAs with 0s
                      filter(control>2, rainout>2) #plant needs to be in more than two transects for control and rainout, filters out any plant species that is in less than two transects

#CAP preparation: 
#turns flowercount into wide format
flowercount.wide <- flowercounts %>% pivot_wider(names_from = plant, values_from = flower_number, values_fill=0) 

#makes dataframe with only plant species and flower number
flowercount.numbers <- flowercount.wide [,plantlist] 

#Adding treatments: 
keep <- rowSums(flowercount.numbers)>4 #keeps plots with more than 4 flowers per plot per week (to make CAP happy)
flowercount.numbers <- flowercount.numbers[keep, ] #stores plots with more than 4 flowers per plot per week to flowercount.numbers replacing old version
flowercount.wide <- flowercount.wide[keep,] %>% left_join(rainout.treatments) #creates wide format to add treatments

#CAP of flower number for each treatment
cap.flowernumber <- capscale(flowercount.numbers~treatment+week_num, data = flowercount.wide)

#anova on the CAP 
anova.cap.flowernumber <- anova(cap.flowernumber, by="term")
#View(anova.cap.flowernumber)
#flower number significant p=0.001; treatment not significant

#within each date, whats the difference between treatments
plot(cap.flowernumber, type="n")
points(cap.flowernumber, col=dnet.color.palette[flowercount.wide$treatment])
text(cap.flowernumber, display="species")

#cap with only treatment - don't use, doesn't include date
#treat.only <- capscale(flowercount.numbers ~treatment, data=flowercount.wide)
#plot(treat.only, type="n")
#points(treat.only, col=dnet.color.palette[flowercount.wide$treatment])
#text(treat.only, display="species")
#text(treat.only, display = "cn", col="magenta")

#nmds to see what species are causing variation in data
#nmds.flowernumber <- metaMDS(flowercount.numbers)

#plot(nmds.flowernumber, type="n")
#points(nmds.flowernumber, col=dnet.color.palette[flowercount.wide$treatment])
#text(cap.flowernumber, display="species")

barplot(t(as.matrix(flowercount.numbers)), col = rainbow(16))


flowercount.summary <- flowercounts %>%  
    filter(!plant %in% c("little white flower", "rock jasmine", "rosa", "agoseris aurantiaca")) %>% 
    left_join(rainout.treatments) %>% 
    group_by(week_num, treatment, plant) %>% summarise(flower_number=sum(flower_number)) 

gg.flowercount <- ggplot(flowercount.summary, aes(y=flower_number, x= paste(week_num, treatment), fill=plant)) +geom_col() + scale_x_discrete(guide = guide_axis(angle=90)) + scale_fill_manual(values = species_colors) + xlab("") + ylab("Flower Number")
print(gg.flowercount)
#Helianthella quinquenervis


ggsave(gg.flowercount, file='flowercount.summary.jpg', width=130, height=105, units= 'mm', dpi=600)

#TODO: clean up repeat species in flowercount.summary


#flower number cap by week 

cap.flowernumber.byweek <- flowercount.wide %>% group_by(week_num) %>% nest() %>% 
                  mutate(model=map(data, ~ capscale(.x[,plantlist] ~ treatment, data=.x))) %>%    
            mutate(test= map(model, ~tidy(anova(.x, by="term")))) %>%  unnest(test)
#TODO: fix plant names in flower count spread sheet
View(cap.flowernumber.byweek)

```

#boxplot for each species and week
```{r}

boxplot.flowercount <- ggplot(flowercounts %>% left_join(rainout.treatments), aes(x=week_num, y=flower_number, color=treatment)) + facet_wrap(vars(plant), scales="free_y") + geom_boxplot() + scale_color_manual(values=dnet.color.palette)

print(boxplot.flowercount)
ggsave(boxplot.flowercount, file='boxplot.flowercount.jpg', width=130, height=105, units= 'mm', dpi=600)
```

#line plot
```{r}
ggplot(flowercounts %>% left_join(rainout.treatments), aes(x=week_num, y=flower_number, color=treatment, group=transectplot)) + facet_wrap(vars(plant), scales="free_y") + geom_line() + scale_color_manual(values=dnet.color.palette)
```

#heatmap - flower number
#TODO: Fix galium with 500 flower count 7b 
```{r}
ggplot(flowercounts %>% left_join(rainout.treatments), aes(x=week_num, y=transectplot, fill=flower_number)) + facet_wrap(vars(plant), nrow = 2, scales="free_y") + geom_tile() + scale_fill_viridis_c() + scale_x_discrete(guide=guide_axis(angle=90))



```

#Visitation Rate

```{r}
#heatmap of vrate and flower count
ggplot(vrate.flowercount %>% left_join(rainout.treatments), aes(x=week_num, y=transectplot, fill=visitsperflowerperhr)) + facet_wrap(vars(plant), nrow = 2, scales="free_y") + geom_tile() + scale_fill_viridis_c() + scale_x_discrete(guide=guide_axis(angle=90))

#
vrate.summary <- vrate.flowercount %>% left_join(rainout.treatments) %>% #adds treament codes to dataframe
                group_by(plant, insect_class, week_num) %>% #selects these columns to include in dataframe
                summarise(visitsperflowerperhr=mean(visitsperflowerperhr)) %>% #takes mean is visits per flower per hr
                mutate(visitsperflowerperhr= if_else(visitsperflowerperhr==0, NA, visitsperflowerperhr)) #replaces zeros with NAs in visits per flower per hr column

#heatmap of vrate summary
ggplot(vrate.summary , aes(x=week_num, y=insect_class, fill=visitsperflowerperhr)) + facet_wrap(vars(plant), nrow = 2, scales="free_y") + geom_tile() + scale_fill_viridis_c() + scale_x_discrete(guide=guide_axis(angle=90))

#compare treatments
#creates a new column from vrate.flowercount called plant_insect_class which lists plant and insect together (see last line) and stores it as a new dataframe
vrate.summary.treatment <- vrate.flowercount %>% left_join(rainout.treatments) %>% 
                        group_by(plant, insect_class, week_num, treatment) %>% 
                        summarise(visitsperflowerperhr=mean(visitsperflowerperhr)) %>%
                        mutate(plant_insect_class = paste0(plant, insect_class)) %>% ungroup()
#removes all entries where visitsperflowerperhr=0, all values must be greater than 0
#the zeros were placeholders, which meant that we did not observe an interaction. We want a dataframe with only observed interactions so that we can plot it and not have a rediculous number of facets (only include facets with non-zero data)
plant.insect.nonzero.list <- vrate.summary.treatment %>% group_by(plant_insect_class) %>%  
                              summarise(visitsperflowerperhr = mean(visitsperflowerperhr)) %>% 
                              filter(visitsperflowerperhr>0)

gg.vrate.summary.line <- ggplot(vrate.summary.treatment %>% filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class), #filters column plant_insect_class but the list of plant x pollinator combinations with visits per flower per hr greater than 0 (which means there is data for them)
            aes(x=week_num, y=visitsperflowerperhr, color=treatment, group = treatment)) + facet_grid(insect_class~plant, scales="free_y") + geom_point() + geom_line()+ scale_color_manual(values=dnet.color.palette)
print(gg.vrate.summary.line)

ggsave(gg.vrate.summary.line, file='vrate.summary.line.jpg', width=375, height=250, units= 'mm', dpi=600)
#anova prep: 

vrate.summary.treatment.transect <- vrate.flowercount %>% left_join(rainout.treatments) %>% #adds treatments
                                      group_by(plant, insect_class, week_num, treatment, transect) %>% #only includes these columns
                                      summarise(visitsperflowerperhr=mean(visitsperflowerperhr)) %>% #takes mean visits per flower per hour
                                      mutate(plant_insect_class = paste0(plant, insect_class), plant_insect_week=paste0(plant_insect_class, week_num)) %>%  #creates column called plant_insect_class by pasting together plant and insect_class, then adds another column called plant_insect_week which pasts together plant_insect_class and week_num
                                      ungroup() #need to ungroup to use in vrate by week cap
  
#number of transects that you have both plant and pollinator data for in control and treatment 
plant.insect.week <- vrate.summary.treatment.transect %>% 
                      filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class) %>% 
                      ungroup() %>% 
                      count(plant_insect_class, treatment, week_num) %>% #only includes columns with plant_insect_class, treatment, week_num and number of transects that plant insect combination are found in (ranges 1-4 transects)
                      pivot_wider(names_from = treatment, values_from = n, values_fill = 0) %>%  #makes treatment column into two separate columns one for control, another for rainout
                      filter(control>0, rainout>0) %>% #filters out plant insect combinations that are not found in any of the transects (must occur in at least one transect)
                      mutate(plant_insect_week= paste0(plant_insect_class, week_num)) %>% #creates a new column with plant, insect_class and week_num pasted together
                      pull(plant_insect_week) #only includes the column plant_insect_week so that we can add it to vrate.summary.treatment.transect


gg.nonzero.vrate.summary.line <- ggplot(vrate.summary.treatment.transect %>% filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class), #filters column plant_insect_class but the list of plant x pollinator combinations with visits per flower per hr greater than 0 (which means there is data for them)
            aes(x=week_num, y=visitsperflowerperhr, color=treatment, group = paste0(treatment, transect))) + facet_grid(insect_class~plant) + geom_point() + geom_line()+ scale_color_manual(values=dnet.color.palette)

print(gg.nonzero.vrate.summary.line)
ggsave(gg.nonzero.vrate.summary.line, file='nonzero.vrate.summary.line.jpg', width=375, height=250, units= 'mm', dpi=600)
#vrate.summary.treatment.transect %>% filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class) %>% ungroup() %>% count(plant_insect_class, treatment) %>% pivot_wider(names_from = treatment, values_from = n) %>%  View()

#anova
#creates a model for each combination of plant and insect group
vrate.anova.model <- vrate.summary.treatment.transect %>% 
                      filter(plant!= "vicia_americana") %>% #vicia is only found in the first week so it breaks to anova model, need to get rid of it
                      filter(plant_insect_week%in%plant.insect.week) %>% #adds visitsperflowerperhr column
                      mutate(week_num=factor(week_num)) %>% #make week_num a factor so anova doesn't break
                      group_by(plant_insect_class) %>% nest() %>% 
                      mutate(model=map(data, ~ lm(visitsperflowerperhr~ treatment*week_num, data=.x))) %>% #makes a model for each plant species x pollinator combination
                      mutate(summary = map(model,~ tidy(anova(.x)))) %>% #summarizes each model made in the line before, tidy makes summary into table
                      select(plant_insect_class, summary) %>%  unnest() #expands dataframe
#WARNING - ignore warning message, just means there's not enough data
#NaN means there is not enough data for the model, you can ignore warning in console this is also the reason (not enough data for some plant insect combos)

View(vrate.anova.model)


  
#TODO: we are having issues with how to compare visitation to a plant species when flowering phenology is different between transects and treatments. 
#issue making sure plant species is flowering/open at the same time in each transect - could restrict the data that the plant has to be open in both transects in order to look at the number of visits
# lump all dates together - boxplot with drought and control, exclude any week that plant is not flowering in both treatments then do a t-test between the two or to get more power lump them all and put in week at random or fixed effect. Model is visitrate against treatment +weeknum
#figure out how to take out weeks that don't have data for both drought and control transects


```

# multivariate analysis on visitation rate
```{r}
#cap by week (6 different caps)
#we did cap by week (instead of cap on whole summer vrate) bc you'd see week effect on whole summer vrate cap but this by week cap says is there a difference by week visitation rate. Maybe there's differences when a certain plant or pollinator is present in the network.

vrate.cap.byweek <- vrate.summary.treatment.transect %>% 
            filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class) %>% #filters column plant_insect_class but the list of plant x pollinator combinations with visits per flower per hr greater than 0 (which means there is data for them)
            pivot_wider(names_from = plant_insect_class, values_from = visitsperflowerperhr, values_fill = 0, id_cols = c(week_num, transect, treatment)) %>% #values_fill=0 fills in zeros for anytime where there is no interaction data, could cause problems bc if transect doesn't have plant you are now setting all those to 0 not bc of treatment but bc there are none of those plants in that transect
            group_by(week_num) %>%  nest() %>% 
            mutate(model=map(data, ~ capscale(.x[,-(1:2)] ~ treatment, data=.x))) %>%   #cap minus the first two cols which are transect and treatment 
            mutate(test= map(model, ~tidy(anova(.x, by="term")))) %>%  unnest(test)

View(vrate.cap.byweek)

```


#For each pollinator group did they make different amounts of visits (to any plant) between treatments? Are there more bees in the drought treatment?
```{r}

ggplot(vrate.summary.treatment.transect %>% filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class) %>% 
                          group_by(insect_class, week_num, treatment, transect) %>% 
                          summarise(visitsperflowerperhr=sum(visitsperflowerperhr)), 
                          aes(x=week_num, y=visitsperflowerperhr, color=treatment, group = paste0(treatment, transect))) + 
                          facet_wrap(vars(insect_class)) + geom_point() + geom_line()+ scale_color_manual(values=dnet.color.palette)
#anova
insect.class.anova <- vrate.summary.treatment.transect %>% 
                      filter(plant!= "vicia_americana") %>% #vicia is only found in the first week so it breaks to anova model, need to get rid of it
                      filter(plant_insect_week%in%plant.insect.week) %>% #adds visitsperflowerperhr column
                      mutate(week_num=factor(week_num)) %>% #make week_num a factor so anova doesn't break
                      group_by(insect_class, week_num, treatment, transect) %>% 
                      summarise(visitsperflowerperhr=sum(visitsperflowerperhr)) %>% 
                      group_by(insect_class) %>% nest() %>% 
                      mutate(model=map(data, ~ lm(visitsperflowerperhr~ treatment*week_num, data=.x))) %>% #makes a model for each plant species x pollinator combination
                      mutate(summary = map(model,~ tidy(anova(.x)))) %>% #summarizes each model made in the line before, tidy makes summary into table
                      select(insect_class, summary) %>%  unnest(summary)

View(insect.class.anova)

```


# For each plant species, is there a difference in pollinator visitation between treatments?

```{r}

ggplot(vrate.summary.treatment.transect %>% filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class) %>% 
                          group_by(plant, week_num, treatment, transect) %>% 
                          summarise(visitsperflowerperhr=sum(visitsperflowerperhr)), 
                          aes(x=week_num, y=visitsperflowerperhr, color=treatment, group = paste0(treatment, transect))) + 
                          facet_wrap(vars(plant)) + geom_point() + geom_line()+ scale_color_manual(values=dnet.color.palette)
#anova
plant.anova <- vrate.summary.treatment.transect %>% 
                      filter(!plant%in% c("vicia_americana", "campanula_rotundifolia")) %>% #vicia is only found in the first week so it breaks to anova model, need to get rid of it
                      filter(plant_insect_week%in%plant.insect.week) %>% #adds visitsperflowerperhr column
                      mutate(week_num=factor(week_num)) %>% #make week_num a factor so anova doesn't break
                      group_by(plant, week_num, treatment, transect) %>% 
                      summarise(visitsperflowerperhr=sum(visitsperflowerperhr)) %>% 
                      group_by(plant) %>% nest() %>% 
                      mutate(model=map(data, ~ lm(visitsperflowerperhr~ treatment*week_num, data=.x))) %>% #makes a model for each plant species x pollinator combination
                      mutate(summary = map(model,~ tidy(anova(.x)))) %>% #summarizes each model made in the line before, tidy makes summary into table
                      select(plant, summary) %>%  unnest(summary)

View(plant.anova)

```

# Is the change in visitation rate predicted by shifts in phenology (flower number)? 
```{r}

ggplot(vrate.flowercount %>% left_join(rainout.treatments) %>% filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class),#filters vrate.flowercount by interactions that have non-zero data
       aes(y=visitsperhr, x=flower_number, color=treatment)) + facet_grid(insect_class~plant, scales="free_y") +geom_point() + geom_smooth(method="lm")+ scale_color_manual(values=dnet.color.palette)


#anova
regression.anova <-vrate.flowercount %>% left_join(rainout.treatments) %>%  
                      filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class) %>% 
                      filter(!plant%in% c("vicia_americana", "campanula_rotundifolia")) %>% #vicia is only found in the first week so it breaks to anova model, need to get rid of it
                      group_by(plant_insect_class, plant) %>% nest() %>% 
                      mutate(model=map(data, ~ lm(visitsperhr~ treatment*flower_number, data=.x))) %>% #makes a model for each plant species x pollinator combination
                      mutate(summary = map(model,~ tidy(anova(.x)))) %>% #summarizes each model made in the line before, tidy makes summary into table
                      select(plant_insect_class, plant, summary) %>%  unnest(summary) %>%  #expands dataframe
                      mutate(adjusted.pvalue= p.adjust(p.value, method="fdr"))

View(regression.anova)
#heterotheca - more flowers get more visits but treatment doesn't have an effect on the average visitation rate per hr (vertical shift) but does effect the slope. There is a difference is visitation rate for the same number of flowers open- the control has more visits even though the drought has same number of open flowers. This could be due to differences in floral rewards, flower density (of all plant species), etc. 


#correction for false discovery rate

# only heterotheca and heliomeris
het.helio.anova <- regression.anova %>%  filter(plant%in%c("heterotheca_villosa", "heliomeris_multiflora"))

View(het.helio.anova)
#het helio plot
ggplot(vrate.flowercount %>% left_join(rainout.treatments) %>% filter(plant_insect_class%in%plant.insect.nonzero.list$plant_insect_class) %>%  filter(plant%in%c("heterotheca_villosa", "heliomeris_multiflora","potentilla_hippiana","erigeron_flagellaris","erigeron_speciosus","collomia_linearis")),#filters vrate.flowercount by interactions that have non-zero data
       
       
       aes(y=visitsperhr, x=flower_number, color=treatment)) + facet_grid(insect_class~plant, scales= "free") +geom_point() + geom_smooth(method="lm")+ scale_color_manual(values=dnet.color.palette)


```

#TODO: clean poll obs dataset -  fix galium 500; what is bee vs solitary bee?


# Network 
```{r}
big.net.long <- vrate %>% group_by(plant, insect_class) %>% summarise(visits=sum(visits)) %>% drop_na(plant) 

big.net <- big.net.long %>% pivot_wider(names_from=insect_class, values_from = visits, values_fill = 0)
  
  
drought.net <- vrate %>%  left_join(rainout.treatments) %>% filter(treatment== "rainout") %>% group_by(plant, insect_class) %>% summarise(visits=sum(visits)) %>% drop_na(plant) %>% pivot_wider(names_from=insect_class, values_from = visits, values_fill = 0) %>%  arrange(plant) %>% #orders rows (plant names) alphabetically
                          column_to_rownames("plant") %>% 
                          relocate(sort(tidyselect::peek_vars()))  #orders columns alphabetically
                          

wet.net <- vrate %>%  left_join(rainout.treatments) %>% filter(treatment== "control") %>% group_by(plant, insect_class) %>% summarise(visits=sum(visits)) %>% drop_na(plant) %>% pivot_wider(names_from=insect_class, values_from = visits, values_fill = 0)  %>%  arrange(plant) %>% #orders rows (plant names) alphabetically
                      column_to_rownames("plant") %>% 
                      relocate(sort(tidyselect::peek_vars()))  #orders columns alphabetically

#make a network for each transect so n=5 for each treatment 
per.transect.networks <- vrate %>% left_join(rainout.treatments) %>% 
                              mutate(transect= if_else(transect==4, 2, transect)) %>%  #adds transect 4 to transect 2 - make sure to remove once all data is entered! 
                              filter(plant%in%majorplant$plant) %>% group_by(plant, insect_class, transect, treatment) %>% summarise(visits= sum(visits)) %>% 
                              drop_na(plant) %>% group_by(transect, treatment) %>% nest() %>% 
                              mutate(network = map(data, ~pivot_wider(.x, names_from=insect_class, values_from = visits, values_fill = 0)  %>%  arrange(plant) %>% #orders rows (plant names) alphabetically
                      column_to_rownames("plant") %>% 
                      relocate(sort(tidyselect::peek_vars())))) #map takes a column and gives back a list
  
```

# Accumulation curve 
```{r}

poll.interaction <- vrate %>% drop_na(plant) %>% mutate(plant_insect_class= paste0(plant, insect_class)) %>%  select(plant_insect_class, visits) %>% mutate(index=row_number()) %>% 
    pivot_wider(names_from=plant_insect_class, values_from = visits) %>% mutate(across(everything(),~replace_na(.x, 0))) %>% select(-index) %>% as.matrix()

spec<- vegan::specaccum(poll.interaction,permutations=100, method="exact")
plot(spec) #y is unique plantxpoll combos, x is number of rows which is like a 15min block; approx 2 interactions per 15min block

```


# p-p web plot

```{r}

plant.pollinator.network.drought <- plotweb(sqrt(drought.net), empty=F, text.rot=90, method="normal")

plant.pollinator.network.wet <- plotweb(sqrt(wet.net), empty=F, text.rot=90, method="normal")

```

# Network Stats

```{r}
drought.net.stats <- networklevel(drought.net) 

wet.net.stats <- networklevel(wet.net)

all.network.stats <- bind_rows(as.data.frame(t(drought.net.stats)), as.data.frame(t(wet.net.stats))) %>% t() %>% as.data.frame() %>% 
                            rename("drought"= V1, "control"= V2)
                          
#transect networks are too small
by.transect.net.stats <- per.transect.networks %>% mutate(stats=map(network, ~ as.data.frame(t(networklevel(.x))))) %>%  unnest(stats)

#plot network stat by transect and treatment
ggplot(by.transect.net.stats %>% pivot_longer(colnames(by.transect.net.stats)[5:52]), aes(y=value, x=treatment, color=treatment)) + facet_wrap(vars(name), scales="free_y") + geom_point()

#run an anova on network stats you're interested in


```

# down sample network 

```{r}
#down sampling allows you to have same number of interactions in both control and drought so you can compare the two networks more easily
#take drought.net and wet.net and take random rows

#number to down sample to

downsample.rows <- 220 #number of rows in down sampled networks

#make mini networks 

downsample.function <- function(manipulation, downsample.rows){
                          vrate %>%  left_join(rainout.treatments) %>% filter(treatment== manipulation) %>% 
                          slice_sample(n=downsample.rows) %>% 
                          group_by(plant, insect_class) %>% summarise(visits=sum(visits)) %>% drop_na(plant) %>% 
                          pivot_wider(names_from=insect_class, values_from = visits, values_fill = 0) %>%  
                          arrange(plant) %>% #orders rows (plant names) alphabetically
                          column_to_rownames("plant") %>% 
                          relocate(sort(tidyselect::peek_vars())) %>%  #orders columns alphabetically
                          networklevel(index=c("connectance", "NODF", "H2", "links per species")) %>% 
                          t() %>% data.frame()
            
                        }

stats.downsample <- tibble(permutation=rep(1:999, times=2), treatment=rep(c("control", "rainout"), each=999)) %>% 
              mutate(stats= map(treatment, ~downsample.function(.x, downsample.rows))) %>% unnest(stats)

# Links per species plot (species = insect class and size)
ggplot(stats.downsample, aes(x=links.per.species, fill= treatment))+ geom_histogram(binwidth = 0.05) + scale_fill_manual(values=dnet.color.palette) + facet_wrap(vars(treatment), ncol = 1)

#t test
#don't use pvalue bc we are just choosing how much data to create with number of permutations
anova(lm(links.per.species~treatment, data=stats.downsample))

```


